### 2021-10-06 code/censusland_code.Rmd
Added census FIPS code + land cover data

---
title: "census-land-data"
output: html_document
---

##Load Dependencies
The following code loads the package dependencies for our analysis:
```{r}
#install.packages("tidycensus", repos = "http://cran.us.r-project.org")
#install.packages("censusapi", repos = "http://cran.us.r-project.org")
#install.packages("censusGeography", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/chron_2.3-55.tgz")
#install.packages("rerddap", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/rerddap_0.6.5.tgz")

library(tidyverse)
library(dplyr) #data wrangling
library(tidycensus) #access census data
library(censusapi)
library(rerddap)
library(RCurl)

#path to data
data_location <- here::here("data","processed_data","cleansampledata.rds")

#load data. 
data <- readRDS(data_location)
```

##Load Census API (load only once)
```{r}
#specify from which package a function is to be used

#tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE) 
```
^This is not a code chunk you will need in every notebook. As long as "install = TRUE", you will only have to do this once. 
#restart R session after installing the first time

##Obtain FIPS codes for sampling sites
```{r}
#create function
latlong2fips <- function(latitude, longitude) {
  url <- "https://geo.fcc.gov/api/census/block/find?format=json&latitude=%f&longitude=%f"
  url <- sprintf(url, latitude, longitude)
  json <- RCurl::getURL(url)
  json <- RJSONIO::fromJSON(json)
  as.character(json$County['FIPS'])
}
#vectorize so function will take more than one argument
latlong2fips_vec <- Vectorize(latlong2fips, vectorize.args = "latitude", "longitude")
```

#accessing NHGIS land use/land cover data from downloaded file
source:https://www.nhgis.org/user-resources/environmental-summaries#download
```{r}
#path to data
data_location1 <- here::here("data","raw_data","landcover2015.csv")

#load data. 
landcover <- read.csv(data_location1)

class(landcover$GISJOIN)
#class = character, so I can use strsplit(). Data source uses extra letters and zeroes in their GIS code. I need to extract the original FIPS code to match my data

#extract state fip from GISJOIN
landcover = landcover %>% mutate(state_fip = substring(landcover$GISJOIN, 2,3)) %>% relocate(state_fip)

#extract county fip from GISJOIN
landcover = landcover %>% mutate(county_fip = substring(landcover$GISJOIN, 5, 7)) %>% relocate(county_fip)

#combine to state, county fip for whole fips codes
landcover$fips = paste0(landcover$state_fip, landcover$county_fip)

landcover <- landcover %>%
  select(fips, everything()) # moving fips column to far left of data set

#filter to clean data set
landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID)

athensfips <- c("13219", "13059")
athensland <- filter(landcoverclean, fips %in% athensfips)

```

### 2021-10-06 code/processing_code/processingscript.R

@@ -8,6 +8,7 @@
 library(readxl) #for loading Excel files
 library(dplyr) #for data processing
 library(here) #to set paths
+library(tidyverse)
 
 #path to data
 data_location1 <- here::here("data","raw_data","mpdata_ALL.xlsx")
@@ -37,10 +38,34 @@ glimpse(rawdata)
 # the notes column won't be neccessary for data analysis. Columns 16, 17, and 18 
 # are also unnecessary. Let's remove those. 
 
-processeddata <- rawdata %>% select(-"notes", -(16:18))
+processeddata <- rawdata %>% select(-"notes", -(17:19))
 
 # assessing missing values
 is.na(processeddata) %>% summary()
+
+# Since there are just a few observations that are missing location (watershed, lat, and long)
+# and/or count data, and there is not a great way to impute them. Since I am working
+# with data that I collected myself, I know that those observations simply didn't
+# get counted, got lost, or site names were misread/input incorrectly. Let's clean 
+# up the data set. 
+
+processeddata <-subset(processeddata, site!="MIDO 504" & site!="MIDO 503")
+
+# I've removed the rows by name instead of filtering out rows with missing watershed information.
+# This is because I want to retain the "control" rows for now, which originated 
+# in the lab and don't have a corresponding sample location
+
+# Another thing I can do is tidy the data by filling in all "id" information (all missing ids for previous
+# samplings would be me (a)). 
+
+processeddata$id[is.na(processeddata$id)]<-"a"
+
+# Next, let's remove observations where the samples were originally recorded in the
+# data sheet, but apparently never got counted during sample processing. Again, 
+# we just want to remove sample type observations and not control type observations. 
+
+sampledata <- processeddata %>% filter(obs_type == "sample") %>% subset(a_count!="NA" & b_count!="NA")
+controldata <- processeddata %>% filter(obs_type == "control")
 
 # There are a lot of missing values for the c_count and d_count. These 
 # variables resulted from the fourth quarterly sampling in the dataset, when
@@ -51,29 +76,45 @@ is.na(processeddata) %>% summary()
 # replicates was counted twice (once by two different volunteers) in order to 
 # assess the efficacy of the training program/reproducability of counts. 
 
-# Some options: 
+# Ideas to address this: 
 # 1. Subset the data into one group of all samples with just a_count
 # and b_count and another group with counts a, b, c, and d for the July sampling
[…]

### 2021-10-07 code/processing_code/uownprocessing.Rmd
added UOWN data
updated README, more data cleaning

---
title: "uownprocessing"
output: html_document
---
This script loads the raw UOWN data, processes and cleans it 
and saves it as Rds file in the processed_data folder. While the microplastics data is the focus of my project, I am interested in comapring UOWN's data on bacteria levels with the recorded microplastic levels at samples sites throughout the study timeframe.

# load packages
```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
```

# path to data & load data
```{r}
data_location1 <- here::here("data","raw_data","UOWN_spring2021.csv")

uowndata <- read.csv(data_location1)
```

# Take a look at data
```{r}
glimpse(uowndata)
```

# Subsetting UOWN data
```{r}
uowndata <- uowndata %>% filter((year == 2020 & month == 11) | (year == 2021 & quarter == 1) | (year == 2021 & quarter == 2 )| (year == 2021 & quarter == 3))
```

Now we have the UOWN data from just the samples dates that we also have the microplastic data for. Right now, only Nov 2020, Feb 2021, and April 2021 data are available from the UOWN website, but July 2021 data should be available soon and I will add it in whenever it is posted. The code chunk above should already include the code to load the July data when the updated file is available.

# Cleaning up the data
```{r}
cleanuowndata <- uowndata %>% select(-"ecoli.method.known", 
                                     -"biological_score",
                                     -"po4.mgL", 
                                     -"conductivity.uscm",
                                     -"no3.mgL",
                                     -"pH") %>%
  rename(site=WSID)

```

# Save UOWN data alone
```{r}
# save data as RDS

# location to save file
save_data_location <- here::here("data","processed_data","cleanuowndata.rds")

saveRDS(cleanuowndata, file = save_data_location)
```

# Combine UOWN data with microplastic sample data
```{r}
# load MP data
data_location <- here::here("data","processed_data","cleansampledata.rds")
cleansampledata <- readRDS(data_location)

# edit site column to match UOWN data
# remove white space in site name 
cleansampledata <- cleansampledata %>% mutate(site = 
                                     gsub(" ","",cleansampledata$site))

# join data sets by site name (cleansampledata) / WSID (cleanuowndata)
joindata <- left_join(cleansampledata, cleanuowndata, by = "site")

# we joined the data by site, but we also need to join it by date. we'll have to clean up some of the columns in order to do that

# edit date format by creating new columns for month and year
cleansampledata <- cleansampledata %>% mutate(year = ifelse(date == "nov_20", "2020", "2021")) %>% mutate(month = ifelse(date == "nov_20", "11", 
                                                                                                              ifelse(date == "feb_21", "2", 
                                                                                                                     ifelse(date == "apr_21", "4", 
                                                                                                                            ifelse(date == "july_21", "7", "no")))))
cleansampledata$month <- as.numeric(cleansampledata$month)
cleansampledata$year <- as.numeric(cleansampledata$year)
# try joining again
joindata <- left_join(cleansampledata, cleanuowndata, by = c("site", "month", "year"))

# clean up joined dataset
joindata <- joindata %>% select(-"day",
                                -"WS", 
                                -"ID")
# save combined data as RDS
save_data_location <- here::here("data", "processed_data", "combodata.rds")

saveRDS(joindata, file = save_data_location)

```

### 2021-10-07 code/processing_code/mpprocessing.R

@@ -14,7 +14,7 @@ library(tidyverse)
 data_location1 <- here::here("data","raw_data","mpdata_ALL.xlsx")
 data_location2 <- here::here("data","raw_data","uown 8.26.xlsx")
 
-#load data. 
+#load microplastics data. 
 rawdata1 <- read_excel(data_location1)
 rawdata2 <- read_excel(data_location2)
 
@@ -117,4 +117,20 @@ saveRDS(cleansampledata, file = save_data_location)
 saveRDS(volunteer, file = save_data_location1)
 saveRDS(controldata, file = save_data_location2)
 
+# Here, I'll also load some data we'll want to look at later. The csv file being 
+# imported contains the coordinates for Athens' three water reclamation facilities (WRF)
+# Since WRF efflunet is a hypothesized predictor of microplastic levels, we'll examine
+# this later in the exploratory analysis. 
 
+# load Water Reclamation Facility location data
+
+#path to data
+data_location3 <- here::here("data","raw_data","wrflocation.csv")
+
+#load WRF data
+wrflocation <- read.csv(data_location3)
+
+#save data as RDS
+save_data_location3 <- here::here("data", "processed_data", "wrflocation.rds")
+
+saveRDS(wrflocation, file = save_data_location3)

### 2021-10-08 code/analysis_code/exploration.Rmd
project part 2 draft

---
title: "exploration"
output: html_document
---

For this project, I began with microplastic count data for various sample sites, variables including site, watershed, latitude, longitude, and a_count and b_count. So far, I have obtained the following additional data:

 - local land cover based on coordinates converted to census FIPS codes, using NGIS data
 - bacterial data from corresponding UOWN sample dates/sites
 - location population levels by zip code, converted from lat/long and joined with census data.

In this document, I will explore relationships between some of these variables.

Quantities of main interest: 
 - particles per liter of microplastics
 - season/sample date
 - population level
 - proximity to water reclamation facility effluent
 - bacteria levels at corresponding samples sites/dates

# load packages and data
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(ggplot2)

#path to data
data_location <- here::here("data","processed_data","popdata.rds")

#load data
mpdata <- readRDS(data_location)

```

# take a look at data
```{r}
summary(mpdata)

# view particles/L by season

mpdata %>% ggplot(aes(x = month, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")

```

### 2021-10-08 code/processing_code/censusprocessing.Rmd
project part 2 draft

---
title: "censusprocessing"
output: html_document
---

At this point, I have written a few processing files loading the microplastics data, land cover/use data, and UOWN data for this project. I'd also like to look at small area population. I think I can achieve this by looking at population by zip code in Athens Clarke County, which is available from the US Census. My goal is to convert the latitude and longitude coordinates for each sample site into a zip code, and then examine whether microplastic levels vary based on local population size/population density.

The data will come from the American Community Survey 5-year estimate 2019, and I will plan to use API. To convert sample coordinates to comparable zip codes, I can use the `tidygeocoder` package.

Citation for zipcode population data (saving for later): 
United States Census Bureau. B01001 SEX BY AGE, 2019 American Community Survey 5-Year Estimates. U.S. Census Bureau, American Community Survey Office. Web. 10 December 2020. http://www.census.gov/.

Downloaded at: https://www.georgia-demographics.com/zip_codes_by_population

```{r}
#load  packages
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(revgeo) #for reverse geocoding coordinates to zipcode

#path to data
data_location <- here::here("data","raw_data","zippop.xlsx")

#load population data. 
zippop <- read_excel(data_location)

#load MP data
data_location1 <- here::here("data","processed_data","combodata.rds")

#load data. 
combodata <- readRDS(data_location1)
```

# convert lat long to zip
```{r}

# use Bing API to convert coordinates to corresponding zip code
zip<-revgeo(longitude=combodata$long, 
                            latitude=combodata$lat, 
                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip")

# add zip variable to combodata
combodata <- combodata %>% mutate(revgeo(longitude=combodata$long, 
                            latitude=combodata$lat, 
                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip") %>%
                              select("zip"))
```

# add population levels based on zip code
```{r}
popdata <- left_join(combodata, zippop, by = c("zip" = "zipcode"))

#save data as RDS
popdata_loc <- here::here("data", "processed_data", "popdata.rds")

saveRDS(popdata, file = popdata_loc)

```

### 2021-10-08 code/analysis_code/exploration.Rmd
project pt 2 finishing touches

@@
 
-mpdata %>% ggplot(aes(x = month, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")
+mpdata %>% ggplot(aes(x = date, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")
 
 ```
+
+ Here, we can see that the levels of microplastics are similar, from a wide lens, through the four sample dates.
+There are a few outliers - it would be interesting to look at the data when excluding outliers.
+I'll have to look into what the best practice for that is.
+Also, for final products, I'll want to reorder the dates so that Nov 2020 is listed first (chronologically), and I will relabel everything to make it neater as well.
+
+ For all samples, let's compare microplastic level and local population.
+
+# MP level vs population
+```{r}
+mpdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point()
+```
+
+I'm not seeing any obvious relationships here.
+Also, population values are appearing discrete rather than continous since there are a limited number of zip codes in the study area, and population was based on zip code.
+
+```{r}
+mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
+```
+
+Something weird is giong on with the values for e.coli.cfu.
+
+```{r}
+class(mpdata$e.coli.cfu)
+```
+
+It looks like that ended up saved as a character instead of integer. Let's convert. 
+# covert character to numeric
+```{r}
+mpdata <- mpdata %>% mutate(e.coli.cfu = as.numeric(e.coli.cfu))
+class(mpdata$e.coli.cfu)
+```
+
+# try plot again
+```{r}
+mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
+```
+
+Again, no obvious relationships here. How about if we looked with a logarithmic scale?
+
+```{r}
+mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1)) + scale_x_continuous(trans='log10') +
+  scale_y_continuous(trans='log10')
+```
+
+Note: missing values are for e.coli.cfu for July 2021 sampling. These values should be available for comparison soon.
+
+Let's look at overall distribution of MP levels
+```{r}
+mpdata %>% ggplot(aes(x = particles_l)) + geom_histogram()
+```
+
+We have some large outliers that make it harder to view the exact distribution of the lower values. Let's filter out the outliers for a moment.
+
+```{r}
+mpdata %>% filter(particles_l < 400) %>% ggplot(aes(x = particles_l)) + geom_histogram()
+```
+
+We see that most levels fall below 100 particles per liter.
+
+```{r}
+summary(mpdata$particles_l)
+```
+
+These summary statistics give us a little more descriptive information about the dataset.
+The other variables that I wanted to look at particles per liter alongside were land cover.
+The process for looking at land cover will be a little more complicated - there a key for variable names corresponding to land types that I need to look closer at, and I am also not sure if the scale of the land cover data will be appropriate.
+Currently, it is at the census FIPS code level (approximately county level).
+Since most of these samples occur within that same area, we don't have other reference areas for comparison.
+I'll have to think about what I can do with this information.
+
+I also wanted to look at proximity of sample sites to water reclamation facilities (WRFs). Let's try that. 
+# load WRF location
+```{r}
+#path to data
+wrfdata_location <- here::here("data","processed_data","wrflocation.rds")
+
+#load data
+wrflocation <- readRDS(wrfdata_location)
+```
+
+Eventually, I would like to be able to view this information on a map - where WRFs are labelled, and samples sites are color-coded by density of microplastics found.
+For now, let's see if we can calculate distance otherwise.
+
+```{r}
+library(geodist)
+
+#calculate distance between sample coordinates and WRF location
+wrfdistance <- geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat)
+```
+
+I think the product we have here is a matrix of the distance of each sample site from each of the three WRFs, in the order that they appear in the WRF data (Cedar Creek, North Oconee, Middle Oconee).
+But, the labelling of the data is not super clear, so I need to verify this to be sure.
+
+```{r}
+test <- mpdata %>% mutate(wrfdist = geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat))
+```
+
+This produced another column like I wanted, but I don't know which WRF the distance was calculated for.
+Maybe I need to save WRF locations as individual lat and long values, and produce three new columns that way?
+However, I am not sure what the final product would look like if we do that, because a sample site could be closely associated with one WRF but much further from another one.
+Maybe I need to narrow this down to proximity to the nearest WRF, so that each sample site is only associated with the distance to ONE WRF, not three.

### 2021-10-08 code/processing_code/censusprocessing.Rmd
project pt 2 finishing touches

@@
 
+This function produces the zip codes I am looking for, but there are observations with missing variables that didn't return addresses and zip codes.
+I'll have to work on troubleshooting this.
+I may be able to manually input the zip codes, but that is not ideal.
+
 # add population levels based on zip code
 ```{r}
+#add zipcode information to my original combined dataset
 popdata <- left_join(combodata, zippop, by = c("zip" = "zipcode"))

### 2021-10-08 products/manuscript/Manuscript.Rmd
project pt 2 finishing touches

---
title: "Microplastics in the Urban Environment"
subtitle: "Exploratory Study of Freshwater Microplastics in Athens, GA"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
#bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found). Removing this line for now so that Manuscript will knit. Will continue to troubleshoot and fix bib formatting. 
csl: ../apa.csl
---

This document will be my final manuscript for the MADA project. I am keeping these draft sections for structure now, until I edit them later.

# Summary/Abstract
_Write a summary of your project._

# Introduction

In the past two decades, the study of microplastics has expanded exponentially. Since the earliest paper mentioning small plastic particles in the ocean was published in the 1970s, before the term “microplastics” even existed, researchers across the globe have quantified microplastic particle levels in both marine and freshwater environments, assessed potential drivers and predictors of microplastic pollution, examined the presence of microplastics in the food chain, and explored the potential human health impacts of ingested microplastics [@Carpenter1972]. Though the body of research on microplastics has grown in recent years, many knowledge gaps remain regarding this pollutant. The term microplastics was first coined in 2004, and today microplastics are defined as plastic particles smaller in size than 5 mm, though some studies examine particles as large as 20 mm [@Thompson2004]. Years of sampling, experimentation, and progress in research methods tell researchers that microplastics are a persistent class of pollutant and are found worldwide in a diverse assemblage of forms [@Barboza2018].

Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014][@Waller2017][@Peng2018]. Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era [@Zalasiewicz2016]. These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017][@Barboza 2018][@Prokic2019][@Evereart2018]. In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans [@Barboza2018]. Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011][@Li2018]. Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014][@Li2018][@Browne2010][@Lebreton2017]. However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2018][@Dris2015]. Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

# Methods and Results

_In most research papers, results and methods are separate. You can combine them here if you find it easier. You are also welcome to structure things such that those are separate sections._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Exploratory analysis

_Use a combination of text/tables/figures to explore and describe your data. You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. FIgures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data._

_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables. Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones._

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

Table \@ref(tab:summarytable) shows a table summarizing the data.

```{r summarytable,  echo=FALSE}
#saving template for formatting later on

#resulttable=readRDS("../../results/summarytable.rds")
#knitr::kable(resulttable, caption = 'Data summary table.')
```

Figure \@ref(fig:resultfigure) shows a scatterplot figure produced by one of the R scripts.

```{r resultfigure,  fig.cap='Analysis figure.', echo=FALSE}
#knitr::include_graphics("../../results/resultfigure.png")
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example table \@ref(tab:resulttable) shows a table summarizing a linear model fit.

```{r resulttable,  echo=FALSE}
#resulttable=readRDS("../../results/resulttable.rds")
#knitr::kable(resulttable, caption = 'Linear model fit table.')
```

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

# References

### 2021-10-08 code/analysis_code/analysisscript.R

@@ -16,54 +16,9 @@ data_location <- here::here("data","processed_data","processeddata.rds")
 #load data. 
 mydata <- readRDS(data_location)
 
-######################################
-#Data exploration/description
-######################################
-#I'm using basic R commands here.
-#Lots of good packages exist to do more.
-#For instance check out the tableone or skimr packages
-
-#summarize data 
-mysummary = summary(mydata)
-
-#look at summary
-print(mysummary)
-
-#do the same, but with a bit of trickery to get things into the 
-#shape of a data frame (for easier saving/showing in manuscript)
-summary_df = data.frame(do.call(cbind, lapply(mydata, summary)))
-
-#save data frame table to file for later use in manuscript
-summarytable_file = here("results", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+#this is where my analysis script for the MADA project will be located. Keeping
+#this file as a placeholder for now. 
 
 
-#make a scatterplot of data
-#we also add a linear regression line to it
-p1 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-
-#look at figure
-plot(p1)
-
-#save figure
-figure_file = here("results","resultfigure.png")
-ggsave(filename = figure_file, plot=p1) 
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-# fit linear model
-lmfit <- lm(Weight ~ Height, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable <- broom::tidy(lmfit)
-
-#look at fit results
-print(lmtable)
-
-# save fit results table  
-table_file = here("results", "resulttable.rds")
-saveRDS(lmtable, file = table_file)
 
   

### 2021-10-08 code/analysis_code/exploration.Rmd
final updates project pt2

@@
 
-mpdata %>% ggplot(aes(x = date, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")
+p1 <- mpdata %>% ggplot(aes(x = date, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")
 
+plot(p1)
+
+#save figure
+figure_file = here("results","concentrationbydate.png")
+ggsave(filename = figure_file, plot=p1)
 ```
@@
 ```{r}
-mpdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point()
+p2 <- mpdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point()
+
+plot(p2)
+
+#save figure
+figure_file = here("results","particlesvpop.png")
+ggsave(filename = figure_file, plot=p1)
 ```
@@
 ```{r}
-mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
+p3 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
+
+plot(p3)
+
+#save figure
+figure_file = here("results","mpvbacteria.png")
+ggsave(filename = figure_file, plot=p3)
 ```
@@
 ```{r}
-mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1)) + scale_x_continuous(trans='log10') +
+p4 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1)) + scale_x_continuous(trans='log10') +
   scale_y_continuous(trans='log10')
+
+plot(p4)
+
+#save figure
+figure_file = here("results","logmpvbacteria.png")
+ggsave(filename = figure_file, plot=p4)
 ```
@@
 ```{r}
-mpdata %>% filter(particles_l < 400) %>% ggplot(aes(x = particles_l)) + geom_histogram()
+p5 <- mpdata %>% filter(particles_l < 400) %>% ggplot(aes(x = particles_l)) + geom_histogram()
+
+plot(p5)
+
+#save figure
+figure_file = here("results","distribution.png")
+ggsave(filename = figure_file, plot=p5)
 ```
@@
 summary(mpdata$particles_l)
+
 ```

### 2021-10-08 products/manuscript/Manuscript.Rmd
final updates project pt2

@@
 ## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
+
+The data that I am using for this project is from my personal undergraduate research project on freshwater microplastics in Athens, GA.
+The study is ongoing and was conducted in collaboration with the Upper Oconee Watershed Network (UOWN) via their quarterly sampling events.
+There is one year’s worth of quarterly data, with the following sample dates: November 2020, February 2021, April 2021, and July 2021.
+The data include variables such as sample site, latitude and longitude, watershed location, and duplicate A and B counts of microplastics observed per filter.
+For the most recent sampling date, July 2021, there is additional data where a second counting session was performed, in order to assess the comparability of parallel counts.
+There are about 136 observations of ~10 variables in total in the dataset.
+The independent variables – site location, coordinates, watershed location, etc – were retrieved from UOWN.
+Values for microplastic counts were recorded based on visual identification via a dissecting microscope, performed by myself and by a small team of <10 volunteers.
+
+Additional data [write up further explanation]:
+ - small area population via census zipcode
+ - UOWN bacteria level data
+ - local land cover
+ - distance from water reclamation facility (WRF)
 
 ## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
+The questions that I want to answer with this data include the following:
+ - Can we characterize microplastics in Athens, GA using a citizen science approach? 
+ -	What are the levels of microplastics like in Athens, GA overall? 
+ -	How do microplastic levels differ based on location within the region?
+
+Questions that I want to answer that will require gathering more data using coordinates/watershed info include:
+ -	Is microplastic concentration impacted by local land use? 
+ -	Is microplastic concentration impacted by localized population levels in specific neighborhoods?
+ -	Does wastewater treatment plant effluent from nearby facilities increase microplastic concentration at affected sites?
+ -	Are microplastic levels correlated to bacteria levels measured at the same sites? 
+These questions also relate to hypothesized predictors of microplastic pollution: population level, land use, wastewater treatment plant effluent, and microbial water quality.
+
+My ideas for how to analyze this data include using non-parametric statistical tests, correlation matrices, and modeling (?) to characterize and compare observations.
 
@@
 
+- Methods for sample collection and processing for UOWN MP water samples
+
 ## Data import and cleaning
@@
 
+Since I am performing importing and cleaning in a separate document, in this section I will provide an overview of my steps/methods for wrangling the data, without actual code in this section.
+
+Import of additional data
+ - census - American Community Survey
+ - NGIS land cover
+ - zip code - Bing reverse geocoding
+ - WRF locations: located on Google Maps, coordinates recorded in Excel spreadsheet
+ - UOWN bacteria data (citation/attribution available on UOWN.org)
+
 ## Exploratory analysis
 
+This section will contain the key products of my exploratory analysis (located in exploration.Rmd) when complete.
+Since I am keeping this document in manuscript style, I will keep processing/exploration/analysis code separate.
 _Use a combination of text/tables/figures to explore and describe your data.

### 2021-10-26 analysis.Rmd
updated pop data, wrf data, land cover FIPS

---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2021-10-26 code/analysis_code/exploration.Rmd
updated pop data, wrf data, land cover FIPS

@@
 Maybe I need to narrow this down to proximity to the nearest WRF, so that each sample site is only associated with the distance to ONE WRF, not three.
+
+#Working on gathering WRF distance
+```{r}
+#set individual WRF locations and save as values
+cedarcreeklat = 33.87351 
+cedarcreeklong = -83.33639
+northoconeelat = 33.93402
+northoconeelong = -83.36018
+middleoconeelat = 33.90862
+middleoconeelong = -83.39214
+
+mpdata <- mpdata %>% mutate(ccwrf = geodist_vec(mpdata$long, mpdata$lat, cedarcreeklong, cedarcreeklat))
+mpdata <- mpdata %>% mutate(nowrf = geodist_vec(mpdata$long, mpdata$lat, northoconeelong, northoconeelat))
+mpdata <- mpdata %>% mutate(miwrf = geodist_vec(mpdata$long, mpdata$lat, middleoconeelong, middleoconeelat))
+
+#Since I only care about the proximity to the closest WRF, my next task is to write code that creates a new variable returning only the lowest value out of ccwrf, nowrf, and miwrf for each sample site.
+
+#subset data into just distance from each WRF
+wrf <- mpdata %>% select("ccwrf", "nowrf", "miwrf") 
+#find minimum distance to any WRF out of all three
+val <- do.call(pmin, c(wrf, na.rm = TRUE)) 
+minwrf <- transform(wrf, min=val, minCol = names(wrf)[max.col(wrf == val, 'first')])
+
+#minwrf contains the variable min, which tells us the distance from the sample site to the nearest WRF. This is all of the information that we need, so no we will re-join to the original DF
+mpwrfdata <- left_join(mpdata, minwrf)
+
+#remove variables we won't need anymore
+mpwrfdata <- mpwrfdata %>% select(-"ccwrf", 
+                                  -"nowrf",
+                                  -"miwrf")
+#rename columns
+mpwrfdata <- mpwrfdata %>% rename("dist" = "min", 
+                     "wrf" = "minCol")
+
+#save as RDS
+saveRDS(mpwrfdata, file = here("data", "processed_data", "mpwrfdata.rds"))
+```
+
+Update 10/25/21
+I've updated the population data by accessing census tract level population rather than zip code level population.
+This should give us more localized population levels and might paint a better picture of the relationship between MP particles per liter and local population level.
+
+#Adding WRF data to updated population data
+```{r}
+#load data
+tractpopdata <- readRDS(here("data", "processed_data", "tractpopdata.rds"))
+
+#add wrf data to tractpopdata
+wrfjoin <- mpwrfdata %>% select("site", "date", "wrf", "dist") #selecting just the variables we want to add
+mpwrfcombo <- left_join(tractpopdata, wrfjoin, by = c("site", "date")) #joining the two datasets
+
+#save as RDS
+saveRDS(mpwrfcombo, file = here("data", "processed_data", "mpwrfcombo.rds"))
+```
+
+Now, `mpwrfcombo` contains the sample info and the new WRF distance info.
+However, something to work on going forward is that a sample site may be near a WRf but not downstream of it.
+It might be more important to know the distance to the nearest upstream WRF.
+I'll have to go back and consider this as I continue my exploration/analysis.
+
+Other than not including landcover data, `mpwrfcombo` is the most comprehensive dataset, including almost all of the data that I have compiled so far.
+
+# Plotting population vs particles per liter w/ new data
+```{r}
+tractpopdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point() 
+```
+
+With this plot, we still don't see an obvious relationship between particles per liter and population level, even with our updated population levels.
+
+```{r}
+tractpopdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point() + scale_x_continuous(trans='log10') +
+  scale_y_continuous(trans='log10')
+```

### 2021-10-26 code/processing_code/censusland_code.Rmd
updated pop data, wrf data, land cover FIPS

--- 2021-10-06 code/censusland_code.Rmd
@@
 
+#clean dataset by selecting only data for the state of Georgia
+landcover <- landcover %>%
+  filter(state_fip == "13")
+
+#get rid of the 4th and the 8th zero in GISJOIN to make compatible with FIPS codes for samples + population count
+landcover = landcover %>% mutate(FIPS = paste(state_fip, county_fip, substring(landcover$GISJOIN, 9, 14), sep = "")) %>% relocate(FIPS)
+
 #filter to clean data set
-landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID)
+landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID, fips)
 
-athensfips <- c("13219", "13059")
-athensland <- filter(landcoverclean, fips %in% athensfips)
+#Now we have landcover data in a form that is compatible with our sample site data (using small geography census info).
+Let's join datasets to create a dataset with sample site info and land cover info
+tractpopdata <- readRDS(here("data", "processed_data", "tractpopdata.rds"))
+athensland <- left_join(tractpopdata, landcoverclean, by = "FIPS")
 
+#save as RDS
+saveRDS(athensland, file = here("data", "processed_data", "athensland.rds"))
 ```

### 2021-10-26 code/processing_code/censusprocessing.Rmd
updated pop data, wrf data, land cover FIPS

@@
 ```
+
+Update 10/25/21
+
+I am realizing that we can get more localized population levels by using census tracts, which are smaller levels of census geography.
+Let's try to use FIPS codes (which include tract info) to obtain population data that might be more informative.
+# Converting coordinates to FIPS
+```{r}
+#create function
+latlong2fips <- function(latitude, longitude) {
+  url <- "https://geo.fcc.gov/api/census/block/find?format=json&latitude=%f&longitude=%f"
+  url <- sprintf(url, latitude, longitude)
+  json <- RCurl::getURL(url)
+  json <- RJSONIO::fromJSON(json)
+  as.character(json$Block['FIPS'])
+}
+#vectorize so function will take more than one argument
+latlong2fips_vec <- Vectorize(latlong2fips, vectorize.args = "latitude", "longitude")
+
+fips <- popdata %>% rowwise() %>% mutate(FIPS = latlong2fips_vec(lat, long))
+
+saveRDS(fips, "fips.RDS")
+```
+
+# FIPS to census tract population
+```{r}
+library(tidycensus)
+#tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE)
+get_decennial("block", state = Georgia, variables = population, year = 2019)
+```
+
+Here, we receive an error message suggesting to use NHGIS data since the most recent small census geography data is not available through `get_decennial()`.
+
+# Using NHGIS data for more recent block-level population counts
+```{r}
+#Downloaded data from http://NHGIS.org
+install.packages("ipumsr")
+library(ipumsr)
+
+#read in census tract level population data
+tract <- read.csv(here("data","raw_data","nhgis0001_csv", "nhgis0001_ds244_20195_tract.csv"))
+
+#clean up dataset to remove unneeded columns/variables
+tract <- tract %>% filter(STATE == "Georgia") %>% select(1:3, 6:9, 12, 40, 43:46)
+
+#Now, we have the FIPS codes for each sample site (down to the tract/block level), and we have population counts by FIPS code.
+
+#First, we need to remove the last four digits of the FIPS code in the fips dataset so that it will correspond with the #FIPS code in the tract dataset.
+We'll also need to remove the G from the beginning of the tract FIPS codes.
+
+#combine individual codes for state, county, tract to get long form FIPS code
+a <- tract %>% mutate(FIPS = paste(STATEA, COUNTYA, sep = "")) %>%
+  mutate(FIPS = paste(FIPS, TRACTA, sep = "0"))
+
+#keep only the variables we need, FIPS code and population size (ALUBE001)
+a <- a %>% select("FIPS", "ALUBE001")
+#remove last four digits from fips codes associates with samples, so that they coordinate with population census tract FIPS codes
+b <- fips %>% mutate(FIPS = (substr(FIPS,1,nchar(FIPS)-4)))
+#combine datasets by FIPS code, so that population size is added to our main dataset based on corresponding FIPS code
+blockpop <- left_join(b, a, by = "FIPS")
+
+#We still have some missing values, but when I search for the FIPS code that are missing population levels, I am able to find them in the tract dataset.
+It looks like the FIPS code/GEOID are sometimes the same but sometimes differ by one digit (usually a zero inserted in the middle of the string somewhere).
+Let's adjust the GEOID variable to make it comparable to the FIPS codes we are still missing population data for, and then we'll join the datasets again to try to fill in missing values.
+
+#let's try GEOID minus first 7 digits
+tract$GEOID <- gsub("^.{0,7}", "", tract$GEOID) 
+#selecting only the variables we want to add in our final combination dataset
+tractpop <- tract %>% select("GEOID", "ALUBE001")
+#joining based on FIPS code to fill in missing population data
+c <- left_join(blockpop, tractpop, by = c("FIPS" = "GEOID"))
+
+#This last attempt was successful and we now have a column that gives us the census tract level population count, with no missing values.
+We no longer need the first column, or the zip code and zip code population data that we looked at initially.
+Let's clean up this dataset for further exploration
+
+#Remove first ALUBE001 column and zip, population columns
+tractpopdata <- c %>% select(-"ALUBE001.x", -"zip", -"population")
+tractpopdata <- rename(tractpopdata, "population" = "ALUBE001.y")
+
+#save as RDS
+saveRDS(tractpopdata, file = here("data", "processed_data", "tractpopdata.rds"))
+
+```

### 2021-10-26 products/manuscript/Manuscript.Rmd
updated pop data, wrf data, land cover FIPS

@@
     toc: false
-#bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found).
+bibliography: here::here("products", "library.bib") #I am having trouble with the bib file (error message: pandoc error 83 file not found).
 Removing this line for now so that Manuscript will knit.

### 2021-10-26 products/manuscript/Manuscript.Rmd
fixed manuscript citation style

--- 2021-10-08 products/manuscript/Manuscript.Rmd
@@
     toc: false
-#bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found).
+bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found).
 Removing this line for now so that Manuscript will knit.
 Will continue to troubleshoot and fix bib formatting.
-csl: ../apa.csl
+Bib file works now, but some citations recieve double brackets.
+Will continue looking into this issue
+csl: elsevier.csl
 ---
@@
 
-Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014][@Waller2017][@Peng2018].
+Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014;@Waller2017;@Peng2018].
 Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era [@Zalasiewicz2016].
-These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017][@Barboza 2018][@Prokic2019][@Evereart2018].
+These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017;@Barboza2018;@Prokic2019;@Everaert2018].
 In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans [@Barboza2018].
-Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011][@Li2018].
-Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014][@Li2018][@Browne2010][@Lebreton2017].
-However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2018][@Dris2015].
+Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011;@Li2018].
+Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014;@Li2018;@Browne2010;@Lebreton2017].
+However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2017;@Dris2015].
 Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].

### 2021-10-27 analysis.Rmd
cleaned landcover data, began modeling analysis

@@
 ```
+
+#load packages and data
+```{r}
+library(ggplot2) #for plotting
+library(broom) #for cleaning up output from lm()
+library(here) #for data loading/saving
+library(tidymodels) #for modeling
+
+data <- readRDS(here("data", "processed_data", "athensland.rds"))
+```
+
+# fit linear model to continuous outcome (particles_l)
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm)
+```
+# linear regression
+```{r}
+#linear regression model specification set up
+lm_mod <- linear_reg() %>% set_engine("lm")
+#estimating/training linear model
+lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
+#view summary of fit
+tidy(lm_fit1)
+```
+# model evaluation
+```{r}
+#split into test and train
+#set seed for reproducible analysis
+set.seed(222)
+#subset 3/4 of data as training set
+data_split <- initial_split(data, prop = 3/4)
+
+#save sets as data frames
+train_data <- training(data_split)
+test_data <- testing(data_split)
+```
+
+# recipe to fit continuous outcome to all predictors
+```{r}
+#recipe for particles per liter, all predictors
+mp_rec <- 
+  recipe(particles_l ~ ., data = train_data)
+
+#linear regression model workflow set up
+mp_wflow <- 
+  workflow() %>% add_model(lm_mod) %>% add_recipe(mp_rec)
+
+#use workflow to prepare recipe and train model with predictors
+mp_fit <- 
+  mp_wflow %>% fit(data = train_data)
+
+#extract model coefficient
+mp_fit %>% extract_fit_parsnip() %>% tidy()
+```
+# evaluate model with ROC and ROC-AUC
+```{r}
+#use trained workflow to predict with test data
+test_data <- test_data[-c(20:28),]
+predict(mp_fit, test_data)
+
+#include probabilities
+mp_aug <- augment(mp_fit, test_data)
+
+#generate ROC curve
+mp_aug %>% 
+  roc_curve(truth = particles.l, .pred) %>% autoplot()
+
+#estimate area under the curve
+mp_aug %>% roc_auc(truth = particles_l, .pred)
+```
+
+Here, I am running into an issue since july_21 data is still missing. Hopefully we'll get that from Megan soon. For now, let's try filtering it out.

### 2021-10-27 code/processing_code/censusland_code.Rmd
cleaned landcover data, began modeling analysis

@@
 landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID, fips)
+#remove empty variables and rename variables according to key described in readme file for raw NHGIS landcover data
+landcoverclean <- landcoverclean %>% select(-(4:67)) #remove variable columns from years 2001 and 2006, keeping only 2011 data 
+landcoverclean <- landcoverclean %>% rename(open_water = AREA11_2011,
+                                            perennial_ice = AREA12_2011, 
+                                            developed_open = AREA21_2011, 
+                                            developed_low = AREA22_2011, 
+                                            developed_medium = AREA23_2011, 
+                                            developed_high = AREA24_2011, 
+                                            barren_land = AREA31_2011, 
+                                            deciduous_forest = AREA41_2011, 
+                                            evergreen_forest = AREA42_2011, 
+                                            mixed_forest = AREA43_2011,
+                                            shrub = AREA52_2011, 
+                                            grassland = AREA71_2011, 
+                                            pasture = AREA81_2011, 
+                                            crops = AREA82_2011, 
+                                            woody_wetlands = AREA90_2011, 
+                                            herb_wetlands = AREA95_2011) 
+landcoverclean <- landcoverclean %>% select(-(20:35), 
+                                            -"perennial_ice") #removing empty perennial ice column, removing proporation variables (keeping area values rather than proportions.
+we have the total area and can recalculate proportion later if needed)
+
+#Based on the NAs in the perennial_ice columns and knowing the context/region of the data, I am thinking that the other NA values for certain land cover types simply indicate that there is no land cover of that type in the area - therefore, these values are essentially zero.
+Let's replace NA values with zeroes so that we are working with complete (not missing) data.
+
+landcoverclean[is.na(landcoverclean)] <- 0 #replace NA with 0
 

### 2021-10-27 analysis.Rmd
working on modeling, cleaning

@@
 ```
+
+Here, where model coefficient are extracted, the terms are individual sites rather than the different variable predictors we want to look like, unlike in last weeks exercise.
+Maybe I need to go back and remove the variables that aren't going to be helpful predictors (ex: site name)
+
 # evaluate model with ROC and ROC-AUC
@@
 
-#generate ROC curve
-mp_aug %>% 
-  roc_curve(truth = particles.l, .pred) %>% autoplot()
+#get RMSE for test data
+mp_aug %>% rmse(truth = particles_l, .pred)
 
-#estimate area under the curve
-mp_aug %>% roc_auc(truth = particles_l, .pred)
 ```
 
+Why are there so many rows resulting? When we did the analysis last week, only one row was the result, and it was just rmse and provided a value around 1.
+
 Here, I am running into an issue since july_21 data is still missing. Hopefully we'll get that from Megan soon. For now, let's try filtering it out.
+
+The error message about levels is indicating that these results appear in the test data but not the train data.
+However, this values should be continuous and not categorical.
+let's check on the class of the variable to see what's going on.
+
+```{r}
+class(test_data$e.coli.cfu)
+```
+
+Okay, e.coli.cfu has been saved as a character rather than numeric/double.
+I encountered this issue earlier in exploration but it looks like I neglected to save an updated version of the dataset.
+Let's go back to the processing code and adjust this.

### 2021-10-28 code/analysis_code/analysis.Rmd
fixing errors, building up analysis portion

---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling

data <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
```

# fit linear model to continuous outcome (particles_l)
```{r}
#attempt to visualize data
#look at particles per L vs population
ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm)
```

# linear regression
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
#view summary of fit
tidy(lm_fit1)
```

# linear regression for one predictor, one outcome
```{r}
#predict particles per liter based on population
lm_fit2 <- lm_mod %>% fit(particles_l ~ population, data = data)
#view summary of fit
tidy(lm_fit2)

#predict particles per liter based on e coli
lm_fit3 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = data)
#view summary of fit
tidy(lm_fit3)

#comparison of fit
comp2_3 <- anova(lm_fit2$fit, lm_fit3$fit, test = "Chisq")
comp2_3
```

Here, we get an error message that models were not fitted to the same size of dataset. This is likely due to missing values in some predictors. To solve this, we can clean up the dataset some more and create a dataset specifically to use for modelling, that has only predictors of interest and no missing values.

#clean up data for modelling
```{r}
data %>% is.na() %>% summary()
complete <- data %>% na.omit()
```

Trying code from above again with the complete dataset

```{r}
#estimating/training linear model
lm_fit4 <- lm_mod %>% fit(particles_l ~ ., data = complete)
#view summary of fit
tidy(lm_fit4)
#predict particles per liter based on population
lm_fit5 <- lm_mod %>% fit(particles_l ~ population, data = complete)
#view summary of fit
tidy(lm_fit5)

#predict particles per liter based on e coli
lm_fit6 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = complete)
#view summary of fit
tidy(lm_fit6)

#comparison of fit
comp5_6 <- anova(lm_fit5$fit, lm_fit6$fit, test = "Chisq")
comp5_6
```

Here, we have a VERY large RSS, which I think indicates a poor model fit for our data. Let's try with some variables removed.

# Subsetting data to remove unimportant variables
```{r}
subset <- complete %>% select(-"site", 
                              "lat", 
                              "long",
                              "date") 
#estimating/training linear model
lm_fit7 <- lm_mod %>% fit(particles_l ~ ., data = subset)
#view summary of fit
tidy(lm_fit7)

#predict particles per liter based on population
lm_fit8 <- lm_mod %>% fit(particles_l ~ population, data = subset)
#view summary of fit
tidy(lm_fit8)

#predict particles per liter based on e coli
lm_fit9 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = subset)
#view summary of fit
tidy(lm_fit9)

#model comparison
comp8_9 <- anova(lm_fit8$fit, lm_fit9$fit, test = "Chisq")
comp8_9
```

Here, we see the same RSS results even after subsetting the data (the values are exactly the same when run on the "complete" data as on the "subset" data.)

# model evaluation, splitting data into train and test
```{r}
#split into test and train
#set seed for reproducible analysis
set.seed(222)
#subset 3/4 of data as training set
data_split <- initial_split(data, prop = 3/4)

#save sets as data frames
train_data <- training(data_split)
test_data <- testing(data_split)
```

# recipe to fit continuous outcome to all predictors
```{r}
#recipe for particles per liter, all predictors
mp_rec <- 
  recipe(particles_l ~ ., data = train_data)

#linear regression model workflow set up
mp_wflow <- 
  workflow() %>% add_model(lm_mod) %>% add_recipe(mp_rec)

#use workflow to prepare recipe and train model with predictors
mp_fit <- 
  mp_wflow %>% fit(data = train_data)

#extract model coefficient
mp_fit %>% extract_fit_parsnip() %>% tidy()
```

Here, we see site names come up first, even though we know site names are not going to be good predictors of microplastic levels. I need to go back and remove the variables that aren't going to be helpful predictors (ex: site name)

# evaluate model with ROC and ROC-AUC
```{r}
#use trained workflow to predict with test data
predict(mp_fit, test_data)

#include probabilities
mp_aug <- augment(mp_fit, test_data)

#get RMSE for test data
mp_aug %>% rmse(truth = particles_l, .pred)

```

Why are there so many rows resulting? When we did the analysis last week, only one row was the result, and it was just rmse and provided a value around 1.

#trying model evaluation with SUBSET
```{r}

```

### 2021-10-28 code/analysis_code/exploration.Rmd
fixing errors, building up analysis portion

@@
 
-```{r}
-mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
-```
-
-Something weird is giong on with the values for e.coli.cfu.
-
-```{r}
-class(mpdata$e.coli.cfu)
-```
-
-It looks like that ended up saved as a character instead of integer. Let's convert. 
-# covert character to numeric
-```{r}
-mpdata <- mpdata %>% mutate(e.coli.cfu = as.numeric(e.coli.cfu))
-class(mpdata$e.coli.cfu)
-```
-
-# try plot again
+# plot e.coli.cfu vs particles_l
 ```{r}
@@
 #subset data into just distance from each WRF
-wrf <- mpdata %>% select("ccwrf", "nowrf", "miwrf") 
+wrf <- mpdata %>% select("ccwrf", "nowrf", "miwrf", "site", "date") 
 #find minimum distance to any WRF out of all three
@@
 
-#minwrf contains the variable min, which tells us the distance from the sample site to the nearest WRF. This is all of the information that we need, so no we will re-join to the original DF
+#minwrf contains the variable min, which tells us the distance from the sample site to the nearest WRF. This is all of the information that we need, so now we will re-join to the original DF
 mpwrfdata <- left_join(mpdata, minwrf)

### 2021-10-28 code/processing_code/censusprocessing.Rmd
fixing errors, building up analysis portion

@@
 
-# FIPS to census tract population
-```{r}
-library(tidycensus)
-#tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE)
-get_decennial("block", state = Georgia, variables = population, year = 2019)
-```
-
-Here, we receive an error message suggesting to use NHGIS data since the most recent small census geography data is not available through `get_decennial()`.
+Here, after attemping to use `get_decennial()` we receive an error message suggesting to use NHGIS data since the most recent small census geography data is not available through `get_decennial()`.
 

### 2021-10-28 code/analysis_code/analysisscript.R

@@ -1,24 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-#this is where my analysis script for the MADA project will be located. Keeping
-#this file as a placeholder for now. 
-
-
-
-  

### 2021-10-28 analysis.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files

--- 2021-10-26 analysis.Rmd
@@
 ```
+
+#load packages and data
+```{r}
+library(ggplot2) #for plotting
+library(broom) #for cleaning up output from lm()
+library(here) #for data loading/saving
+library(tidymodels) #for modeling
+
+data <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
+```
+
+# fit linear model to continuous outcome (particles_l)
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Particles/Liter")
+```
+
+# particles per liter vs population with outliers filtered out
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+data %>% filter(particles_l < 750) %>% ggplot(aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs.
+Population", x = "Population", y = "Particles/Liter")
+```
+
+# fit linear model to continuous outcome (particles_l) with log transformation
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+ggplot(data, aes(x = population, y = particles_l)) + geom_point() +
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Log(Particles/Liter)")
+```
+
+# visualize particles vs e.coli.cfu 
+```{r}
+ggplot(data, aes(x = e.coli.cfu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. E. coli CFU", x = "CFU/mL", y = "Particles/Liter")
+```
+
+# particles per liter vs e.coli.cfu with log transformation
+```{r}
+ggplot(data, aes(x = e.coli.cfu, y = particles_l)) + geom_point() + scale_x_continuous(trans='log10') +
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. E. coli CFU", x = "Log(CFU/mL)", y = "Log(Particles/Liter)")
+```
+
+# linear regression
+```{r}
+#linear regression model specification set up
+lm_mod <- linear_reg() %>% set_engine("lm")
+#estimating/training linear model
+lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
+#view summary of fit
+tidy(lm_fit1)
+```
+
+# linear regression for one predictor, one outcome
+```{r}
+#predict particles per liter based on population
+lm_fit2 <- lm_mod %>% fit(particles_l ~ population, data = data)
+#view summary of fit
+tidy(lm_fit2)
+
+#predict particles per liter based on e coli
+lm_fit3 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = data)
+#view summary of fit
+tidy(lm_fit3)
+
+#comparison of fit
+comp2_3 <- anova(lm_fit2$fit, lm_fit3$fit, test = "Chisq")
+comp2_3
+```
+
+Here, we get an error message that models were not fitted to the same size of dataset.
+This is likely due to missing values in some predictors.
+To solve this, we can clean up the dataset some more and create a dataset specifically to use for modelling, that has only predictors of interest and no missing values.
+
+#clean up data for modelling
+```{r}
+data %>% is.na() %>% summary()
+complete <- data %>% na.omit()
+```
+
+Trying code from above again with the complete dataset
+
+```{r}
+#estimating/training linear model
+lm_fit4 <- lm_mod %>% fit(particles_l ~ ., data = complete)
+#view summary of fit
+tidy(lm_fit4)
+#predict particles per liter based on population
+lm_fit5 <- lm_mod %>% fit(particles_l ~ population, data = complete)
+#view summary of fit
+tidy(lm_fit5)
+
+#predict particles per liter based on e coli
+lm_fit6 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = complete)
+#view summary of fit
+tidy(lm_fit6)
+
+#comparison of fit
+comp5_6 <- anova(lm_fit5$fit, lm_fit6$fit, test = "Chisq")
+comp5_6
+```
+
+Here, we have a VERY large RSS, which I think indicates a poor model fit for our data. Let's try with some variables removed.
+
+# Subsetting data to remove unimportant variables
+```{r}
+subset <- complete %>% select(-"site", 
+                              "lat", 
+                              "long",
+                              "date") 
+#estimating/training linear model
+lm_fit7 <- lm_mod %>% fit(particles_l ~ ., data = subset)
+#view summary of fit
+tidy(lm_fit7)
+
+#predict particles per liter based on population
+lm_fit8 <- lm_mod %>% fit(particles_l ~ population, data = subset)
+#view summary of fit
+tidy(lm_fit8)
+
+#predict particles per liter based on e coli
+lm_fit9 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = subset)
+#view summary of fit
+tidy(lm_fit9)
+
+#model comparison
+comp8_9 <- anova(lm_fit8$fit, lm_fit9$fit, test = "Chisq")
+comp8_9
+```
+
+Here, we see the same RSS results even after subsetting the data (the values are exactly the same when run on the "complete" data as on the "subset" data.)
+
+# model evaluation, splitting data into train and test
+```{r}
+#split into test and train
+#set seed for reproducible analysis
+set.seed(222)
+#subset 3/4 of data as training set
+data_split <- initial_split(data, prop = 3/4)
+
+#save sets as data frames
+train_data <- training(data_split)
+test_data <- testing(data_split)
+```
+
+# recipe to fit continuous outcome to all predictors
+```{r}
+#recipe for particles per liter, all predictors
+mp_rec <- 
+  recipe(particles_l ~ ., data = train_data)
+
+#linear regression model workflow set up
+mp_wflow <- 
+  workflow() %>% add_model(lm_mod) %>% add_recipe(mp_rec)
+
+#use workflow to prepare recipe and train model with predictors
+mp_fit <- 
+  mp_wflow %>% fit(data = train_data)
+
+#extract model coefficient
+mp_fit %>% extract_fit_parsnip() %>% tidy()
+```
+
+Here, we see site names come up first, even though we know site names are not going to be good predictors of microplastic levels.
+I need to go back and remove the variables that aren't going to be helpful predictors (ex: site name)
+
+# evaluate model with RMSE
+```{r}
+#use trained workflow to predict with test data
+predict(mp_fit, test_data)
+
+#include probabilities
+mp_aug <- augment(mp_fit, test_data)
+
+#get RMSE for test data
+mp_aug %>% rmse(truth = particles_l, .pred)
+
+```
+
+Why are there so many rows resulting? When we did the analysis last week, only one row was the result, and it was just rmse and provided a value around 1.
+
+#trying model evaluation with SUBSET
+```{r}
+
+```

### 2021-10-28 code/analysis_code/exploration.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files

--- 2021-10-08 code/analysis_code/exploration.Rmd
@@
 library(ggplot2)
+library(geodist) #to calculate distance from sample site to each water reclamation facility
 
 #path to data
-data_location <- here::here("data","processed_data","popdata.rds")
+data_location <- here::here("data","processed_data","mpwrfcombo.rds")
 
@@
 
-# take a look at data
+# MP concentration by season
 ```{r}
@@
 
-p1 <- mpdata %>% ggplot(aes(x = date, y = particles_l)) + geom_boxplot(aes()) + labs(title = "Microplastic Concentration")
+mpdata$date <- factor(mpdata$date , levels=c("nov_20", "feb_21", "apr_21", "july_21"))
+
+p1 <- mpdata %>% ggplot(aes(x = date, y = particles_l)) + 
+  geom_boxplot(aes()) + 
+  labs(title = "Microplastic Concentration by Sample Date", x = "Date", y = "Particles/Liter") + 
+  scale_x_discrete(labels=c("Fall 2020",
+                            "Winter 2021",
+                            "Spring 2021",
+                            "Summer 2021"))
 
@@
 #save figure
-figure_file = here("results","concentrationbydate.png")
-ggsave(filename = figure_file, plot=p1)
+ggsave(filename = here("results","concentrationbydate.png"), plot=p1)
 ```
@@
 ```{r}
-p2 <- mpdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point()
+p2 <- mpdata %>% ggplot(aes(x = population, y = particles_l)) + geom_point() +
+  labs(title = "Microplastic Concentration vs Population", x = "Population", y = "Particles/Liter")
 
@@
 #save figure
-figure_file = here("results","particlesvpop.png")
-ggsave(filename = figure_file, plot=p1)
+ggsave(filename = here("results","particlesvpop.png"), plot=p2)
 ```
@@
 
+# particles vs population with log transformation
 ```{r}
-mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
+p3 <- mpdata %>% ggplot(aes(x = population, y = particles_l)) + 
+  geom_point() + 
+  scale_y_continuous(trans='log10') +
+  labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Log(Particles/Liter)")
+
+plot(p3)
+#save figure
+figure_file = here("results","logparticlesvpop.png")
+ggsave(filename = figure_file, plot=p3)
 ```
 
-Something weird is giong on with the values for e.coli.cfu.
+# plot e.coli.cfu vs particles_l
+```{r}
+p4 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + 
+  geom_point() + 
+  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
+  labs(title = "Microplastic Concentration vs. E. coli CFU", x = "CFU", y = "Particles/Liter")
 
-```{r}
-class(mpdata$e.coli.cfu)
-```
-
-It looks like that ended up saved as a character instead of integer. Let's convert. 
-# covert character to numeric
-```{r}
-mpdata <- mpdata %>% mutate(e.coli.cfu = as.numeric(e.coli.cfu))
-class(mpdata$e.coli.cfu)
-```
-
-# try plot again
-```{r}
-p3 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1))
-
-plot(p3)
+plot(p4)
 
 #save figure
-figure_file = here("results","mpvbacteria.png")
-ggsave(filename = figure_file, plot=p3)
+figure_file = here("results","mpvcfu.png")
+ggsave(filename = figure_file, plot=p4)
 ```
@@
 ```{r}
-p4 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust=1)) + scale_x_continuous(trans='log10') +
-  scale_y_continuous(trans='log10')
+p5 <- mpdata %>% ggplot(aes(x = e.coli.cfu, y = particles_l)) + 
+  geom_point() + 
+  theme(axis.text.x = element_text(angle = 90, hjust=1)) + 
+  scale_x_continuous(trans='log10') +
+  scale_y_continuous(trans='log10') + 
+  labs(title = "Microplastic Concentration vs. E. coli CFU", x = "Log(CFU)", y = "Log(Particles/Liter)")
 
-plot(p4)
+plot(p5)
 
 #save figure
-figure_file = here("results","logmpvbacteria.png")
-ggsave(filename = figure_file, plot=p4)
+figure_file = here("results","logmpvcfu.png")
+ggsave(filename = figure_file, plot=p5)
 ```
 
-Note: missing values are for e.coli.cfu for July 2021 sampling. These values should be available for comparison soon.
+Let's look at overall distribution of MP levels.
 
-Let's look at overall distribution of MP levels
+# Distribution of MP concentration
 ```{r}
-mpdata %>% ggplot(aes(x = particles_l)) + geom_histogram()
+mpdata %>% ggplot(aes(x = particles_l)) + geom_histogram() + 
+  labs(title = "Distribution of Microplastic Concentration", x = "Particles/Liter", y = "Observations")
 ```
@@
 
+# Distribution of MP concentration without outliers
 ```{r}
-p5 <- mpdata %>% filter(particles_l < 400) %>% ggplot(aes(x = particles_l)) + geom_histogram()
+p6 <- mpdata %>% filter(particles_l < 400) %>% ggplot(aes(x = particles_l)) + geom_histogram() + 
+  labs(title = "Distribution of Microplastic Concentration", x = "Particles/Liter", y = "Observations")
 
-plot(p5)
+plot(p6)
 
@@
 figure_file = here("results","distribution.png")
-ggsave(filename = figure_file, plot=p5)
+ggsave(filename = figure_file, plot=p6)
 ```
@@
 We see that most levels fall below 100 particles per liter. 
-
+# Summary statistics
 ```{r}
 summary(mpdata$particles_l)
-
 ```
@@
 
-I also wanted to look at proximity of sample sites to water reclamation facilities (WRFs). Let's try that. 
-# load WRF location
+# correlation matrices
 ```{r}
-#path to data
-wrfdata_location <- here::here("data","processed_data","wrflocation.rds")
+library(GGally) #for produces correlation matrices
 
-#load data
-wrflocation <- readRDS(wrfdata_location)
+#subsetting data to keep only variables we are interested in for the corr plot
+mpdatacorr <- mpdata %>% select(-"quarter", 
+                                -"month", 
+                                -"lat",
+                                -"long", 
+                                -"year")
+
+#correlation using MP, UOWN, population data
+p7 <- ggcorr(mpdatacorr, 
+       label_alpha = TRUE) + labs(title = "Microplastic Predictor Correlation Matrix")
+
+plot(p7)
+#save figure
+figure_file = here("results","matrix.png")
+ggsave(filename = figure_file, plot=p7)
 ```
 
-Eventually, I would like to be able to view this information on a map - where WRFs are labelled, and samples sites are color-coded by density of microplastics found.
-For now, let's see if we can calculate distance otherwise.
+# exploring land use with a correlation matrix
+```{r}
+athensland <- readRDS(here("data", "processed_data", "athensland.rds"))
 
-```{r}
-library(geodist)
+#subset data to view in plot
+athenslandcorr <- athensland %>% select(-"long", 
+                                        -"lat")
 
-#calculate distance between sample coordinates and WRF location
-wrfdistance <- geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat)
+#land use correlation matrix
+p8 <- ggcorr(athenslandcorr, 
+       label_alpha = TRUE) + labs(title = "Land Use Correlation Matrix")
+
+plot(p8)
+#save figure
+figure_file = here("results","landmatrix.png")
+ggsave(filename = figure_file, plot=p8)
 ```
-
-I think the product we have here is a matrix of the distance of each sample site from each of the three WRFs, in the order that they appear in the WRF data (Cedar Creek, North Oconee, Middle Oconee).
-But, the labelling of the data is not super clear, so I need to verify this to be sure.
-
-```{r}
-test <- mpdata %>% mutate(wrfdist = geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat))
-```
-
-This produced another column like I wanted, but I don't know which WRF the distance was calculated for.
-Maybe I need to save WRF locations as individual lat and long values, and produce three new columns that way?
-However, I am not sure what the final product would look like if we do that, because a sample site could be closely associated with one WRF but much further from another one.
-Maybe I need to narrow this down to proximity to the nearest WRF, so that each sample site is only associated with the distance to ONE WRF, not three.

### 2021-10-28 code/processing_code/wrfprocessing.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files

---
title: "wrfprocessing"
output: html_document
---

I also wanted to look at proximity of sample sites to water reclamation facilities (WRFs). Let's try that. 
# load WRF location
```{r}
#load MP data
mpdata <- readRDS(here("data", "processed_data", "popdata.rds"))
#path to data
wrfdata_location <- here::here("data","processed_data","wrflocation.rds")

#load data
wrflocation <- readRDS(wrfdata_location)
```

Eventually, I would like to be able to view this information on a map - where WRFs are labelled, and samples sites are color-coded by density of microplastics found. For now, let's see if we can calculate distance otherwise.

```{r}
#calculate distance between sample coordinates and WRF location
wrfdistance <- geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat)
```

I think the product we have here is a matrix of the distance of each sample site from each of the three WRFs, in the order that they appear in the WRF data (Cedar Creek, North Oconee, Middle Oconee). But, the labelling of the data is not super clear, so I need to verify this to be sure.

```{r}
test <- mpdata %>% mutate(wrfdist = geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat))
```

This produced another column like I wanted, but I don't know which WRF the distance was calculated for. Maybe I need to save WRF locations as individual lat and long values, and produce three new columns that way? However, I am not sure what the final product would look like if we do that, because a sample site could be closely associated with one WRF but much further from another one. Maybe I need to narrow this down to proximity to the nearest WRF, so that each sample site is only associated with the distance to ONE WRF, not three.

#Working on gathering WRF distance
```{r}
#set individual WRF locations and save as values
cedarcreeklat = 33.87351 
cedarcreeklong = -83.33639
northoconeelat = 33.93402
northoconeelong = -83.36018
middleoconeelat = 33.90862
middleoconeelong = -83.39214

mpdata <- mpdata %>% mutate(ccwrf = geodist_vec(mpdata$long, mpdata$lat, cedarcreeklong, cedarcreeklat))
mpdata <- mpdata %>% mutate(nowrf = geodist_vec(mpdata$long, mpdata$lat, northoconeelong, northoconeelat))
mpdata <- mpdata %>% mutate(miwrf = geodist_vec(mpdata$long, mpdata$lat, middleoconeelong, middleoconeelat))

#Since I only care about the proximity to the closest WRF, my next task is to write code that creates a new variable returning only the lowest value out of ccwrf, nowrf, and miwrf for each sample site.

#subset data into just distance from each WRF
wrf <- mpdata %>% select("ccwrf", "nowrf", "miwrf", "site", "date") 
#find minimum distance to any WRF out of all three
val <- do.call(pmin, c(wrf, na.rm = TRUE)) 
minwrf <- transform(wrf, min=val, minCol = names(wrf)[max.col(wrf == val, 'first')])

#minwrf contains the variable min, which tells us the distance from the sample site to the nearest WRF. This is all of the information that we need, so now we will re-join to the original DF
mpwrfdata <- left_join(mpdata, minwrf)

#remove variables we won't need anymore
mpwrfdata <- mpwrfdata %>% select(-"ccwrf", 
                                  -"nowrf",
                                  -"miwrf")
#rename columns
mpwrfdata <- mpwrfdata %>% rename("dist" = "min", 
                     "wrf" = "minCol")

#save as RDS
saveRDS(mpwrfdata, file = here("data", "processed_data", "mpwrfdata.rds"))
```

Update 10/25/21
I've updated the population data by accessing census tract level population rather than zip code level population. This should give us more localized population levels and might paint a better picture of the relationship between MP particles per liter and local population level.

#Adding WRF data to updated population data
```{r}
#load data
tractpopdata <- readRDS(here("data", "processed_data", "tractpopdata.rds"))

#add wrf data to tractpopdata
wrfjoin <- mpwrfdata %>% select("site", "date", "wrf", "dist") #selecting just the variables we want to add
mpwrfcombo <- left_join(tractpopdata, wrfjoin, by = c("site", "date")) #joining the two datasets

glimpse(mpwrfcombo)
#dist variable is saved as character when it needs to be numeric
#convert dist from character to numeric
mpwrfcombo <- mpwrfcombo %>% mutate(dist = as.numeric(dist))

#save as RDS
saveRDS(mpwrfcombo, file = here("data", "processed_data", "mpwrfcombo.rds"))
```

Now, `mpwrfcombo` contains the sample info and the new WRF distance info. However, something to work on going forward is that a sample site may be near a WRf but not downstream of it. It might be more important to know the distance to the nearest upstream WRF. I'll have to go back and consider this as I continue my exploration/analysis.

Other than not including landcover data, `mpwrfcombo` is the most comprehensive dataset, including almost all of the data that I have compiled so far.

### 2021-10-28 data/processed_data/data_index.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files

---
title: "readme"
output: html_document
---

# Index of Processed Data, Alphabetical
1. *athensland.rds*
  - contains all sample data combined with tract population level, tract land use, UOWN data
2. *cleansamplesdata.rds*
  - contains JUST original microplastics data, no additional information (109 observations, 6 variables)
3. *cleanuowndata.rds*
  - contains UOWN data corresponding to MP sampling, but not yet combined with MP data
4. *combodata.rds*
  - combination of MP and UOWN data (contains particles_l and e.coli, visual score, turbidity)
5. *controldata.rds*
 - subset from original MP sample data - contains only control observations, no sample observations (technical controls from laboratory methods)
6. *fips.rds*
 - contains MP, UOWN, zipcode population, adds FIPS codes to sample observations (does not include population level)
7. *mpwrfcombo.rds*
 - contains MP, UOWN, population, and WRF distance variables
8. *mpwrfdata.rds*
 - intermediate dataset containing MP, UOWN, population, and WRF distances variables prior to updated population info. we may be able to remove this later
9. *popdata.rds*
 - contains MP and UOWN data, and uses zipcode to get population levels. After creating this dataset, I realized census tract population count would be more appropriate. We may be able to remove this later as it is not used in later exploration/analysis
10. *tractpopdata.rds*
 - contains MP, UOWN, FIPS, and tract population level data
10. *volunteer.rds*
 - subset of MP data containing just July_21 observations and volunteer ID numbers. May be used to compare volunteer MP counting accuracy
11. *wrflocation.rds*
 - contains coordinates of Athens' 3 WRF. Initially attempted to import data and use to calculate distance to nearest WRF, but later ended up manually assigning values in the Rmd. We may be able to remove this later

### 2021-10-28 products/manuscript/Manuscript.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files

@@
 ---
-title: "Microplastics in the Urban Environment"
-subtitle: "Exploratory Study of Freshwater Microplastics in Athens, GA"
+title: "Plastic in the Urban Freshwater Environment"
+subtitle: "An Exploratory Study of Microplastics in the Upper Oconee Watershed via Community-Based Efforts"
 author: Amelia M. Foley
@@
     number_sections: true
+    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
   bookdown::html_document2: 
     toc: false
-bibliography: library.bib #I am having trouble with the bib file (error message: pandoc error 83 file not found).
-Removing this line for now so that Manuscript will knit.
-Will continue to troubleshoot and fix bib formatting.
-Bib file works now, but some citations recieve double brackets.
-Will continue looking into this issue
+bibliography: library.bib 
 csl: elsevier.csl
@@
 
-This document will be my final manuscript for the MADA project. I am keeping these draft sections for structure now, until I edit them later.
+# Abstract
+With growing awareness of plastic pollution and its detrimental environmental impacts, there has been increased attention on the study of microplastics.
+These small debris result from the weathering and breakdown of plastic waste from urban runoff, agricultural runoff, ocean dumping, stormwater, waste management, and other sources.
+Microplastics pose a threat to both aquatic life and human health due to their chemical makeup, their build-up in the environment, and their suspected long-term biological effects.
+As their ubiquity becomes more apparent, so does the need for accessible, affordable, and effective methods for the widespread study and quantification of microplastic pollution.
 
-# Summary/Abstract
-_Write a summary of your project._
+This study takes place in Athens, GA, an urban area which houses a population of about 127,000.
+The project relies on local volunteer citizen science efforts for water sampling, processing, and analysis.
+Given the local and regional ecological significance of the Oconee River—including its transport of freshwater to the Atlantic Ocean—this study aims to collect relevant data regarding microplastics in the Upper Oconee Watershed, which includes 2 rivers and 15 creeks in Athens-Clarke County.
+
+Through community-based sample collection and analysis, volunteers quantified microplastics through membrane filtration and visual identification.
+Preliminary data consistently show that local freshwater microplastic levels are greater than the U.S.
+average in the Upper Oconee Watershed, a portion of the Oconee River basin with both a high population size and high population density relative to other regions of Georgia.
+Average and median concentrations of 95 particles/L and 46 particles/L, respectively, were recorded throughout one year of seasonal sampling.
+Further visual and statistical analysis will confirm the project’s results, and discussion of these results in context will reveal the efficacy of this community-based approach to the quantification of microplastics.
 
@@
 
-# Methods and Results
-
-_In most research papers, results and methods are separate. You can combine them here if you find it easier. You are also welcome to structure things such that those are separate sections._
+# Methods
 
@@
 
+# Results
+
 # Discussion

### 2021-10-28 code/analysis_code/exploration.Rmd
adding more plots to exploration.Rmd

@@
 library(geodist) #to calculate distance from sample site to each water reclamation facility
+library(GGally) #for producing correlation matrices
 
@@
 ```{r}
-library(GGally) #for produces correlation matrices
-
 #subsetting data to keep only variables we are interested in for the corr plot
@@
 ```
+
+# concentration at each site throughout seasons
+```{r}
+p9 <- mpdata %>% ggplot(aes(x = site, y = particles_l, color = date)) + 
+  geom_point() + scale_color_discrete(name = "Date", 
+                                     breaks= c("nov_20", "feb_21", "apr_21", "july_21"), 
+                                     labels = c("Fall 2020", "Winter 2021", "Spring 2021", "Summer 2021")) +
+  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
+  labs(title = "Seasonal Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) #removing outliers to better visualize majority of concentrations
+
+plot(p9)
+#save figure
+figure_file = here("results","seasonalconc_bysite.png")
+ggsave(filename = figure_file, plot=p9)
+```
+
+# View concentration by watershed
+```{r}
+p10 <- mpdata %>% ggplot(aes(x = watershed, y = particles_l, color = )) + 
+  geom_boxplot()  +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(title = "Microplastic Concentration by Watershed", x = "Watershed", y = "Particles/Liter") + ylim(0,500)
+
+plot(p10)
+#save figure
+figure_file = here("results","conc_bywatershed.png")
+ggsave(filename = figure_file, plot=p10)
+```
+# Boxplot for individual sites throughout season
+```{r}
+p11 <- mpdata %>% ggplot(aes(x = site, y = particles_l, color = )) + 
+  geom_boxplot()  +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500)
+
+plot(p11)
+#save figure
+figure_file = here("results","conc_bysite.png")
+ggsave(filename = figure_file, plot=p11)
+```
+# concentration over time by watershed
+```{r}
+# calculate mean concentration in each watershed for each sample date
+mean_data <- group_by(mpdata, watershed, date) %>%
+             summarise(particles_l = mean(particles_l, na.rm = TRUE))
+
+# create plot using mean data
+p12 <- mean_data %>% ggplot(aes(x = date, y = particles_l, color = watershed, group = watershed)) + 
+  geom_point() + geom_path() + scale_color_discrete(name = "Watershed")+ 
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(title = "Watershed Microplastic Concentration Over Time", x = "Watershed", y = "Particles/Liter") + scale_x_discrete(labels=c("Fall 2020",
+                            "Winter 2021",
+                            "Spring 2021",
+                            "Summer 2021"))+ 
+  ylim(0,300) #removing outliers to better visualize majority of concentrations
+
+plot(p12)
+#save figure
+figure_file = here("results","watershed_overtime.png")
+ggsave(filename = figure_file, plot=p12)
+```
+
+Nice!
+Going forward, I should think about whether to keep or remove the watersheds that were only sampled in the last (July) sampling.
+I also see that some watershed are missing intermediate data points and are not connected.
+Would we want to connect and skip over a season, or leave them unconnected (looking at Oconee River)

### 2021-10-29 code/analysis_code/analysis.Rmd
finishing project pt 3

@@
 library(tidymodels) #for modeling
+library(gridExtra) #for exporting tables
 
@@
 #look at particles per L vs population
-ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm)
+m1 <- ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Particles/Liter")
+
+plot(m1)
+#save figure
+ggsave(filename = here("results","m_concvpop.png"), plot=m1)
+```
+
+# particles per liter vs population with outliers filtered out
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+m2 <- data %>% filter(particles_l < 750) %>% ggplot(aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs.
+Population", x = "Population", y = "Particles/Liter")
+
+plot(m2)
+#save figure
+ggsave(filename = here("results","m_concvpop_no_outlier.png"), plot=m2)
+```
+
+# fit linear model to continuous outcome (particles_l) with log transformation
+```{r}
+#attempt to visualize data
+#look at particles per L vs population
+m3 <- ggplot(data, aes(x = population, y = particles_l)) + geom_point() +
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Log(Particles/Liter)")
+
+plot(m3)
+#save figure
+ggsave(filename = here("results","m_logconcvpop.png"), plot=m3)
+```
+
+# visualize particles vs e.coli.cfu 
+```{r}
+m4 <- ggplot(data, aes(x = e.coli.cfu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. E. coli CFU", x = "CFU/mL", y = "Particles/Liter")
+
+plot(m4)
+#save figure
+ggsave(filename = here("results","m_concvcfu.png"), plot=m4)
+```
+
+# particles per liter vs e.coli.cfu with log transformation
+```{r}
+m5 <- ggplot(data, aes(x = e.coli.cfu, y = particles_l)) + geom_point() + scale_x_continuous(trans='log10') +
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. E. coli CFU", x = "Log(CFU/mL)", y = "Log(Particles/Liter)")
+
+plot(m5)
+#save figure
+ggsave(filename = here("results","m_logconcvcfu.png"), plot=m5)
 ```
@@
 ```{r}
+data <- data %>% select("particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist")
+
 #linear regression model specification set up
@@
 lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
+#save table
+lm_fit1_table <- tidy(lm_fit1)
 #view summary of fit
-tidy(lm_fit1)
+(lm_fit1)
+#save file
+saveRDS(lm_fit1_table, file = here("results", "lm_fit_table.rds"))
 ```
@@
 #comparison of fit
-comp2_3 <- anova(lm_fit2$fit, lm_fit3$fit, test = "Chisq")
-comp2_3
+#comp2_3 <- anova(lm_fit2$fit, lm_fit3$fit, test = "Chisq")
+#comp2_3
+#these lines of code are kept to show how I got the original error message about models not being fitted to the same size dataset, but I have cut them out so that the code will run smoothly
 ```
@@
 
-#clean up data for modelling
+# create subset of data with only complete cases
 ```{r}
-data %>% is.na() %>% summary()
-complete <- data %>% na.omit()
+data %>% is.na() %>% summary() #check out missing cases
+complete <- data %>% na.omit() #save complete cases
 ```
@@
 Trying code from above again with the complete dataset
-
+# Modeling with complete cases only
 ```{r}
@@
 
-Here, we have a VERY large RSS, which I think indicates a poor model fit for our data. Let's try with some variables removed.
+Here, we have a VERY large RSS, which I think indicates a poor model fit for our data.
+Let's try with some variables removed.
+If an RSS of 0 is a perfect model, our RSS of 2 million is either a *very bad* model or I am making some sort of error/misinterpreting(?).
 
-# Subsetting data to remove unimportant variables
+# Modeling single predictors with complete cases
 ```{r}
-subset <- complete %>% select(-"site", 
-                              "lat", 
-                              "long",
-                              "date") 
-#estimating/training linear model
-lm_fit7 <- lm_mod %>% fit(particles_l ~ ., data = subset)
-#view summary of fit
-tidy(lm_fit7)
-
 #predict particles per liter based on population
-lm_fit8 <- lm_mod %>% fit(particles_l ~ population, data = subset)
+lm_fit8 <- lm_mod %>% fit(particles_l ~ population, data = complete)
 #view summary of fit
@@
 #predict particles per liter based on e coli
-lm_fit9 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = subset)
+lm_fit9 <- lm_mod %>% fit(particles_l ~ e.coli.cfu, data = complete)
 #view summary of fit
@@
 
-# recipe to fit continuous outcome to all predictors
+# recipe to fit particles/L outcome to all predictors
 ```{r}
@@
 
-# evaluate model with ROC and ROC-AUC
+# evaluate model with RMSE
 ```{r}
@@
 
-#trying model evaluation with SUBSET
+#trying model evaluation with turbidity - linear regression
 ```{r}
+#estimating/training linear model
+turb_fit1 <- lm_mod %>% fit(particles_l ~ turbidity.ntu, data = data)
+#view summary of fit
+tidy(turb_fit1)
+
+#visualize linear model
+t1 <- ggplot(data, aes(x = turbidity.ntu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs.
+Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")
+
+#save figure
+ggsave(filename = here("results","m_concvturbidity.png"), plot=t1)
+```
+
+#evaluating fit using turbidity as predictor
+```{r}
+#recipe for particles per liter, all predictors
+turb_rec <- 
+  recipe(particles_l ~ turbidity.ntu, data = train_data)
+
+#linear regression model workflow set up
+turb_wflow <- 
+  workflow() %>% add_model(lm_mod) %>% add_recipe(turb_rec)
+
+#use workflow to prepare recipe and train model with predictors
+turb_fit <- 
+  turb_wflow %>% fit(data = train_data)
+
+#extract model coefficient
+turb_fit %>% extract_fit_parsnip() %>% tidy()
+
+#use trained workflow to predict with test data
+predict(turb_fit, test_data)
+
+#include probabilities
+turb_aug <- augment(turb_fit, test_data)
+
+#get RMSE for test data
+turb_aug %>% rmse(truth = particles_l, .pred)
 
 ```
+
+Here, I am still confused about why so many rmse values are resulting.
+Earlier, I was getting multiple rmse values but thought it was because I was building a model with all predictors.
+Here, I am using one predictor but still getting multiple rmse values.
+If I remember correctly, there was only one rmse value in the exercises we worked on in the previous weeks.

### 2021-10-29 code/analysis_code/exploration.Rmd
finishing project pt 3

@@
 library(ggplot2)
-library(geodist) #to calculate distance from sample site to each water reclamation facility
 library(GGally) #for producing correlation matrices
@@
 
+```
+
+# Summary statistics
+```{r}
+summary <- summary(mpdata)
+summary
+#save table
+saveRDS(summary, file = here("results", "summarytable.rds"))
 ```

### 2021-10-29 code/processing_code/censusland_code.Rmd
fixing errors, building up analysis portion
finishing project pt 3

@@
 ```{r}
+library(dplyr) #for data processing
+library(here) #to set paths
+library(tidyverse)
 #install.packages("tidycensus", repos = "http://cran.us.r-project.org")
@@
 
-#path to data
-data_location <- here::here("data","processed_data","cleansampledata.rds")
-
-#load data. 
-data <- readRDS(data_location)
 ```
@@
 
+#cleanup resultant dataset
+athensland <- athensland %>% select(-"fips", 
+                              -"FIPS", 
+                              -"AREA", 
+                              -"quarter", 
+                              -"month",
+                              -"year")
+
 #save as RDS

### 2021-10-29 code/processing_code/censusprocessing.Rmd
various edits
update exploration figures, manuscript additions, adding data index and populating readme files
finishing project pt 3

@@
 
-saveRDS(fips, "fips.RDS")
+saveRDS(fips, here("data", "processed_data", "fips.RDS"))
 ```
@@
 #Downloaded data from http://NHGIS.org
-install.packages("ipumsr")
 library(ipumsr)

### 2021-10-29 code/processing_code/wrfprocessing.Rmd
finishing project pt 3

@@
 ```{r}
+library(here)
+library(geodist) #to calculate distance from sample site to each water reclamation facility
+library(dplyr) #for data processing
+library(tidyverse)
+
 #load MP data

### 2021-10-29 products/manuscript/Manuscript.Rmd
finishing project pt 3

@@
 
-Additional data [write up further explanation]:
- - small area population via census zipcode
- - UOWN bacteria level data
- - local land cover
- - distance from water reclamation facility (WRF)
-
 ## Questions/Hypotheses to be addressed
@@
 
-My ideas for how to analyze this data include using non-parametric statistical tests, correlation matrices, and modeling (?) to characterize and compare observations.
-
 # Methods
@@
 
-- Methods for sample collection and processing for UOWN MP water samples
+### Description of study area
+The Upper Oconee Watershed is located in Northeast Georgia, spanning from Gainesville on the northern end to Lake Sinclair in Milledgeville, Georgia, on the southern end.
+The Upper Oconee Watershed encompasses 7590 km2, and includes more than 4000 km of streams [@Fisher].
+Land within the Upper Oconee Watershed is historically agricultural, but has been subject to increasing urbanization in the past two decades; the study area contains developed and densely populated urban residential areas [@Cho; @Fisher].
+The Oconee Watershed is one of 14 river basins in the state of Georgia; the Oconee River feeds into the Altamaha River, which eventually deposits into the Atlantic Ocean off the Georgia coast (source: UOWN website).
+The Oconee River begins where the North and Middle Oconee Rivers meet, south of Athens, Georgia.
+The North and Middle Oconee Rivers begin in Lula, GA, and Brassleton, GA, respectively.
+
+The Upper Oconee Watershed is regularly monitored by the Upper Oconee Watershed Network (UOWN), an Athens-based monitoring, education, advocacy, and recreation organization.
+Quarterly sampling events occur throughout the year, and the present research project was made possible via collaboration and eventual partnership with UOWN and the quarterly sampling events.
+The UOWN study area consists of 600 km2.
+
+### Experimental design: sample collection and processing
+1-liter grab samples of water were collected quarterly from Fall 2020 through Summer 2021, in November, February, April, and July by UOWN community members.
+Samples were collected from 25 to 30 sites on each sampling date.
+The number of sites was decreased from previous sampling events due to COVID-19 public health restrictions with volunteers.
+
+Samples were processed via vacuum filtration, which was performed in 150 mL volumes in duplicate for each site.
+Samples were pumped through gridded 0.45 micron membranes (Whatman) and immediately covered and stored covered in plastic petri dishes at room temperature for up to 8 weeks.
+Filters were visually inspected using a dissecting microscope with magnification ranging from [INSERT].
+Microplastic particles were identified based on the following characteristic features: unnatural homogenous color, unnatural shape, homogenous texture, homogenous width (no tapered end) (CITE).
+Microplastic concentration (No/L) was calculated by dividing the total microplastic particle count by the sample volume.
+Duplicates (samples A and B for each site) were averaged to obtain a singular concentration value for each sample site.
+
+### Controls
+To evaluate contamination in the lab, blank filters were used to gauge airborne contamination and contamination within processing methods.
+Blank filters were placed in uncovered petri dishes on each work bench during sample filtration and processing.
+DI and MilliQ water were filtered in 150 mL volumes, in duplicate, to produce controls to represent potential contamination from filtration processes.
+The average contamination per sample was = INSERT.
 
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
 
@@
 
-Import of additional data
- - census - American Community Survey
- - NGIS land cover
- - zip code - Bing reverse geocoding
- - WRF locations: located on Google Maps, coordinates recorded in Excel spreadsheet
- - UOWN bacteria data (citation/attribution available on UOWN.org)
+To supplement my analysis, additional data was imported as detailed below.
+
+### UOWN Quarterly Data
+Public data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org
+### Population
+NHGIS [@Mansen]
+### Land Cover
+NHGIS [@Mansen]
+### Water Reclamation Facilities
+Coordinates for Athens-Clarke county's three water reclamation facilities: North Oconee, Middle Oconee, and Cedar Creek were manually located via Google Maps.
 
@@
 Since I am keeping this document in manuscript style, I will keep processing/exploration/analysis code separate.
-_Use a combination of text/tables/figures to explore and describe your data.
-You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-FIgures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data._
-
-_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-Plots produced here can be scatterplots, boxplots, violinplots, etc.
-Tables can be simple 2x2 tables or larger ones._
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
 
 Table \@ref(tab:summarytable) shows a table summarizing the data.
+```{r summarytable,  echo=FALSE}
 
-```{r summarytable,  echo=FALSE}
-#saving template for formatting later on
+#summarytable=readRDS("../../results/summarytable.rds")
+#knitr::kable(summarytable, caption = 'Data summary table.')
 
-#resulttable=readRDS("../../results/summarytable.rds")
-#knitr::kable(resulttable, caption = 'Data summary table.')
+#I eventually want a nice looking summary statistic table. This is a place holder for now  
 ```
 
-Figure \@ref(fig:resultfigure) shows a scatterplot figure produced by one of the R scripts.
+### Distribution of Microplastic Concentration
 
-```{r resultfigure,  fig.cap='Analysis figure.', echo=FALSE}
-#knitr::include_graphics("../../results/resultfigure.png")
+Figure \@ref(fig:distribution) shows a histogram of microplastic concentration observations.
+The minimum concentration is 16.67 particles/L and the maximum is 1193.33 particles/L.
+The mean concentration is 104.39 particles/L, and the median is 66.67 particles/L.
+
+```{r distribution,  fig.cap='Distribution of Microplastic Concentration', echo=FALSE}
+knitr::include_graphics("../../results/distribution.png")
+```
+
+Microplastic concentrations remained in similar ranges throughout the study period. Figure \@ref(fig:concentrationbydate) shows a boxplot of concentrations by sample date.
+
+```{r concentrationbydate,  fig.cap='Particles/L by Sample Date', echo=FALSE}
+knitr::include_graphics("../../results/concentrationbydate.png")
+```
+
+There is some seasonal variation in concentration at each individual site. Figure \@ref(fig:seasonalconc-bysite) shows a plot of concentrations at each site.
+
+```{r seasonalconc-bysite,  fig.cap='Seasonal Variation in Particles/L', echo=FALSE}
+knitr::include_graphics("../../results/seasonalconc_bysite.png")
+```
+
+There are similar microplastic levels throughout the watersheds within the Upper Oconee.
+Some watersheds experienced greater variation in microplastic levels than other watersheds.
+Figure \@ref(fig:conc-bywatershed) shows the microplastic concentrations by watershed.
+
+```{r conc-bywatershed,  fig.cap='Watershed Microplastic Concentrations', echo=FALSE}
+knitr::include_graphics("../../results/conc_bywatershed.png")
+```
+
+Figure \@ref(fig:watershed-overtime) shows a line graph of the mean watershed microplastic concentrations at each seasonal sampling date.
+
+```{r watershed-overtime,  fig.cap='Microplastic Concentration Over Time', echo=FALSE}
+knitr::include_graphics("../../results/watershed_overtime.png")
+```
+
+### Predictors
+Population, land cover/use, and bacteria levels are hypothesized predictors of microplastic concentration.
+Figure \@ref(fig:particlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.
+
+```{r particlesvpop,  fig.cap="Particles/L vs Population", echo=FALSE}
+knitr::include_graphics("../../results/particlesvpop.png")
+```
+
+```{r logmpvcfu,  fig.cap='Log particles/L vs CFU', echo=FALSE}
+knitr::include_graphics("../../results/logmpvcfu.png")
+```
+
+Figure \@ref(fig:matrix) and Figure \@ref(fig:landmatrix) show correlation matrices for the hypothesized predictor and for the different categories of land use.
+
+```{r matrix,  fig.cap='Predictor matrix', echo=FALSE}
+knitr::include_graphics("../../results/matrix.png")
+```
+
+```{r landmatrix,  fig.cap='Land cover matrix', echo=FALSE}
+knitr::include_graphics("../../results/landmatrix.png")
 ```
@@
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
+Preliminary modeling reveals that there is not a strong relationship between microplastic concentration and population level. Figure \@ref(fig:m-concvpop-no-outlier) demonstrates a linear model fit.
 
-Example table \@ref(tab:resulttable) shows a table summarizing a linear model fit.
+```{r m-concvpop-no-outlier,  fig.cap='Concentration vs Population Linear Model', echo=FALSE}
+knitr::include_graphics("../../results/m_concvpop_no_outlier.png")
+```
 
-```{r resulttable,  echo=FALSE}
-#resulttable=readRDS("../../results/resulttable.rds")
-#knitr::kable(resulttable, caption = 'Linear model fit table.')
+Figure \@ref(fig:m-logconcvcfu) shows a linear model of microplastic concentration vs CFU (both variables log-transformed).
+
+```{r m-logconcvcfu,  fig.cap='Concentration vs CFU Linear Model', echo=FALSE}
+knitr::include_graphics("../../results/m_logconcvcfu.png")
+```
+
+Figure \@ref(fig:concvturbidity) demonstrates a linear model of particles/L vs turbidity.
+
+```{r m_concvturbidity,  fig.cap='Concentration vs Turbidity Linear Model', echo=FALSE}
+knitr::include_graphics("../../results/m_concvturbidity.png")
+```
+
+Table \@ref(tab:lmfittable) shows a table summarizing a linear model fit predicting particles/L with 6 predictors.
+
+```{r lmfittable,  echo=FALSE}
+lm_fit_table=readRDS("../../results/lm_fit_table.rds")
+knitr::kable(lm_fit_table, caption = 'Linear model fit table.')
 ```

### 2021-11-02 code/processing_code/censusland_code.Rmd
add API key, remove incorrect files
added API key to censusprocessing.Rmd to attempt to fix run error

@@
 ```{r}
-#specify from which package a function is to be used
-
 #tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE) 

### 2021-11-02 code/processing_code/censusprocessing.Rmd
add API key, remove incorrect files
added API key to censusprocessing.Rmd to attempt to fix run error

@@
 ```{r}
+#load census API key
+tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE, overwrite = TRUE)
+
 #create function

### 2021-11-11 code/analysis_code/Untitled.Rmd
pt 4 - continue modeling, analysis

---
title: "land_model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-11-11 code/analysis_code/analysis.Rmd
ungroup data for analysis
pt 4 - continue modeling, analysis

@@
 ```{r}
-data <- data %>% select("particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist")
-
+data <- data %>% select("particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist", "watershed")
+#save data subset
+saveRDS(data, file = here("data", "processed_data", "data_subset.rds"))
 #linear regression model specification set up
@@
 comp8_9
+
 ```
@@
 
-Why are there so many rows resulting? When we did the analysis last week, only one row was the result, and it was just rmse and provided a value around 1.
-
 #trying model evaluation with turbidity - linear regression
@@
 Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")
-
+t1
 #save figure
@@
 
-Here, I am still confused about why so many rmse values are resulting.
-Earlier, I was getting multiple rmse values but thought it was because I was building a model with all predictors.
-Here, I am using one predictor but still getting multiple rmse values.
-If I remember correctly, there was only one rmse value in the exercises we worked on in the previous weeks.
+Here, we see that a model with turbidity as the single predictor produces an RMSE of 58. This is significantly better than the full, all-predictor model which had an RMSE of 101.
+
+Going forward, perhaps we could add in additional predictors one at a time to determine what predictors may improve the model performance.

### 2021-11-11 code/analysis_code/land_model.Rmd
pt 4 - continue modeling, analysis

---
title: "land_model"
output: html_document
---

In this portion of the analysis, we will include the land use data for the sample sites to see if the additional predictors can improve our model fit.

# load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(rpart)
library(glmnet)
library(ranger)
library(rpart.plot)  # for visualizing a decision tree
library(vip)         # for variable importance plots

#load cleaned land data. 
data <- readRDS(here("data", "processed_data", "athensland.rds"))
data <- data %>% select(-"site", 
                        -"lat",
                        -"long",
                        -"date")
```

# split data into train and test subsets
```{r}
# set seed for reproducible analysis (instead of random subset each time)
set.seed(123)
#subset 3/4 of data as training set
data_split <- initial_split(data, 
                            prop = 7/10, 
                            strata = particles_l) #stratify by MP concentration for balanced outcome

#save sets as data frames
train_data <- training(data_split)
test_data <- testing(data_split)
```

# Cross validation
We want to perform 5-fold CV, 5 times repeated
```{r}
#create folds (resample object)
set.seed(123)
folds <- vfold_cv(train_data, 
                  v = 5, 
                  repeats = 5,
                  strata = particles_l) #folds is set up to perform our CV

#linear model set up
lm_mod <- linear_reg() %>% 
            set_engine('lm') %>% 
            set_mode('regression')

#create recipe for data and fitting and make dummy variables
MP_rec <- recipe(particles_l ~ ., data = train_data) %>% step_dummy(all_nominal())

#workflow set up
MP_wflow <- 
  workflow() %>% add_model(lm_mod) %>% add_recipe(MP_rec)

#use workflow to prepare recipe and train model with predictors
MP_fit <- 
  MP_wflow %>% fit(data = train_data)

#extract model coefficient
MP_fit %>% extract_fit_parsnip() %>% tidy()
```

# Null model performance
```{r}
#recipe for null model
null_train_rec <- recipe(particles_l ~ 1, data = train_data) #predicts mean of outcome

#null model workflow incorporating null model recipe
null_wflow <- workflow() %>% add_model(lm_mod) %>% add_recipe(null_train_rec)

# I want to check and make sure that the null model worked as it was supposed to, so I want to view the predictions and make sure they are all the mean of the outcome
#get fit for train data using null workflow
nullfittest <- null_wflow %>% fit(data = train_data)
#get predictions based on null model
prediction <- predict(nullfittest, train_data)
test_pred <- predict(nullfittest, test_data)
#the predictions for the train and test data are all the same mean value, so this tells us the null model was set up properly

#Now, we'll use fit_resamples based on the tidymodels tutorial for CV/resampling (https://www.tidymodels.org/start/resampling/)
#fit model with training data
null_fit_train <- fit_resamples(null_wflow, resamples = folds)

#get results
metrics_null_train <- collect_metrics(null_fit_train)
#RMSE for null train fit is 109.4

#repeat for test data
null_test_rec <- recipe(particles_l ~ 1, data = test_data) #predicts mean of outcome
null_test_wflow <- workflow() %>% add_model(lm_mod) %>% add_recipe(null_test_rec) #sets workflow with new test recipe
null_fit_test <- fit_resamples(null_test_wflow, resamples = folds) #performs fit
metrics_null_test <- collect_metrics(null_fit_test) #gets fit metrics
#RMSE for null test fit is 109.4
```

# Model tuning and fitting
Include:
1. Model specification
2. Workflow definition
3. Tuning grid specification
4. Tuning w/ cross-validation + `tune_grid()`

## LASSO model
```{r}
#based on tidymodels tutorial: case study
#cross validation
set.seed(123)
cell_folds <- vfold_cv(train_data)

#model specification
lasso <- linear_reg(penalty = tune()) %>% set_engine("glmnet") %>% set_mode("regression")

#set workflow
lasso_wf <- workflow() %>% add_model(lasso) %>% add_recipe(MP_rec)

#tuning grid specification
lasso_grid <- tibble(penalty = 10^seq(-3, 0, length.out = 30))

#tuning with CV and tune_grid
lasso_res <- lasso_wf %>% tune_grid(resamples = cell_folds, 
                                    grid = lasso_grid, 
                                    control = control_grid(save_pred = TRUE), 
                                    metrics = metric_set(rmse))
#view model metrics
lasso_res %>% collect_metrics()

#select top models
top_lasso <- 
  lasso_res %>% show_best("rmse") %>% arrange(penalty)
top_lasso #view

#see best lasso
best_lasso <- lasso_res %>% select_best()
best_lasso #view

#finalize workflow with top model
lasso_final_wf <- lasso_wf %>% finalize_workflow(best_lasso)

#fit model with finalized WF
lasso_fit <- lasso_final_wf %>% fit(train_data)
```

### LASSO plots
```{r}
#diagnostics
autoplot(lasso_res)
#calculate residuals
lasso_resid <- lasso_fit %>%
  augment(train_data) %>% #this will add predictions to our df
  select(.pred, particles_l) %>%
  mutate(.resid = particles_l - .pred) #manually calculate residuals

#model predictions from tuned model vs actual outcomes
lasso_pred_plot <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
  labs(title = "Predictions vs Actual Outcomes: LASSO", x = "Particles/L Outcome", y = "Particles/L Prediction")
lasso_pred_plot

#plot residuals vs predictions
lasso_resid_plot <- ggplot(lasso_resid, aes(y = .resid, x = .pred)) + geom_point() + 
  labs(title = "Predictions vs Residuals: LASSO", x = "Particles/L Prediction", y = "Residuals")
lasso_resid_plot #view plot

#compare to null model
metrics_null_train #view null RMSE for train data
lasso_res %>% show_best(n=1) #view RMSE for best lasso model

```

RMSE for best LASSO model on the land data is 127.5, compared to 109.4 RMSE for the null model. This tells us our model is not good.

In the perfect outcome/prediction plot, you hope to see data scattered along a 45 degree diagonal line. Instead, data is clustered. However, the scales of the x and y axis are different which may impact how we are viewing the data. Let's visualize with a better scale.

```{r}
#model predictions from tuned model vs actual outcomes SCALE ADJUSTED
lasso_pred_plot_adj <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
  labs(title = "Predictions vs Actual Outcomes: LASSO", x = "Particles/L Outcome", y = "Particles/L Prediction") + ylim(0, 1000) + xlim(0,1000)
lasso_pred_plot_adj
```

Our outcome vs prediction plot isn't terrible but also isn't great. It looks like predictions are consistently higher than the actual outcomes, since the points fall below the hypothetical 45 degree diagonal line that would represent a perfect fit.

#plot for how the number of predictors included in the LASSO model changes with the tuning parameter
```{r}
x <- lasso_fit$fit$fit$fit 
plot(x, "lambda")

summary(best_lasso)

coefficients <- lasso_fit %>% extract_fit_parsnip() %>% tidy()
lasso_fit %>% extract_fit_parsnip() %>% tidy()

```

It appears that the addition of land use data does not make our model more informative. In fact, it increases the RMSE to be higher than that of the null model.

```

### 2021-11-11 code/analysis_code/modeling.Rmd
pt 4 - continue modeling, analysis

--- 2021-11-11 code/analysis_code/land_model.Rmd
@@
 ---
-title: "land_model"
+title: "modeling"
 output: html_document
@@
 
-In this portion of the analysis, we will include the land use data for the sample sites to see if the additional predictors can improve our model fit.
+In this portion of the analysis, we will use the LASSO method for regularization of the model.
 
@@
 
-#load cleaned land data. 
-data <- readRDS(here("data", "processed_data", "athensland.rds"))
-data <- data %>% select(-"site", 
-                        -"lat",
-                        -"long",
-                        -"date")
+#path to data
+#note the use of the here() package and not absolute paths
+data_location <- here::here("data","processed_data","processeddata.rds")
+
+#load cleaned data. 
+data <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
 ```
@@
 
-RMSE for best LASSO model on the land data is 127.5, compared to 109.4 RMSE for the null model. This tells us our model is not good.
+next I want to see what predictors are used in the best lasso model. also need to compare to null model
 
-In the perfect outcome/prediction plot, you hope to see data scattered along a 45 degree diagonal line.
-Instead, data is clustered.
-However, the scales of the x and y axis are different which may impact how we are viewing the data.
-Let's visualize with a better scale.
-
-```{r}
-#model predictions from tuned model vs actual outcomes SCALE ADJUSTED
-lasso_pred_plot_adj <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
-  labs(title = "Predictions vs Actual Outcomes: LASSO", x = "Particles/L Outcome", y = "Particles/L Prediction") + ylim(0, 1000) + xlim(0,1000)
-lasso_pred_plot_adj
-```
-
-Our outcome vs prediction plot isn't terrible but also isn't great.
-It looks like predictions are consistently higher than the actual outcomes, since the points fall below the hypothetical 45 degree diagonal line that would represent a perfect fit.
+RMSE for best LASSO model is 100.5, compared to 109.4 RMSE for the null model. Still need to check included predictors for the best LASSO model.
 
@@
 
-It appears that the addition of land use data does not make our model more informative. In fact, it increases the RMSE to be higher than that of the null model.
+By looking at the predictors used in this fit, we see there are a lot of uninformative predictors included, such as latitude, longitude, site ID, FIPS, etc.
+Let's look at a subset of predictors to build a model.
+
+Try modeling with data subset, including these variables: "particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist".
+We'll do this in a separate Rmd document called `subset_modeling.Rmd`
 

### 2021-11-11 code/analysis_code/subset_modeling.Rmd
pt 4 - continue modeling, analysis

--- 2021-11-11 code/analysis_code/land_model.Rmd
@@
 ---
-title: "land_model"
+title: "subset_modeling"
 output: html_document
@@
 
-In this portion of the analysis, we will include the land use data for the sample sites to see if the additional predictors can improve our model fit.
+In this portion of the analysis, we will perform modeling on a smaller subset of data in attempt to improve model performance.
 
@@
 
-#load cleaned land data. 
-data <- readRDS(here("data", "processed_data", "athensland.rds"))
-data <- data %>% select(-"site", 
-                        -"lat",
-                        -"long",
-                        -"date")
+#path to data
+#note the use of the here() package and not absolute paths
+data_location <- here::here("data","processed_data","processeddata.rds")
+
+#load cleaned USBSET data. 
+data <- readRDS(here("data", "processed_data", "data_subset.rds"))
 ```
@@
 
-RMSE for best LASSO model on the land data is 127.5, compared to 109.4 RMSE for the null model. This tells us our model is not good.
+next I want to see what predictors are used in the best lasso model. also need to compare to null model
+
+RMSE for best LASSO model on the subset data is 113.09, compared to 109.4 RMSE for the null model.
+This is definitely not a well-performing model, since we want a better RMSE for any model we build than the RMSE for the null model.
 
@@
 
-Our outcome vs prediction plot isn't terrible but also isn't great.
-It looks like predictions are consistently higher than the actual outcomes, since the points fall below the hypothetical 45 degree diagonal line that would represent a perfect fit.
+Now that we have adjusted the scale, our outcome vs prediction plot looks decent, and there don't appear to be systematic deviations.
+This indicates that our LASSO model may be flexible enough for our data.
 
@@
 
-It appears that the addition of land use data does not make our model more informative. In fact, it increases the RMSE to be higher than that of the null model.
+Try modeling with data subset, including these variables: "particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist".
+
+The results we get from the initial machine learning modeling and from the modeling of this subset data are not great. Let's think of some ways that we can go about improving our model. 
+- An option is variable/predictor removal. We tried that with the subset data. We could attempt further subsets going forward. 
+- Variable transformation. Some microplastic studies use log transformation on the concentration values to get a normal distribution. 
+- Alternatives: account for different distribution.
+The distribution of microplastic concentration may be considered more Poisson distributed than normally distributed.
+First, we can test to see whether it falls in either of these distribution categories, and based on the results we can apply a GLM with Poisson distribution.
+(Tranformation could make intrepretability more difficult - may want to avoid)
+- however, regularization may accomplish our goals more efficiently - we are advised that subset selection is not always a great idea and doesn't neccesarily do the best job of addressing overfitting
+- could try ridge regression for alternate form of regularization 
+- could try decision tree (though we learned these usually don't perform as well as LASSO)
+
+```{r}
+#check for Poisson distribution
+#https://stackoverflow.com/questions/59809960/how-do-i-know-if-my-data-fit-a-poisson-distribution-using-r
+dispersion_test <- function(x) 
+{
+  res <- 1-2 * abs((1 - pchisq((sum((x - mean(x))^2)/mean(x)), length(x) - 1))-0.5)
+
+  cat("Dispersion test of count data:\n",
+      length(x), " data points.\n",
+      "Mean: ",mean(x),"\n",
+      "Variance: ",var(x),"\n",
+      "Probability of being drawn from Poisson distribution: ", 
+      round(res, 3),"\n", sep = "")
+
+  invisible(res)
+}
+
+set.seed(1)
+x <- data$particles_l
+
+dispersion_test(x)
+```
 

### 2021-11-11 products/manuscript/Manuscript.Rmd
pt 4 - continue modeling, analysis

@@
     number_sections: true
+    global_numbering: true
+    fig_caption: yes
     #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted

### 2021-11-11 code/processing_code/ARCHIVE.Rmd
fix censusprocesseing.Rmd, add to manuscript

---
title: "ARCHIVE"
output: html_document
---

# convert lat long to zip
```{r}
#path to data
data_location <- here::here("data","raw_data","zippop.xlsx")

#load population data. 
zippop <- read_excel(data_location)

# use Bing API to convert coordinates to corresponding zip code
zip<-revgeo(longitude=combodata$long, 
                            latitude=combodata$lat, 
                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip")

# add zip variable to combodata
combodata <- combodata %>% mutate(revgeo(longitude=combodata$long, 
                            latitude=combodata$lat, 
                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip") %>%
                              select("zip"))

#add zipcode information to my original combined dataset
popdata <- left_join(combodata, zippop, by = c("zip" = "zipcode"))
```

### 2021-11-11 code/processing_code/censusprocessing.Rmd
fix censusprocesseing.Rmd, add to manuscript

--- 2021-10-28 code/processing_code/censusprocessing.Rmd
@@
 
-#path to data
-data_location <- here::here("data","raw_data","zippop.xlsx")
-
-#load population data. 
-zippop <- read_excel(data_location)
-
-#load MP data
-data_location1 <- here::here("data","processed_data","combodata.rds")
-
 #load data. 
-combodata <- readRDS(data_location1)
-```
-
-# convert lat long to zip
-```{r}
-
-# use Bing API to convert coordinates to corresponding zip code
-zip<-revgeo(longitude=combodata$long, 
-                            latitude=combodata$lat, 
-                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip")
-
-# add zip variable to combodata
-combodata <- combodata %>% mutate(revgeo(longitude=combodata$long, 
-                            latitude=combodata$lat, 
-                            provider = 'bing', API = 'AoRXmf4yiTmdhPeNAXK2kjZBU1cSCI-_8fWgIzPu45Rh6PMu_TP1Rx_ZTuNEdvQJ', output="frame", item = "zip") %>%
-                              select("zip"))
+combodata <- readRDS(here("data", "processed_data", "combodata.rds"))
 ```
@@
 
-# add population levels based on zip code
+# renaming variable differentiate use in processing
 ```{r}
-#add zipcode information to my original combined dataset
-popdata <- left_join(combodata, zippop, by = c("zip" = "zipcode"))
-
+popdata <- combodata
 #save data as RDS
-popdata_loc <- here::here("data", "processed_data", "popdata.rds")
-
-saveRDS(popdata, file = popdata_loc)
+saveRDS(popdata, file = here("data", "processed_data", "popdata.rds"))
 
@@
 Let's try to use FIPS codes (which include tract info) to obtain population data that might be more informative.
-# Converting coordinates to FIPS
+I've removed the previous code that converted coordinates to zip code and got population based on zip code.
+This code is now located in `ARCHIVE.Rmd`
+
+# Loading data with FIPS variable, converted from coordinates
+
+FIPS codes were obtained through a census API run on the local computer, then saved locally and copied into the project directory under `raw_data` for this analysis.
+This ensures that the code will be reproducible and consistent even if the API changes in the future.
 ```{r}
-#create function
-latlong2fips <- function(latitude, longitude) {
-  url <- "https://geo.fcc.gov/api/census/block/find?format=json&latitude=%f&longitude=%f"
-  url <- sprintf(url, latitude, longitude)
-  json <- RCurl::getURL(url)
-  json <- RJSONIO::fromJSON(json)
-  as.character(json$Block['FIPS'])
-}
-#vectorize so function will take more than one argument
-latlong2fips_vec <- Vectorize(latlong2fips, vectorize.args = "latitude", "longitude")
-
-fips <- popdata %>% rowwise() %>% mutate(FIPS = latlong2fips_vec(lat, long))
-
-saveRDS(fips, "fips.RDS")
+fips <- readRDS(here("data", "raw_data", "fips_copy.RDS"))
+#saveRDS(fips, here("data", "processed_data", "fips.RDS"))
 ```
@@
 ```{r}
-#Downloaded data from http://NHGIS.org
-install.packages("ipumsr")
+#Downloaded data from http://NHGIS.org and placed in "raw_data" for this R project. 
 library(ipumsr)
@@
 #Remove first ALUBE001 column and zip, population columns
-tractpopdata <- c %>% select(-"ALUBE001.x", -"zip", -"population")
+tractpopdata <- c %>% select(-"ALUBE001.x")
 tractpopdata <- rename(tractpopdata, "population" = "ALUBE001.y")

### 2021-11-11 products/manuscript/Manuscript.Rmd
fix censusprocesseing.Rmd, add to manuscript

@@
 
+## Background
 In the past two decades, the study of microplastics has expanded exponentially.
@@
 
+### Pollution Sources
+Microplastics are classified by source as either primary or secondary.
+Primary microplastics are those which are originally manufactured as microplastics, such as microbeads; primary microplastics are often found in personal care products, industrial scrubbers used for cleaning, plastic powder used for molding, and virgin resin used in plastic production (EFSA, 2016).
+Secondary microplastics begin as larger plastics and degrade into smaller pieces as the result of exposure to environmental stressors (e.g.
+sunlight, wind, water) (Murphy et al., 2016).
+For example, secondary microplastics occur when ultraviolet light exposure catalyzes photo-oxidation in plastic, leading to brittle pieces that are even more likely to break down (Wright & Kelly, 2017).
+A secondary microplastic source that is less obvious than the fragmentation of plastic debris is synthetic clothing.
+Clothes made of fibers such as nylon, polyester, rayon, or spandex shed microfibers during washing cycles; these fibers are later deposited into the environment through wastewater treatment plant effluent (Murphy et al., 2016).
+Most experts in the field agree that secondary microplastics are the most significant contributor to microplastic pollution, though a broad range of microplastic sources are recognized (Tibbetts et al., 2018).
+Researchers recognize wastewater treatment plant (WWTP) effluent as a major vector for both primary and secondary microplastics, and therefore WWTP effluent has been one area of focus in microplastics literature (Murphy et al., 2016; McCormick et al., 2016; Li et al., 2018, Tibbetts et al., 2018, Harrison et al., 2018).
+A team of scientists in Scotland examined WWTPs as a source of microplastics in the aquatic environment by sampling for microplastics at different stages of wastewater treatment; the team calculated a 98% reduction of microplastics from influent to effluent (Murphy et al., 2016).
+While an efficient removal rate, Murphy and collaborators estimate that 65 million microplastics are still released from the plant daily due to the large volume of effluent.
+Numerous other studies have evaluated WWTP effluent in relation to microplastic concentration as well (McCormick et al., 2016; Li et al., 2018, Tibbetts et al., 2018, Harrison et al., 2018).
+Other predictors of microplastics particle concentration that researchers consider include land use (i.e.
+natural, agricultural, industrial, developed, forest & woodland, etc.) and population (total size, density, dwellings) (Tibbetts et al., 2018; Barrows et al, 2018; Townsend et al., 2019; Yonkos et al, 2014).
+Compared to the ocean, freshwater is the dominant source of microplastic due to land-based industrial and domestic pollution (Luo et al., 2019).
+Rivers transport both point-source and diffuse pollution to the ocean, contributing an estimated 1.15 to 2.41 million tons of plastic waste annually (Lebreton et al, 2017).
+Factors including anthropogenic impact, pollution sources, and hydrodynamics affect the rate of microplastic accumulation and transport (Luo et al., 2019; CITE).
+Researchers continue to focus on identifying the variables influencing microplastic concentration in present day research (CITE).
+
+### Ecological Consequences
+Despite their small size, microplastics have significant and quantifiable impact on aquatic ecosystems (CITE).
+Since microplastics permeate both marine and freshwater environments, aquatic animals easily mistake the particles for food and ingest them, allowing the pollutant to accumulate in their bodies; studies have shown that organisms ingest microplastics either by accident or by mistaking them for food (Wright & Kelly, 2017; Carbery, O’Connor, & Palanisami, 2018, Tanaka & Takada, 2016).
+Microplastics have been identified in the digestive tracts of fish and birds of prey, in mussels and other bivalves, and in natural populations of a crustacean species (Tanaka & Takada, 2016; Carlin et al., 2020; van Cauwenberghe & Janssen, 2014; Devriese et al., 2015; Murray and Cowie, 2011).
+Microplastics are dangerous to organisms due to the physical accumulation of the particles themselves, and because they act as a transport mechanism for toxic chemicals in three main ways: (1) the plastic itself is toxic, (2) the plastic adsorbs and later releases environmental pollutants, and (3) the plastic leaches chemical additives (Mato et al, 2001; Bouwmeester et al., 2015).
+The hydrophobic surface and high surface area to volume ratio of microplastics causes the adsorption and concentration of organic pollutants such as polychlorinated bisphenols (PCBs), dichlorodiphenyldichloroethylene (DDE), and nonylphenol (NP) (Wright & Kelly, 2017; Carbery, O’Connor, & Palanisami, 2018).
+Evidence of trophic transfer of microplastics through the food chain suggests that microplastics and the toxins that accompany them pose a threat to aquatic life further up the food chain, and to human health through consumption of seafood (Barboza et al 2018, Carbery et al, 2018).
+
+### Relevance
 Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014;@Waller2017;@Peng2018].
@@
 Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].
+
+### Present Research
+	This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, and wastewater treatment plant proximity (CITE – who hypothesizes this?).
+We hypothesized that microplastics would be present in the Upper Oconee Watershed, that levels would be greater in more highly-trafficked areas with greater anthropogenic impact, and that microplastic pollution would correlate to other indicators of water quality (CITE).
 
@@
 # Discussion
+This exploratory pilot study demonstrates the presence of freshwater microplastic pollution in the Upper Oconee Watershed and makes an initial attempt to quantify the level of pollution throughout this region of Athens, GA.
+At the start of this project, there was little to no existing research on microplastics in Athens.
+Therefore, the data resulting from this study establish the presence of microplastics in the area [mention support of other studies showing MP in urban vs rural areas].
+
+Overall, concentration levels were found to be higher than expected based on some previous studies of regional freshwater microplastic pollution.
+However, the body of research on microplastics as a whole reflects mixed evidence of concentrations.
+Recorded levels in the literature range from 1 or fewer particles per liter at some sites to greater than 1000 particles per liter at others.
+There is growing evidence that microplastic pollution in permeates ecosystems in unexpected, supposedly pristine regions across the globe; one study found between 22 and 117 particles per liter in the Antarctic, demonstrating that microplastic particles are being found in places you would never expect.
+## Variation of Microplastic Concentration
+Limitations to the scope of this project include that it was designed as a student-led pilot with minimal funding.
+Visual identification via dissecting microscope was the method used due to its simplicity and accessibility; visual identification is used in most baseline microplastic studies, but can be up to 70% inaccurate and therefore is usually accompanied by spectroscopy technology which allows chemical identification of suspected microplastic particles and increasing the accuracy of quantification results.
+
+This year-long pilot study is followed by a continued partnership with the Upper Oconee Watershed Network, who will continue sampling for microplastics quarterly through citizen-science volunteer efforts.
+Areas of focus for the future include furthering temporal data collection, refining quantification methods, observing the relationship between microplastic levels and other water quality indicators, and exploring other hypothesized water quality predictors.
+While limited in scope, this study provides valuable initial data on the presence of microplastics in the region and establishes a basis for continued and refined research on the growing subject of freshwater microplastics.
 

### 2021-11-12 code/analysis_code/exploration.Rmd
edit manuscript, expand modeling
removed EDA plots from manuscript, added RF and tree models, created grid for results

@@
 library(GGally) #for producing correlation matrices
+library(gridExtra) #
 
@@
 
+Let's start with some summary statistics. We want to see the overall summary of microplastic concentration - mean, median, range, etc.
+
 # Summary statistics
@@
 
+We have a total of 109 sites. Particles/L ranges from 16 to 1193, with a mean of 104 and a median of 67. This tells us the data is right-skewed (we have outliers on the upper end of the data).
+
+Our observations can be categorized by sample date, since there were four different sample events that occurred. Let's see if there is any temporal relationship in the microplastic concentrations.
+
 # MP concentration by season
@@
 
-mpdata$date <- factor(mpdata$date , levels=c("nov_20", "feb_21", "apr_21", "july_21"))
+mpdata$date <- factor(mpdata$date , levels=c("nov_20", "feb_21", "apr_21", "july_21")) #reorder dates to appear chronologically
 
+# make boxplot for microplastic concentration by season
 p1 <- mpdata %>% ggplot(aes(x = date, y = particles_l)) + 
@@
 
+One of our main hypothesized predictors is population.
+Population size is a suppossed proxy for anthropogenic impact.
+We want to examine if microplastic levels increase in areas (census tracts) with greater populations.
+
 For all samples, let's compare microplastic level and local population.
@@
 I'm not seeing any obvious relationships here.
-Also, population values are appearing discrete rather than continous since there are a limited number of zip codes in the study area, and population was based on zip code.
+Also, population values are appearing discrete/in clusters rather than continous since there are a limited number of census tracts codes in the study area, and population was based on census tract.
+
+Another way to visualize microplastic concentration vs population may be to log transform the microplastic levels.
+This would normalize the data, and the log transformation method has been used in previously published studies that analyze microplastic levels.
 
@@
 ```
+
+This gives us a plot that is a little nicer to look at, but we still don't see a clear relationship between micrplastic level and local population.
+
+Another potential predictor is e.coli colony forming units.
+Similar to population count, e.coli levels may be a proxy for human impact/nearby pollution sources.
+We want to see if there is a relationship between microplastic pollution and bacteria levels at the sample sites.
 
@@
 
-Again, no obvious relationships here. How about if we looked with a logarithmic scale?
+Again, no obvious relationships here.
+We also see some pretty low values in this plot.
+How about if we looked with a logarithmic scale?
+Log transformation is often used for bacterial data, and can also be used with microplastic levels.
 
+# log(MP) vs log(CFU) log transformed
 ```{r}
@@
 
-Let's look at overall distribution of MP levels.
+The log transformation may make these values easier to work with. Still no clear relationships, however.
+
+Now, let's look at overall distribution of MP levels.
 
@@
 ```
+
+Most observations fall below 300 particles/L, with a few large outliers.
 
@@
 
-We see that most levels fall below 100 particles per liter. 
-# Summary statistics
-```{r}
-summary(mpdata$particles_l)
-```
+In this graph, we can more clearly see that most levels fall below 100 particles per liter.
 
-These summary statistics give us a little more descriptive information about the dataset.
-The other variables that I wanted to look at particles per liter alongside were land cover.
-The process for looking at land cover will be a little more complicated - there a key for variable names corresponding to land types that I need to look closer at, and I am also not sure if the scale of the land cover data will be appropriate.
-Currently, it is at the census FIPS code level (approximately county level).
-Since most of these samples occur within that same area, we don't have other reference areas for comparison.
-I'll have to think about what I can do with this information.
+Another way to look at relationships among predictors and outcome is through correlation matrices. These allow us to view multiple relationships at once.
 
@@
 
+There are no bright red or bright blue boxes here, so we don't neccesarily have strong relationships among variables.
+We see that temperature and e.coli are positively correlated, as are temperature and turbidity.
+There is a positive relationship between turbidity and population.
+The correlation between particles/L and population is very weak - nearly completely neutral.
+The only moderate correlation is between turbidity and particles/L.
+There appears to be a very weak negative correlation between e.coli and particles/L which is unexpected.
+However, we are basing this off of colors and not specific coefficients or p-values, so this is a vague estimate.
+
+Since we have a lot of different variables for land use, the correlation matrix is a good option to explore these and view many at the same time.
+
 # exploring land use with a correlation matrix
@@
 
+In this plot, we are mainly looking at the row containing particles/L (the very bottom row). We do not see any strong correlations between microplastic level and local land use.
+
+We also may want to see the variability of microplastic concentration at each individual site (rather than aggregate concentration) througout the seasons/sample dates. Let's take a look.
+
 # concentration at each site throughout seasons
@@
 
+In this plot, we see that some sites experienced small differences in concentration between seasons, while other experienced large differences between seasons.
+
+Similarly, let's see if there are any distictions between watersheds, since we can group the sample sites based on what watershed they are located in.
+
 # View concentration by watershed
@@
 ```
+
+A few watersheds exhibit more variation/range in microplastic level than other watersheds.
+This also may be due to the number of sites located in each watershed (some watersheds contain very few sites while others contain many).
+It may be useful to include an "n =" on this graph in the future.
+
+Instead of viewing a dot/point for each sample data, let's view the concentrations at each site throughout the year as a boxplot to visualize variability, averages, etc.
+
 # Boxplot for individual sites throughout season
@@
 ```
+
+There are a few sites that have quite a large range of values, though the majority of the sites have a small to moderate range.
+It must also be noted that not all sample sites were accounted for at each sample date.
+So, some sites may only have 1 or 2 observations, while other sites may have 4 observations.
+
+I want to visualize the change (or lack of change) in MP concentration over time.
+However, that is a lot to take in looking at individual sites.
+Let's view the average concentrations in each watershed for each sample date.
+Then, we can look at those watershed averages over time throughout the study period.
+
 # concentration over time by watershed
@@
 
+Looking at the data this way, we do see somewhat of a pattern.
+It appears that the range of microplastic concentration decreases from Fall 2020 to Summer 2021.
+This may be due to some particularly large outliers in the earlier sample dates.
+Also, we may consider human error.
+Since I counted these microplastics myself, it may be that I initially over-counted, including false positives (particles that I mistook for microplastics, when they may have been organic matter, etc).
+Essentially, methods may have improved/refined over time.
+
 Nice!

### 2021-11-12 code/analysis_code/modeling.Rmd
edit manuscript, expand modeling
removed EDA plots from manuscript, added RF and tree models, created grid for results

@@
 
-In this portion of the analysis, we will use the LASSO method for regularization of the model.
+In this portion of the analysis, we will use the LASSO method for regularization of the model. We'll also try out a decision tree
 
@@
 library(vip)         # for variable importance plots
+library(gridExtra)
 
@@
 lasso_pred_plot <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
-  labs(title = "Predictions vs Actual Outcomes: LASSO", x = "Particles/L Outcome", y = "Particles/L Prediction")
+  labs(title = "LASSO: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
 lasso_pred_plot
@@
 lasso_resid_plot <- ggplot(lasso_resid, aes(y = .resid, x = .pred)) + geom_point() + 
-  labs(title = "Predictions vs Residuals: LASSO", x = "Particles/L Prediction", y = "Residuals")
+  labs(title = "LASSO: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
 lasso_resid_plot #view plot
@@
 
-#plot for how the number of predictors included in the LASSO model changes with the tuning parameter
+# plot for how the number of predictors included in the LASSO model changes with the tuning parameter
 ```{r}
@@
 
+# Decision tree
+Next, we'll try out a decision tree model
+
+## Decision tree
+```{r}
+#going based off of tidymodels tutorial: tune parameters
+#since we already split our data into test and train sets, we'll continue to use those here. they are `train_data` and `test_data`
+
+#model specification
+tune_spec <- 
+  decision_tree(
+    cost_complexity = tune(),
+    tree_depth = tune()
+  ) %>% 
+  set_engine("rpart") %>% 
+  set_mode("regression")
+tune_spec
+
+#tuning grid specification
+tree_grid <- grid_regular(cost_complexity(),
+                          tree_depth(),
+                          levels = 5)
+tree_grid
+
+#cross validation
+set.seed(123)
+cell_folds <- vfold_cv(train_data)
+
+#workflow
+set.seed(123)
+
+tree_wf <- workflow() %>%
+  add_model(tune_spec) %>%
+  add_recipe(MP_rec)
+
+#model tuning with `tune_grid()`
+tree_res <- 
+  tree_wf %>% 
+  tune_grid(
+    resamples = cell_folds,
+    grid = tree_grid
+    )
+tree_res %>% collect_metrics()
+#Here we see 25 candidate models, and the RMSE and Rsq for each
+tree_res %>% autoplot() #view plot
+
+#select the best decision tree model
+best_tree <- tree_res %>% select_best("rmse")
+best_tree #view model details
+
+#finalize model workflow with best model
+tree_final_wf <- tree_wf %>%
+  finalize_workflow(best_tree)
+
+#fit model
+tree_fit <- 
+  tree_final_wf %>% fit(train_data)
+tree_fit
+
 ```
+
+### Decision tree plots
+```{r}
+#diagnostics
+autoplot(tree_res)
+#calculate residuals - originally got stuck trying out lots of different methods for this.
+took inspiration from Zane's code to manually calculate residuals rather than using some of the built in functions that I could not get to cooperate
+tree_resid <- tree_fit %>%
+  augment(train_data) %>% #this will add predictions to our df
+  select(.pred, particles_l) %>%
+  mutate(.resid = particles_l - .pred) #manually calculate residuals
+
+#model predictions from tuned model vs actual outcomes
+tree_pred_plot <- ggplot(tree_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
+  labs(title = "Decision Tree: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
+tree_pred_plot
+
+#plot residuals vs predictions
+tree_resid_plot <- ggplot(tree_resid, aes(y = .resid, x = .pred)) + geom_point() + 
+  labs(title = "Decision Tree: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
+tree_resid_plot #view plot
+
+#compare to null model
+metrics_null_train #view null RMSE for train data
+tree_res %>% show_best(n=1) #view RMSE for best decision tree model
+```
+
+The best decision tree has an RMSE of 102.2, which is better than the null model RMSE of 109.
+However, the decision tree did not perform as well well as the Lasso regularized model, which had an RMSE of about 100.
+
+# Random forest
+Finally, let's try out a random forest.
+
+### split data into train and test subsets with only COMPLETE CASES
+```{r}
+data %>% is.na() %>% summary() #check out missing cases
+complete <- data %>% na.omit() #save complete cases
+complete <- complete %>% select(-"visual_score")
+
+# set seed for reproducible analysis (instead of random subset each time)
+set.seed(123)
+#subset 3/4 of data as training set
+data_split <- initial_split(complete, 
+                            prop = 7/10, 
+                            strata = particles_l) #stratify by MP concentration for balanced outcome
+
+#save sets as data frames
+train_data <- training(data_split)
+test_data <- testing(data_split)
+
+#reset recipe with adjusted, complete data
+MP_rec <- recipe(particles_l ~ ., data = train_data) %>% step_dummy(all_nominal())
+#reset cell_folds
+set.seed(123)
+cell_folds <- vfold_cv(train_data)
+```
+
+## Random forest
+```{r}
+#based on tidymodels tutorial: case study
+library(parallel)
+cores <- parallel::detectCores()
+cores
+#model specification
+r_forest <- rand_forest(mtry = tune(), min_n = tune(), trees = tune()) %>% set_engine("ranger", num.threads = cores) %>% set_mode("regression")
+
+#set workflow
+r_forest_wf <- workflow() %>% add_model(r_forest) %>% add_recipe(MP_rec)
+
+#tuning grid specification
+rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6), min_n = c(40,50,60), trees = c(500,1000)  )
+#what we will tune:
+r_forest %>% parameters()
+
+#tuning with CV and tune_grid
+
+r_forest_res <- 
+  r_forest_wf %>%
+  tune_grid(resamples = cell_folds, 
+            grid = rf_grid, 
+            control = control_grid(save_pred = TRUE), 
+            metrics = metric_set(rmse))
+
+#view top models
+r_forest_res %>% show_best(metric = "rmse")
+
+#view plot of models performance
+autoplot(r_forest_res)
+
+#select best model
+rf_best <- r_forest_res %>% select_best(metric = "rmse")
+rf_best
+
+#finalize workflow with top model
+rf_final_wf <- r_forest_wf %>% finalize_workflow(rf_best)
+
+#fit model with finalized WF
+rf_fit <- rf_final_wf %>% fit(train_data)
+```
+Here, we get an error message that all models failed, note: missing value in visual_score. Let's go back and use only complete cases for the random forest.
+
+The issue was in "cell_folds" which had originated earlier in the Rmd and didn't include the updated complete case dataset.
+Added a line of code to reset cell_folds at this stage of the Rmd in order to use complete cases only.
+We still get some error messages but are able to run the model.
+
+The best random forest models have RMSEs of about 108.
+This is slightly better than the null model, but not significantly.
+The random forest model does not perform as well as the decision tree or LASSO model.
+
+### Random forest plots
+```{r}
+#diagnostics
+autoplot(r_forest_res)
+#calculate residuals
+rf_resid <- rf_fit %>%
+  augment(train_data) %>% #this will add predictions to our df
+  select(.pred, particles_l) %>%
+  mutate(.resid = particles_l - .pred) #manually calculate residuals
+
+#model predictions from tuned model vs actual outcomes
+rf_pred_plot <- ggplot(rf_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
+  labs(title = "Random Forest: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
+rf_pred_plot
+
+#plot residuals vs predictions
+rf_resid_plot <- ggplot(rf_resid, aes(y = .resid, x = .pred)) + geom_point() + 
+  labs(title = "Random Forest: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
+rf_resid_plot #view plot
+
+#compare to null model
+metrics_null_train #view null RMSE for train data
+r_forest_res %>% show_best(n=1) #view RMSE for best decision tree model
+```
+
+```{r}
+model_grid <- grid.arrange(lasso_pred_plot, lasso_resid_plot, rf_pred_plot, rf_resid_plot, tree_pred_plot, tree_resid_plot, ncol = 3)
+
+ggsave(model_grid, filename = here("results", "model_grid.png"))
+```
+
+I'll need to go back and edit this image later so that the titles aren't cut off and the axis scales match up properly so the plots look right.
+I might also want to add a 45 degree diagonal line to the outcomes vs predictions plot, and some other features to improve appearance.

### 2021-11-12 products/manuscript/Manuscript.Rmd
edit manuscript, expand modeling
removed EDA plots from manuscript, added RF and tree models, created grid for results

@@
 
-There are similar microplastic levels throughout the watersheds within the Upper Oconee.
-Some watersheds experienced greater variation in microplastic levels than other watersheds.
-Figure \@ref(fig:conc-bywatershed) shows the microplastic concentrations by watershed.
-
-```{r conc-bywatershed,  fig.cap='Watershed Microplastic Concentrations', echo=FALSE}
-knitr::include_graphics("../../results/conc_bywatershed.png")
-```
-
 Figure \@ref(fig:watershed-overtime) shows a line graph of the mean watershed microplastic concentrations at each seasonal sampling date.
@@
 Population, land cover/use, and bacteria levels are hypothesized predictors of microplastic concentration.
-Figure \@ref(fig:particlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.
+Figure \@ref(fig:logparticlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.
 
-```{r particlesvpop,  fig.cap="Particles/L vs Population", echo=FALSE}
-knitr::include_graphics("../../results/particlesvpop.png")
+```{r logparticlesvpop,  fig.cap="Particles/L vs Population", echo=FALSE}
+knitr::include_graphics("../../results/logparticlesvpop.png")
 ```
@@
 
-Figure \@ref(fig:matrix) and Figure \@ref(fig:landmatrix) show correlation matrices for the hypothesized predictor and for the different categories of land use.
-
-```{r matrix,  fig.cap='Predictor matrix', echo=FALSE}
-knitr::include_graphics("../../results/matrix.png")
-```
-
-```{r landmatrix,  fig.cap='Land cover matrix', echo=FALSE}
-knitr::include_graphics("../../results/landmatrix.png")
-```
-
 ## Full analysis
 
-Preliminary modeling reveals that there is not a strong relationship between microplastic concentration and population level. Figure \@ref(fig:m-concvpop-no-outlier) demonstrates a linear model fit.
+Preliminary modeling reveals that there is not a strong relationship between microplastic concentration and population level. Figure \@ref(fig:m-logconcvpop) demonstrates a linear model fit.
 
-```{r m-concvpop-no-outlier,  fig.cap='Concentration vs Population Linear Model', echo=FALSE}
-knitr::include_graphics("../../results/m_concvpop_no_outlier.png")
+```{r m-logconcvpop-no-outlier,  fig.cap='Concentration vs Population Linear Model', echo=FALSE}
+knitr::include_graphics("../../results/m_logconcvpop.png")
 ```
@@
 knitr::kable(lm_fit_table, caption = 'Linear model fit table.')
+```
+
+Beyond the basic linear model, we have applied additional methods to improve model performance, including LASSO regularization and building decision trees and random forests for model comparison.
+The predictions, outcomes, and residuals resulting from each type of plot are demonstrated in Figure \@ref(fig:model-grid)
+
+```{r model-grid, fig.cap='Model Quality', echo=FALSE}
+knitr::include_graphics("../../results/model_grid.png")
 ```

### 2021-11-12 products/manuscript/Supplement.Rmd
edit manuscript, expand modeling
removed EDA plots from manuscript, added RF and tree models, created grid for results

--- 2021-11-11 products/manuscript/Manuscript.Rmd
@@
 ---
-title: "Plastic in the Urban Freshwater Environment"
+title: "Supplementary Materials: Plastic in the Urban Freshwater Environment"
 subtitle: "An Exploratory Study of Microplastics in the Upper Oconee Watershed via Community-Based Efforts"
@@
 ---
-
-# Abstract
-With growing awareness of plastic pollution and its detrimental environmental impacts, there has been increased attention on the study of microplastics.
-These small debris result from the weathering and breakdown of plastic waste from urban runoff, agricultural runoff, ocean dumping, stormwater, waste management, and other sources.
-Microplastics pose a threat to both aquatic life and human health due to their chemical makeup, their build-up in the environment, and their suspected long-term biological effects.
-As their ubiquity becomes more apparent, so does the need for accessible, affordable, and effective methods for the widespread study and quantification of microplastic pollution.
-
-This study takes place in Athens, GA, an urban area which houses a population of about 127,000.
-The project relies on local volunteer citizen science efforts for water sampling, processing, and analysis.
-Given the local and regional ecological significance of the Oconee River—including its transport of freshwater to the Atlantic Ocean—this study aims to collect relevant data regarding microplastics in the Upper Oconee Watershed, which includes 2 rivers and 15 creeks in Athens-Clarke County.
-
-Through community-based sample collection and analysis, volunteers quantified microplastics through membrane filtration and visual identification.
-Preliminary data consistently show that local freshwater microplastic levels are greater than the U.S.
-average in the Upper Oconee Watershed, a portion of the Oconee River basin with both a high population size and high population density relative to other regions of Georgia.
-Average and median concentrations of 95 particles/L and 46 particles/L, respectively, were recorded throughout one year of seasonal sampling.
-Further visual and statistical analysis will confirm the project’s results, and discussion of these results in context will reveal the efficacy of this community-based approach to the quantification of microplastics.
-
-# Introduction
-
-In the past two decades, the study of microplastics has expanded exponentially.
-Since the earliest paper mentioning small plastic particles in the ocean was published in the 1970s, before the term “microplastics” even existed, researchers across the globe have quantified microplastic particle levels in both marine and freshwater environments, assessed potential drivers and predictors of microplastic pollution, examined the presence of microplastics in the food chain, and explored the potential human health impacts of ingested microplastics [@Carpenter1972].
-Though the body of research on microplastics has grown in recent years, many knowledge gaps remain regarding this pollutant.
-The term microplastics was first coined in 2004, and today microplastics are defined as plastic particles smaller in size than 5 mm, though some studies examine particles as large as 20 mm [@Thompson2004].
-Years of sampling, experimentation, and progress in research methods tell researchers that microplastics are a persistent class of pollutant and are found worldwide in a diverse assemblage of forms [@Barboza2018].
-
-Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica [@Free2014;@Waller2017;@Peng2018].
-Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era [@Zalasiewicz2016].
-These models also suggest that microplastic concentrations will only increase in the future [@Geyer2017;@Barboza2018;@Prokic2019;@Everaert2018].
-In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans [@Barboza2018].
-Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment [@Andrady2011;@Li2018].
-Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems [@Free2014;@Li2018;@Browne2010;@Lebreton2017].
-However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area [@Besseling2017;@Dris2015].
-Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies [@Lebreton2017].
-
-## Description of data and data source
-
-The data that I am using for this project is from my personal undergraduate research project on freshwater microplastics in Athens, GA.
-The study is ongoing and was conducted in collaboration with the Upper Oconee Watershed Network (UOWN) via their quarterly sampling events.
-There is one year’s worth of quarterly data, with the following sample dates: November 2020, February 2021, April 2021, and July 2021.
-The data include variables such as sample site, latitude and longitude, watershed location, and duplicate A and B counts of microplastics observed per filter.
-For the most recent sampling date, July 2021, there is additional data where a second counting session was performed, in order to assess the comparability of parallel counts.
-There are about 136 observations of ~10 variables in total in the dataset.
-The independent variables – site location, coordinates, watershed location, etc – were retrieved from UOWN.
-Values for microplastic counts were recorded based on visual identification via a dissecting microscope, performed by myself and by a small team of <10 volunteers.
-
-## Questions/Hypotheses to be addressed
-The questions that I want to answer with this data include the following:
- - Can we characterize microplastics in Athens, GA using a citizen science approach? 
- -	What are the levels of microplastics like in Athens, GA overall? 
- -	How do microplastic levels differ based on location within the region?
-
-Questions that I want to answer that will require gathering more data using coordinates/watershed info include:
- -	Is microplastic concentration impacted by local land use? 
- -	Is microplastic concentration impacted by localized population levels in specific neighborhoods?
- -	Does wastewater treatment plant effluent from nearby facilities increase microplastic concentration at affected sites?
- -	Are microplastic levels correlated to bacteria levels measured at the same sites? 
-These questions also relate to hypothesized predictors of microplastic pollution: population level, land use, wastewater treatment plant effluent, and microbial water quality.
-
-# Methods
-
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-
-### Description of study area
-The Upper Oconee Watershed is located in Northeast Georgia, spanning from Gainesville on the northern end to Lake Sinclair in Milledgeville, Georgia, on the southern end.
-The Upper Oconee Watershed encompasses 7590 km2, and includes more than 4000 km of streams [@Fisher].
-Land within the Upper Oconee Watershed is historically agricultural, but has been subject to increasing urbanization in the past two decades; the study area contains developed and densely populated urban residential areas [@Cho; @Fisher].
-The Oconee Watershed is one of 14 river basins in the state of Georgia; the Oconee River feeds into the Altamaha River, which eventually deposits into the Atlantic Ocean off the Georgia coast (source: UOWN website).
-The Oconee River begins where the North and Middle Oconee Rivers meet, south of Athens, Georgia.
-The North and Middle Oconee Rivers begin in Lula, GA, and Brassleton, GA, respectively.
-
-The Upper Oconee Watershed is regularly monitored by the Upper Oconee Watershed Network (UOWN), an Athens-based monitoring, education, advocacy, and recreation organization.
-Quarterly sampling events occur throughout the year, and the present research project was made possible via collaboration and eventual partnership with UOWN and the quarterly sampling events.
-The UOWN study area consists of 600 km2.
-
-### Experimental design: sample collection and processing
-1-liter grab samples of water were collected quarterly from Fall 2020 through Summer 2021, in November, February, April, and July by UOWN community members.
-Samples were collected from 25 to 30 sites on each sampling date.
-The number of sites was decreased from previous sampling events due to COVID-19 public health restrictions with volunteers.
-
-Samples were processed via vacuum filtration, which was performed in 150 mL volumes in duplicate for each site.
-Samples were pumped through gridded 0.45 micron membranes (Whatman) and immediately covered and stored covered in plastic petri dishes at room temperature for up to 8 weeks.
-Filters were visually inspected using a dissecting microscope with magnification ranging from [INSERT].
-Microplastic particles were identified based on the following characteristic features: unnatural homogenous color, unnatural shape, homogenous texture, homogenous width (no tapered end) (CITE).
-Microplastic concentration (No/L) was calculated by dividing the total microplastic particle count by the sample volume.
-Duplicates (samples A and B for each site) were averaged to obtain a singular concentration value for each sample site.
-
-### Controls
-To evaluate contamination in the lab, blank filters were used to gauge airborne contamination and contamination within processing methods.
-Blank filters were placed in uncovered petri dishes on each work bench during sample filtration and processing.
-DI and MilliQ water were filtered in 150 mL volumes, in duplicate, to produce controls to represent potential contamination from filtration processes.
-The average contamination per sample was = INSERT.
-
-## Data import and cleaning
-
-Since I am performing importing and cleaning in a separate document, in this section I will provide an overview of my steps/methods for wrangling the data, without actual code in this section.
-
-To supplement my analysis, additional data was imported as detailed below.
-
-### UOWN Quarterly Data
-Public data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org
-### Population
-NHGIS [@Mansen]
-### Land Cover
-NHGIS [@Mansen]
-### Water Reclamation Facilities
-Coordinates for Athens-Clarke county's three water reclamation facilities: North Oconee, Middle Oconee, and Cedar Creek were manually located via Google Maps.
 
@@
 ```
-
-# Results
-
-# Discussion
-
-## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
-
-## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
-
-## Conclusions
-_What are the main take-home messages?_
-
-# References

### 2021-11-13 code/analysis_code/mapping.R
code files

@@ -1,0 +1,194 @@
+#Load Libraries
+library(raster)
+library(mapview)
+library(tidyverse)
+library(dplyr)
+library(sf)
+library(spatstat)
+library(sp)
+library(here)
+
+#Load Data 
+sitedata = readRDS(here("data", "processed_data", "athensland.rds"))
+
+#Save sites as a simple shapefile
+sites_sf = na.omit(sitedata) %>% st_as_sf(coords = c('long', 'lat'))
+st_crs(sites_sf) <- 4326
+st_crs(sites_sf)
+#Base Map
+
+#First, we want to create a basemap of Athens-Clarke County, and include hydrology of the
+#watershed if possible 
+
+#Found shapefiles for hydrology on ACC Website here: https://data-athensclarke.opendata.arcgis.com/datasets/hydropoly?geometry=-83.511%2C33.912%2C-83.280%2C33.962
+#acc_hydro = st_read("./acc_map/Hydropoly.shp")
+#acc_hydro = st_transform(acc_hydro,CRS("+proj=longlat"))
+#mapview(acc_hydro)  
+
+mapview(sites_sf, xcol = "long", ycol = "lat", zcol = "particles_l")
+
+#GA Hydrology accessed from NHD USGS Database
+ga_hydro = st_read(here("data", "raw_data", "shape", "NHDFlowline.shp"))
+sites_sf = st_as_sf(sitedata, coords = c('long', 'lat'))
+st_crs(sites_sf) <- 4326
+st_crs(ga_hydro) <- 4326
+ga_hydro = st_transform(ga_hydro, crs = 4326)
+
+mapview(ga_hydro)
+
+ggplot() + geom_sf(data = ga_hydro, stat = "sf", position = "identity", mapping = aes()) + 
+  xlim(-83.6, -83.3) + ylim(33.8, 34.0) + 
+  geom_point(data = sites_sf, aes(x = long, y = lat, color = particles_l))
+
+map <- mapview(ga_hydro, layer.name = "Upper Oconee Watershed", color = "skyblue3", alpha = .7) + 
+  mapview(sites_sf, zcol = "particles_l", legend = mapviewGetOption("legend"), layer.name = "Microplastics per L")
+
+mapshot(map, url = here("results", "map.html"))
+
+## Seasonal Subsets
+fall <- sitedata %>% filter(date == "nov_20")
+winter <- sitedata %>% filter(date == "feb_21")
[…]

### 2021-11-13 products/manuscript/Manuscript.Rmd
results

@@
 Microplastics are classified by source as either primary or secondary.
-Primary microplastics are those which are originally manufactured as microplastics, such as microbeads; primary microplastics are often found in personal care products, industrial scrubbers used for cleaning, plastic powder used for molding, and virgin resin used in plastic production (EFSA, 2016).
+Primary microplastics are those which are originally manufactured as microplastics, such as microbeads; primary microplastics are often found in personal care products, industrial scrubbers used for cleaning, plastic powder used for molding, and virgin resin used in plastic production [@EFSA2016].
 Secondary microplastics begin as larger plastics and degrade into smaller pieces as the result of exposure to environmental stressors (e.g.
-sunlight, wind, water) (Murphy et al., 2016).
-For example, secondary microplastics occur when ultraviolet light exposure catalyzes photo-oxidation in plastic, leading to brittle pieces that are even more likely to break down (Wright & Kelly, 2017).
+sunlight, wind, water) [@Murphy2016].
+For example, secondary microplastics occur when ultraviolet light exposure catalyzes photo-oxidation in plastic, leading to brittle pieces that are even more likely to break down [@Wright2017].
 A secondary microplastic source that is less obvious than the fragmentation of plastic debris is synthetic clothing.
-Clothes made of fibers such as nylon, polyester, rayon, or spandex shed microfibers during washing cycles; these fibers are later deposited into the environment through wastewater treatment plant effluent (Murphy et al., 2016).
-Most experts in the field agree that secondary microplastics are the most significant contributor to microplastic pollution, though a broad range of microplastic sources are recognized (Tibbetts et al., 2018).
-Researchers recognize wastewater treatment plant (WWTP) effluent as a major vector for both primary and secondary microplastics, and therefore WWTP effluent has been one area of focus in microplastics literature (Murphy et al., 2016; McCormick et al., 2016; Li et al., 2018, Tibbetts et al., 2018, Harrison et al., 2018).
-A team of scientists in Scotland examined WWTPs as a source of microplastics in the aquatic environment by sampling for microplastics at different stages of wastewater treatment; the team calculated a 98% reduction of microplastics from influent to effluent (Murphy et al., 2016).
+Clothes made of fibers such as nylon, polyester, rayon, or spandex shed microfibers during washing cycles; these fibers are later deposited into the environment through wastewater treatment plant effluent [@Murphy2016].
+Most experts in the field agree that secondary microplastics are the most significant contributor to microplastic pollution, though a broad range of microplastic sources are recognized [@Tibbetts2018].
+Researchers recognize wastewater treatment plant (WWTP) effluent as a major vector for both primary and secondary microplastics, and therefore WWTP effluent has been one area of focus in microplastics literature [@Murphy2016;@McCormick2014a;@Li2018;@Tibbetts2018;@Harrison2018].
+A team of scientists in Scotland examined WWTPs as a source of microplastics in the aquatic environment by sampling for microplastics at different stages of wastewater treatment; the team calculated a 98% reduction of microplastics from influent to effluent [@Murphy2016].
 While an efficient removal rate, Murphy and collaborators estimate that 65 million microplastics are still released from the plant daily due to the large volume of effluent.
-Numerous other studies have evaluated WWTP effluent in relation to microplastic concentration as well (McCormick et al., 2016; Li et al., 2018, Tibbetts et al., 2018, Harrison et al., 2018).
+Numerous other studies have evaluated WWTP effluent in relation to microplastic concentration as well [@McCormick2014a;@Li2018,@Tibbetts2018,@Harrison2018].
 Other predictors of microplastics particle concentration that researchers consider include land use (i.e.
-natural, agricultural, industrial, developed, forest & woodland, etc.) and population (total size, density, dwellings) (Tibbetts et al., 2018; Barrows et al, 2018; Townsend et al., 2019; Yonkos et al, 2014).
-Compared to the ocean, freshwater is the dominant source of microplastic due to land-based industrial and domestic pollution (Luo et al., 2019).
-Rivers transport both point-source and diffuse pollution to the ocean, contributing an estimated 1.15 to 2.41 million tons of plastic waste annually (Lebreton et al, 2017).
-Factors including anthropogenic impact, pollution sources, and hydrodynamics affect the rate of microplastic accumulation and transport (Luo et al., 2019; CITE).
+natural, agricultural, industrial, developed, forest & woodland, etc.) and population (total size, density, dwellings) [@Tibbetts2018; @Barrows2018;@Townsend2019; @Yonkos2014].
+Compared to the ocean, freshwater is the dominant source of microplastic due to land-based industrial and domestic pollution [@Luo2019].
+Rivers transport both point-source and diffuse pollution to the ocean, contributing an estimated 1.15 to 2.41 million tons of plastic waste annually [@Lebreton2017].
+Factors including anthropogenic impact, pollution sources, and hydrodynamics affect the rate of microplastic accumulation and transport [@Luo2019; CITE].
 Researchers continue to focus on identifying the variables influencing microplastic concentration in present day research (CITE).
@@
 Despite their small size, microplastics have significant and quantifiable impact on aquatic ecosystems (CITE).
-Since microplastics permeate both marine and freshwater environments, aquatic animals easily mistake the particles for food and ingest them, allowing the pollutant to accumulate in their bodies; studies have shown that organisms ingest microplastics either by accident or by mistaking them for food (Wright & Kelly, 2017; Carbery, O’Connor, & Palanisami, 2018, Tanaka & Takada, 2016).
-Microplastics have been identified in the digestive tracts of fish and birds of prey, in mussels and other bivalves, and in natural populations of a crustacean species (Tanaka & Takada, 2016; Carlin et al., 2020; van Cauwenberghe & Janssen, 2014; Devriese et al., 2015; Murray and Cowie, 2011).
-Microplastics are dangerous to organisms due to the physical accumulation of the particles themselves, and because they act as a transport mechanism for toxic chemicals in three main ways: (1) the plastic itself is toxic, (2) the plastic adsorbs and later releases environmental pollutants, and (3) the plastic leaches chemical additives (Mato et al, 2001; Bouwmeester et al., 2015).
-The hydrophobic surface and high surface area to volume ratio of microplastics causes the adsorption and concentration of organic pollutants such as polychlorinated bisphenols (PCBs), dichlorodiphenyldichloroethylene (DDE), and nonylphenol (NP) (Wright & Kelly, 2017; Carbery, O’Connor, & Palanisami, 2018).
-Evidence of trophic transfer of microplastics through the food chain suggests that microplastics and the toxins that accompany them pose a threat to aquatic life further up the food chain, and to human health through consumption of seafood (Barboza et al 2018, Carbery et al, 2018).
+Since microplastics permeate both marine and freshwater environments, aquatic animals easily mistake the particles for food and ingest them, allowing the pollutant to accumulate in their bodies; studies have shown that organisms ingest microplastics either by accident or by mistaking them for food [@Wright2017;@Carbery2018;@Tanaka2016].
+Microplastics have been identified in the digestive tracts of fish and birds of prey, in mussels and other bivalves, and in natural populations of a crustacean species [@Tanaka2016;@Carlin2020;@VanCauwenberghe2014;@Devriese2015;@Murray2011].
+Microplastics are dangerous to organisms due to the physical accumulation of the particles themselves, and because they act as a transport mechanism for toxic chemicals in three main ways: (1) the plastic itself is toxic, (2) the plastic adsorbs and later releases environmental pollutants, and (3) the plastic leaches chemical additives [@Mato2001;@Bouwmeester2015].
+The hydrophobic surface and high surface area to volume ratio of microplastics causes the adsorption and concentration of organic pollutants such as polychlorinated bisphenols (PCBs), dichlorodiphenyldichloroethylene (DDE), and nonylphenol (NP) [@Wright2017;@Carbery2018].
+Evidence of trophic transfer of microplastics through the food chain suggests that microplastics and the toxins that accompany them pose a threat to aquatic life further up the food chain, and to human health through consumption of seafood [@Barboza2018;@Carbery2018].
 
@@
 ### Present Research
-	This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, and wastewater treatment plant proximity (CITE – who hypothesizes this?).
-We hypothesized that microplastics would be present in the Upper Oconee Watershed, that levels would be greater in more highly-trafficked areas with greater anthropogenic impact, and that microplastic pollution would correlate to other indicators of water quality (CITE).
+This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, and wastewater treatment plant proximity (@Murphy2016).
+We hypothesized that microplastics would be present in the Upper Oconee Watershed, that levels would be greater in more highly-trafficked areas with greater anthropogenic impact, and that microplastic pollution would correlate to other indicators of water quality [@McCormick2014a].
 
@@
 Filters were visually inspected using a dissecting microscope with magnification ranging from [INSERT].
-Microplastic particles were identified based on the following characteristic features: unnatural homogenous color, unnatural shape, homogenous texture, homogenous width (no tapered end) (CITE).
+Microplastic particles were identified based on the following characteristic features: unnatural homogenous color, unnatural shape, homogenous texture, homogenous width (no tapered end)[@Hidalgo-Ruz2012].
 Microplastic concentration (No/L) was calculated by dividing the total microplastic particle count by the sample volume.
@@
 
+### Mapping
+I created some maps for concentrations at each sample sites throughout the seasons (saved as html files "map_season.html") for now, but I am unsure of how best to display these maps in the manuscript word document output.
+```{r}
+#mapview::knit_print.mapview(here("results", "map_spring.html"))
+```
+
 ## Full analysis
@@
 
+Based on the results of the three different models, the LASSO model is the best option for this dataset, though the minimal difference in RMSE when compared to the null model suggests that even though LASSO is the better model method compared to others, it still does not produce a great model for predicting microplastic concentration.
+
 # Results

### 2021-11-14 code/analysis_code/mapping.R
mapping

@@ -1,3 +1,7 @@
+#Intro
+#This script generates maps of sample sites for each season of sampling, color coded to correspond to MP concentration.
+#This script was developed with the help of Megan Lott
+
 #Load Libraries
 library(raster)
 library(mapview)
@@ -11,12 +15,6 @@ library(here)
 #Load Data 
 sitedata = readRDS(here("data", "processed_data", "athensland.rds"))
 
-#Save sites as a simple shapefile
-sites_sf = na.omit(sitedata) %>% st_as_sf(coords = c('long', 'lat'))
-st_crs(sites_sf) <- 4326
-st_crs(sites_sf)
-#Base Map
-
 #First, we want to create a basemap of Athens-Clarke County, and include hydrology of the
 #watershed if possible 
 
@@ -25,25 +23,6 @@ st_crs(sites_sf)
 #acc_hydro = st_transform(acc_hydro,CRS("+proj=longlat"))
 #mapview(acc_hydro)  
 
-mapview(sites_sf, xcol = "long", ycol = "lat", zcol = "particles_l")
-
-#GA Hydrology accessed from NHD USGS Database
-ga_hydro = st_read(here("data", "raw_data", "shape", "NHDFlowline.shp"))
-sites_sf = st_as_sf(sitedata, coords = c('long', 'lat'))
-st_crs(sites_sf) <- 4326
-st_crs(ga_hydro) <- 4326
-ga_hydro = st_transform(ga_hydro, crs = 4326)
-
-mapview(ga_hydro)
-
-ggplot() + geom_sf(data = ga_hydro, stat = "sf", position = "identity", mapping = aes()) + 
-  xlim(-83.6, -83.3) + ylim(33.8, 34.0) + 
-  geom_point(data = sites_sf, aes(x = long, y = lat, color = particles_l))
-
-map <- mapview(ga_hydro, layer.name = "Upper Oconee Watershed", color = "skyblue3", alpha = .7) + 
-  mapview(sites_sf, zcol = "particles_l", legend = mapviewGetOption("legend"), layer.name = "Microplastics per L")
-
-mapshot(map, url = here("results", "map.html"))
 
 ## Seasonal Subsets
 fall <- sitedata %>% filter(date == "nov_20")
@@ -58,33 +37,24 @@ sites_sf = na.omit(fall) %>% st_as_sf(coords = c('long', 'lat'))
 st_crs(sites_sf) <- 4326
 st_crs(sites_sf)
[…]

### 2021-11-17 code/analysis_code/mapping.R
Remove mapping files

@@ -1,109 +1,0 @@
-#Intro
-#This script generates maps of sample sites for each season of sampling, color coded to correspond to MP concentration.
-#This script was developed with the help of Megan Lott
-
-#Load Libraries
-library(raster)
-library(mapview)
-library(tidyverse)
-library(dplyr)
-library(sf)
-library(spatstat)
-library(sp)
-library(here)
-
-#Load Data 
-sitedata = readRDS(here("data", "processed_data", "athensland.rds"))
-
-#First, we want to create a basemap of Athens-Clarke County, and include hydrology of the
-#watershed if possible 
-
-#Found shapefiles for hydrology on ACC Website here: https://data-athensclarke.opendata.arcgis.com/datasets/hydropoly?geometry=-83.511%2C33.912%2C-83.280%2C33.962
-#acc_hydro = st_read("./acc_map/Hydropoly.shp")
-#acc_hydro = st_transform(acc_hydro,CRS("+proj=longlat"))
-#mapview(acc_hydro)  
-
-
-## Seasonal Subsets
-fall <- sitedata %>% filter(date == "nov_20")
-winter <- sitedata %>% filter(date == "feb_21")
-spring <- sitedata %>% filter(date == "apr_21")
-summer <- sitedata %>% filter(date == "july_21")
-     
-## Fall 2020 Map
-                         
-#Save sites as a simple shapefile
-sites_sf = na.omit(fall) %>% st_as_sf(coords = c('long', 'lat'))
-st_crs(sites_sf) <- 4326
-st_crs(sites_sf)
-#Base Map
-#view sites only
-mapview(sites_sf, xcol = "long", ycol = "lat", zcol = "particles_l")
-
-#GA Hydrology accessed from NHD USGS Database
-#We only need to set ga_hydro once - from here on we just edit the samples sites by season
-ga_hydro = st_read(here("data", "raw_data", "shape", "NHDFlowline.shp"))
-sites_sf = st_as_sf(fall, coords = c('long', 'lat'))
-st_crs(sites_sf) <- 4326
-st_crs(ga_hydro) <- 4326
-ga_hydro = st_transform(ga_hydro, crs = 4326)
-
-#view hydrology map
-mapview(ga_hydro)
-
-#sites and hydrology
-map_fall <- mapview(ga_hydro, layer.name = "Upper Oconee Watershed", color = "skyblue3", alpha = .7) + 
[…]

### 2021-11-18 products/proposal.Rmd
Recover manuscript

---
title: "Project Proposal"
output: html_document
---

```{r, include=FALSE}
#load needed packages. make sure they are installed.
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
```

# Project Part 1
The data that I will use for this project is from my personal undergraduate research project on freshwater microplastics in Athens, GA. The study is ongoing and was conducted in collaboration with the Upper Oconee Watershed Network (UOWN) via their quarterly sampling events. There is one year’s worth of quarterly data, with the following sample dates: November 2020, February 2021, April 2021, and July 2021. The data include variables such as sample site, latitude and longitude, watershed location, and duplicate A and B counts of microplastics observed per filter. For the most recent sampling date, July 2021, there is additional data where a second counting session was performed, in order to assess the comparability of parallel counts. There are about 136 observations of ~10 variables in total in the dataset. The independent variables – site location, coordinates, watershed location, etc – were retrieved from UOWN. Values for microplastic counts were recorded based on visual identification via a dissecting microscope, performed by myself and by a small team of <10 volunteers.

The questions that I want to answer with this data include the following:
 - Can we characterize microplastics in Athens, GA using a citizen science approach? 
 -	What are the levels of microplastics like in Athens, GA overall? 
 -	How do microplastic levels differ based on location within the region?

Questions that I want to answer that will require gathering more data using coordinates/watershed info include:
 -	Is microplastic concentration impacted by local land use? 
 -	Is microplastic concentration impacted by localized population levels in specific neighborhoods?
 -	Does wastewater treatment plant effluent from nearby facilities increase microplastic concentration at affected sites?
 -	Are microplastic levels correlated to bacteria levels measured at the same sites? 
These questions also relate to hypothesized predictors of microplastic pollution: population level, land use, wastewater treatment plant effluent, and microbial water quality.

My ideas for how to analyze this data include using non-parametric statistical tests, correlation matrices, and modeling (?) to characterize and compare observations.

There are just a few aspects of data collection for the July quarterly (which I will finish up in the next week or so) but I already have most of my data. Note that the July data (about 1/4 of the data) is still being finalized and is not currently included in the data below.

## Data Preview
```{r}
#path to data
data_location <- here::here("data","raw_data","mpdata_ALL.xlsx")

rawdata <- readxl::read_excel(data_location)

dplyr::glimpse(rawdata)

summary(rawdata)
```

I have been working on a preliminary introduction to this study, which provides background information on the data and may be helpful. My working draft is below (improved system for references coming soon).

## Background
In the past two decades, the study of microplastics has expanded exponentially. Since the earliest paper mentioning small plastic particles in the ocean was published in the 1970s, before the term “microplastics” even existed, researchers across the globe have quantified microplastic particle levels in both marine and freshwater environments, assessed potential drivers and predictors of microplastic pollution, examined the presence of microplastics in the food chain, and explored the potential human health impacts of ingested microplastics (Carpenter et al., 1972). Though the body of research on microplastics has grown in recent years, many knowledge gaps remain regarding this pollutant. The term microplastics was first coined in 2004, and today microplastics are defined as plastic particles smaller in size than 5 mm, though some studies examine particles as large as 20 mm (Thompson et al., 2004; Yang et al., 2019; Hidalgo et al., 2012). Years of sampling, experimentation, and progress in research methods tell researchers that microplastics are a persistent class of pollutant and are found worldwide in a diverse assemblage of forms (Barboza et al., 2018).

Through decades of research, the scientific community proves that microplastics are ubiquitous, having identified microplastics in oceans, rivers, lakes, and other water bodies across the globe, including remote mountain lakes, the bottom of the ocean, and in ocean sediments offshore of Antarctica (Free et al., 2014; Waller et al 2017, Peng et al., 2018). Modeling of current and future plastic production and pollution indicates that plastics are so pervasive that they have been considered as a geological indicator of the proposed Anthropocene era (Zalasiewicz et al., 2016). These models also suggest that microplastic concentrations will only increase in the future (Geyer, Jambeck, & Law, 2017; Barboza et al., 2018, Prokic et al., 2019; Evereart et al., 2018). In a 2018 paper, Barboza and collaborators state that the increase in environmental microplastics is particularly concerning due to the contaminants’ small size, the limited technology for quantifying their presence, and their potential for adversely affecting both marine biota and humans (Barboza et al., 2018). Further, research shows that microplastic pollution is largely due to land-based sources; ocean-based sources such as commercial fishing, vessels, and other activities, contribute only 20% of the total plastic debris, while terrestrial activity contributes the other 80% of pollution in the marine environment (Andrady et al., 2011; Li et al., 2018). Microplastics with various terrestrial origins mainly enter the marine environment via rivers, releasing most microplastics to the ocean and retaining some in freshwater systems (Free et al., 2014; Li et al., 2018; Browne et al., 2010; Lebreton et al., 2017). However, a majority of microplastics research to date has focused on marine settings rather than freshwater, indicating that more studies are required in this area (Besseling et al., 2018; Dris et al., 2015). Examining freshwater systems as the dominant source of microplastics will lead to a better overall understanding of microplastics pollution input and therefore insight into the scale of the issue and mitigation strategies (Lebreton et al., 2017).

### 2021-11-29 code/analysis_code/analysis.Rmd
project pt 5

@@
 library(gridExtra) #for exporting tables
+library(gridExtra) #for savings images in grids
 
@@
 m3 <- ggplot(data, aes(x = population, y = particles_l)) + geom_point() +
-  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Population", x = "Population", y = "Log(Particles/Liter)")
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "MP Concentration vs. Population", x = "Population", y = "Log(Particles/Liter)")
 
@@
 m5 <- ggplot(data, aes(x = e.coli.cfu, y = particles_l)) + geom_point() + scale_x_continuous(trans='log10') +
-  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. E. coli CFU", x = "Log(CFU/mL)", y = "Log(Particles/Liter)")
+  scale_y_continuous(trans='log10') + geom_smooth(method = lm) + labs(title = "MP Concentration vs. E. coli CFU", x = "Log(CFU/mL)", y = "Log(Particles/Liter)")
 
@@
 
-Trying code from above again with the complete dataset
+Let's try the code from above again with the complete dataset
+
 # Modeling with complete cases only
@@
 ```
-
-Here, we have a VERY large RSS, which I think indicates a poor model fit for our data.
-Let's try with some variables removed.
-If an RSS of 0 is a perfect model, our RSS of 2 million is either a *very bad* model or I am making some sort of error/misinterpreting(?).
 
@@
 #visualize linear model
-t1 <- ggplot(data, aes(x = turbidity.ntu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs.
-Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")
+t1 <- ggplot(data, aes(x = turbidity.ntu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration vs. Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")
 t1
@@
 Going forward, perhaps we could add in additional predictors one at a time to determine what predictors may improve the model performance.
+
+# linear model for distance from WRF
+```{r}
+#estimating/training linear model
+wrf_fit1 <- lm_mod %>% fit(particles_l ~ dist, data = data)
+#view summary of fit
+tidy(wrf_fit1)
+
+#visualize linear model
+wrf1 <- ggplot(data, aes(x = dist, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration vs. WRF Distance", x = "WRF Distance (m)", y = "Particles/Liter")
+wrf1
+#save figure
+ggsave(filename = here("results","m_concvwrfdist.png"), plot=wrf1)
+```
+# save grid image of results
+```{r}
+# arrange grid with population, e coli, WRF, and turbidity predictive models
+bivar_grid <- grid.arrange(m3, m5, t1, wrf1, ncol = 2)
+
+# save and export image
+ggsave(bivar_grid, filename = here("results", "bivar_grid.png"))
+```

### 2021-11-29 code/analysis_code/exploration.Rmd
project pt 5

@@
 library(gridExtra) #
+library(gtsummary)
 

### 2021-11-29 code/analysis_code/land_model.Rmd
project pt 5

@@
 ```{r}
-#based on tidymodels tutorial: case study
+
 #cross validation
@@
 #model specification
-lasso <- linear_reg(penalty = tune()) %>% set_engine("glmnet") %>% set_mode("regression")
+#lasso <- linear_reg(penalty = tune()) %>% set_engine("glmnet") %>% set_mode("regression")
+lasso <- linear_reg() %>%
+  set_mode("regression") %>%           
+  set_engine("glmnet") %>%
+  set_args(penalty = tune(), mixture = 1)
 

### 2021-11-29 code/analysis_code/modeling.Rmd
project pt 5

@@
 library(vip)         # for variable importance plots
-library(gridExtra)
+library(gridExtra) # for savings images in grids
 
@@
 lasso_pred_plot <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
-  labs(title = "LASSO: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
+  labs(title = "LASSO", x = "Outcome (Particles/L)", y = "Prediction (Particles/L)")+ ylim(0, 1000) + xlim(0,1000)
 lasso_pred_plot
+ggsave(lasso_pred_plot, filename = here("results", "lasso_pred.png"))
 
@@
 lasso_resid_plot <- ggplot(lasso_resid, aes(y = .resid, x = .pred)) + geom_point() + 
-  labs(title = "LASSO: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
+  labs(title = "LASSO", x = "Prediction (Particles/L)", y = "Residuals")
 lasso_resid_plot #view plot
@@
 tree_pred_plot <- ggplot(tree_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
-  labs(title = "Decision Tree: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
+  labs(title = "Decision Tree", x = "Outcome (Particles/L)", y = "Prediction (Particles/L)")+ ylim(0, 1000) + xlim(0,1000)
 tree_pred_plot
@@
 tree_resid_plot <- ggplot(tree_resid, aes(y = .resid, x = .pred)) + geom_point() + 
-  labs(title = "Decision Tree: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
+  labs(title = "Decision Tree", x = "Prediction (Particles/L)", y = "Residuals")
 tree_resid_plot #view plot
@@
 However, the decision tree did not perform as well well as the Lasso regularized model, which had an RMSE of about 100.
+Based on the outcome/prediction plot, it appears that the model is making one of two predictions for any given actual outcome.
+This indicates that this is not a good model, as we don't see the expected 45 degree diagonal line where predictions correlate closely with outcomes.
 
@@
 #subset 3/4 of data as training set
-data_split <- initial_split(complete, 
+complete_data_split <- initial_split(complete, 
                             prop = 7/10, 
@@
 #save sets as data frames
-train_data <- training(data_split)
-test_data <- testing(data_split)
+complete_train_data <- training(complete_data_split)
+complete_test_data <- testing(complete_data_split)
 
 #reset recipe with adjusted, complete data
-MP_rec <- recipe(particles_l ~ ., data = train_data) %>% step_dummy(all_nominal())
+MP_rec <- recipe(particles_l ~ ., data = complete_train_data) %>% step_dummy(all_nominal())
 #reset cell_folds
 set.seed(123)
-cell_folds <- vfold_cv(train_data)
+cell_folds <- vfold_cv(complete_train_data)
 ```
@@
 #fit model with finalized WF
-rf_fit <- rf_final_wf %>% fit(train_data)
+rf_fit <- rf_final_wf %>% fit(complete_train_data)
 ```
@@
 rf_resid <- rf_fit %>%
-  augment(train_data) %>% #this will add predictions to our df
+  augment(complete_train_data) %>% #this will add predictions to our df
   select(.pred, particles_l) %>%
@@
 rf_pred_plot <- ggplot(rf_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
-  labs(title = "Random Forest: Pred vs Outcomes", x = "Particles/L Outcome", y = "Particles/L Prediction")
+  labs(title = "Random Forest", x = "Outcome (Particles/L)", y = "Prediction (Particles/L)")+ ylim(0, 1000) + xlim(0,1000)
 rf_pred_plot
@@
 rf_resid_plot <- ggplot(rf_resid, aes(y = .resid, x = .pred)) + geom_point() + 
-  labs(title = "Random Forest: Pred vs Residuals", x = "Particles/L Prediction", y = "Residuals")
+  labs(title = "Random Forest", x = "Prediction (Particles/L)", y = "Residuals")
 rf_resid_plot #view plot
@@
 
+# Model comparison plots
 ```{r}
-model_grid <- grid.arrange(lasso_pred_plot, lasso_resid_plot, rf_pred_plot, rf_resid_plot, tree_pred_plot, tree_resid_plot, ncol = 3)
+model_grid <- grid.arrange(lasso_pred_plot, rf_pred_plot, tree_pred_plot, lasso_resid_plot, rf_resid_plot, tree_resid_plot, ncol = 3)
 
@@
 
-I'll need to go back and edit this image later so that the titles aren't cut off and the axis scales match up properly so the plots look right.
-I might also want to add a 45 degree diagonal line to the outcomes vs predictions plot, and some other features to improve appearance.
+# Model selection : LASSO
+```{r}
+#Based on model performance plots and RMSE, we've selected the LASSO model. Let's evaluate it with the test data.
+
+#fit to test data
+last_lasso_fit <- lasso_final_wf %>% last_fit(data_split)
+last_lasso_fit %>% collect_metrics()
+```
+
+It appears that the lasso model receives an RMSE of ~300 when run with the test data.
+Compared to the RMSE of 100 with the train data, this is not great performance and indicates that the model overfits the train data.
+
+# Variable importance
+```{r}
+library(parsnip)
+var_importance <- last_lasso_fit %>% 
+  purrr::pluck(".workflow", 1) %>%   
+  workflows::extract_fit_parsnip() %>% 
+  vip(num_features = 20)
+
+ggsave(var_importance, filename = here("results", "var_importance.png"))
+```
+
+# Final lasso plots
+```{r}
+#diagnostics
+autoplot(lasso_res)
+#calculate residuals
+last_lasso_resid <- last_lasso_fit %>%
+  augment() %>% #this will add predictions to our df
+  select(.pred, particles_l) %>%
+  mutate(.resid = particles_l - .pred) #manually calculate residuals
+
+#model predictions from tuned model vs actual outcomes
+last_lasso_pred_plot <- ggplot(last_lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
+  labs(title = "LASSO: Pred vs Outcomes", x = "Test Outcome (Particles/L)", y = "Test Prediction (Particles/L)")+ ylim(0, 250) + xlim(0,250)
+last_lasso_pred_plot
+ggsave(last_lasso_pred_plot, filename = here("results", "last_lasso_pred.png"))
+
+#plot residuals vs predictions
+last_lasso_resid_plot <- ggplot(last_lasso_resid, aes(y = .resid, x = .pred)) + geom_point() + 
+  labs(title = "LASSO: Test Pred vs Residuals", x = "Test Prediction (Particles/L)", y = "Residuals")
+last_lasso_resid_plot #view plot
+
+#compare to null model
+metrics_null_train #view null RMSE for train data
+last_lasso_fit %>% collect_metrics() #view RMSE for best lasso model
+
+```

### 2021-11-29 code/analysis_code/subset_modeling.Rmd
code files
project pt 5

@@
 In this portion of the analysis, we will perform modeling on a smaller subset of data in attempt to improve model performance.
+The original aim of this document was to perform feature selection/subsetting for model improvement.
+However, the implementation of the LASSO method eliminated the need for this by essentially automating subset selection.
+The contents of this document are needed for the final analysis/manuscript product, but are kept in the project repository for understanding of the whole analysis process.
 
@@
 #path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
 
@@
 #model specification
-lasso <- linear_reg(penalty = tune()) %>% set_engine("glmnet") %>% set_mode("regression")
-
+#lasso <- linear_reg(penalty = tune()) %>% set_engine("glmnet") %>% set_mode("regression")
+lasso <- linear_reg() %>%
+  set_mode("regression") %>%           
+  set_engine("glmnet") %>%
+  set_args(penalty = tune(), mixture = 1)
 #set workflow
@@
 
+# Model selection : LASSO
 ```{r}
-#check for Poisson distribution
-#https://stackoverflow.com/questions/59809960/how-do-i-know-if-my-data-fit-a-poisson-distribution-using-r
-dispersion_test <- function(x) 
-{
-  res <- 1-2 * abs((1 - pchisq((sum((x - mean(x))^2)/mean(x)), length(x) - 1))-0.5)
+#We've selected the LASSO model. Let's evaluate it with the test data.
 
-  cat("Dispersion test of count data:\n",
-      length(x), " data points.\n",
-      "Mean: ",mean(x),"\n",
-      "Variance: ",var(x),"\n",
-      "Probability of being drawn from Poisson distribution: ", 
-      round(res, 3),"\n", sep = "")
-
-  invisible(res)
-}
-
-set.seed(1)
-x <- data$particles_l
-
-dispersion_test(x)
+#fit to test data
+last_lasso_fit <- lasso_final_wf %>% last_fit(data_split)
+last_lasso_fit %>% collect_metrics()
 ```
 
+It appears that the lasso model receives an RMSE of ~197 when run with the test data.
+Compared to the RMSE of 113 with the train data, this is not great performance and indicates that the model overfits the train data.
+
+# Variable importance
+```{r}
+library(parsnip)
+last_lasso_fit %>% 
+  purrr::pluck(".workflow", 1) %>%   
+  workflows::extract_fit_parsnip() %>% 
+  vip(num_features = 20)
+
 ```
+
+This plot shows us the variable importance rankings for the LASSO model, and we see that the watershed identity of each each sample site has importance for predicting microplastics concentration, while the hypothesized predictors (population, distance from WWTP, turbidity, e.coli.cfu, visual score) do not have importance in this model.
+
+```

### 2021-11-29 code/processing_code/censusprocessing.Rmd
project pt 5

@@
 
-This function produces the zip codes I am looking for, but there are observations with missing variables that didn't return addresses and zip codes.
-I'll have to work on troubleshooting this.
-I may be able to manually input the zip codes, but that is not ideal.
-
-# renaming variable differentiate use in processing
+# renaming variable to differentiate use in processing
 ```{r}

### 2021-11-29 code/processing_code/map_images.Rmd
project pt 5

---
title: "map_images"
output: html_document
---

This document imports the map images available in results/maps and compiles them into one image containing 4 quads. The maps were generated by code on the author's local computer and then uploaded to this project repository due to issues with the mapping packages/files corrupting the GitHub/GitKraken interface.

```{r}
library(gridExtra)
library(png)
library(grid)
library(here)
library(tidyverse)

# map images for each sample date are saved in results/maps as 4 individual images
# here, we will combine the images in to a grid for use in the manuscript

map_fall <- readPNG(here("results", "maps", "fall_map.png"))
map_winter <- readPNG(here("results", "maps", "winter_map.png"))
map_spring <- readPNG(here("results", "maps", "spring_map.png"))
map_summer <- readPNG(here("results", "maps", "summer_map.png"))

map_fall

map_grid <- grid.arrange(rasterGrob(map_fall), rasterGrob(map_winter), rasterGrob(map_spring), rasterGrob(map_summer), ncol = 2)

ggsave(map_grid, filename = here("results", "map_grid.png"))
```

### 2021-11-29 products/manuscript/Manuscript.Rmd
project pt 5

@@
     number_sections: true
-    global_numbering: true
-    fig_caption: yes
-    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
+    fig_caption: true
   bookdown::html_document2: 
@@
 bibliography: library.bib 
-csl: elsevier.csl
+csl: apa.csl
 ---
@@
 Rivers transport both point-source and diffuse pollution to the ocean, contributing an estimated 1.15 to 2.41 million tons of plastic waste annually [@Lebreton2017].
-Factors including anthropogenic impact, pollution sources, and hydrodynamics affect the rate of microplastic accumulation and transport [@Luo2019; CITE].
-Researchers continue to focus on identifying the variables influencing microplastic concentration in present day research (CITE).
+Factors including anthropogenic impact, pollution sources, and hydrodynamics affect the rate of microplastic accumulation and transport [@Luo2019].
+Researchers continue to focus on identifying the variables influencing microplastic concentration in present day research.
 
 ### Ecological Consequences
-Despite their small size, microplastics have significant and quantifiable impact on aquatic ecosystems (CITE).
+Despite their small size, microplastics have significant and quantifiable impact on aquatic ecosystems (@Carbery2018;@Murray2011).
 Since microplastics permeate both marine and freshwater environments, aquatic animals easily mistake the particles for food and ingest them, allowing the pollutant to accumulate in their bodies; studies have shown that organisms ingest microplastics either by accident or by mistaking them for food [@Wright2017;@Carbery2018;@Tanaka2016].
@@
 ### Present Research
-This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, and wastewater treatment plant proximity (@Murphy2016).
+This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, bacteria level, and wastewater treatment plant proximity (@Murphy2016).
+Thus, the key questions that we ask in this study are: what are microplastic levels like in and aroudn Athens, GA?
+Do microplastic levels vary by season?
+Does local population level, local wastewater treatment, land cover, or general water quality influence microplastic levels?
 We hypothesized that microplastics would be present in the Upper Oconee Watershed, that levels would be greater in more highly-trafficked areas with greater anthropogenic impact, and that microplastic pollution would correlate to other indicators of water quality [@McCormick2014a].
-
-## Description of data and data source
-
-The data that I am using for this project is from my personal undergraduate research project on freshwater microplastics in Athens, GA.
-The study is ongoing and was conducted in collaboration with the Upper Oconee Watershed Network (UOWN) via their quarterly sampling events.
-There is one year’s worth of quarterly data, with the following sample dates: November 2020, February 2021, April 2021, and July 2021.
-The data include variables such as sample site, latitude and longitude, watershed location, and duplicate A and B counts of microplastics observed per filter.
-For the most recent sampling date, July 2021, there is additional data where a second counting session was performed, in order to assess the comparability of parallel counts.
-There are about 136 observations of ~10 variables in total in the dataset.
-The independent variables – site location, coordinates, watershed location, etc – were retrieved from UOWN.
-Values for microplastic counts were recorded based on visual identification via a dissecting microscope, performed by myself and by a small team of <10 volunteers.
-
-## Questions/Hypotheses to be addressed
-The questions that I want to answer with this data include the following:
- - Can we characterize microplastics in Athens, GA using a citizen science approach? 
- -	What are the levels of microplastics like in Athens, GA overall? 
- -	How do microplastic levels differ based on location within the region?
-
-Questions that I want to answer that will require gathering more data using coordinates/watershed info include:
- -	Is microplastic concentration impacted by local land use? 
- -	Is microplastic concentration impacted by localized population levels in specific neighborhoods?
- -	Does wastewater treatment plant effluent from nearby facilities increase microplastic concentration at affected sites?
- -	Are microplastic levels correlated to bacteria levels measured at the same sites? 
-These questions also relate to hypothesized predictors of microplastic pollution: population level, land use, wastewater treatment plant effluent, and microbial water quality.
 
@@
 ## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
 
@@
 Samples were pumped through gridded 0.45 micron membranes (Whatman) and immediately covered and stored covered in plastic petri dishes at room temperature for up to 8 weeks.
-Filters were visually inspected using a dissecting microscope with magnification ranging from [INSERT].
+Filters were visually inspected using a stereo microscope with magnification ranging from 10x to 40x.
 Microplastic particles were identified based on the following characteristic features: unnatural homogenous color, unnatural shape, homogenous texture, homogenous width (no tapered end)[@Hidalgo-Ruz2012].
+Visual identification was performed by a team of trained citizen-science volunteers.
 Microplastic concentration (No/L) was calculated by dividing the total microplastic particle count by the sample volume.
 Duplicates (samples A and B for each site) were averaged to obtain a singular concentration value for each sample site.
+
+The combination of all four sample dates resulted in a total of 136 observations of 10 variables.
 
@@
 DI and MilliQ water were filtered in 150 mL volumes, in duplicate, to produce controls to represent potential contamination from filtration processes.
-The average contamination per sample was = INSERT.
 
-## Data import and cleaning
-
-Since I am performing importing and cleaning in a separate document, in this section I will provide an overview of my steps/methods for wrangling the data, without actual code in this section.
-
+## Data Import
 To supplement my analysis, additional data was imported as detailed below.
@@
 ### UOWN Quarterly Data
-Public data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org
+This study relies on bublic data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org.
+UOWN supplied information on samples site names, coordinates, and corresponding watersheds during primary data collection.
+Following microplastic analysis, UOWN data on the following variables for corresponding samples sites and dates was imported to supplement analysis: E.
+coli CFU/100mL,  temperature, turbidity, conductivity, and visual score.
+
 ### Population
-NHGIS [@Mansen]
+Population data was retrieved from the NHGIS database [@Manson].
+Population was considered at the census-tract level, and the corresponding census-tract for each water sample site was determined via conversion of coordinates to census FIPS codes.
+
 ### Land Cover
-NHGIS [@Mansen]
+Land cover data was retrieved from the NHGIS database [@Manson].
+Land cover was considered at the census-tract level, and the corresponding census-tract for each water sample sites was determined via conversion of coordinates to census FIPS codes.
+
 ### Water Reclamation Facilities
@@
 
-## Exploratory analysis
+## Data Cleaning
+The data was cleaned to adjust variable format, calculate microplastic concentration averages, subset sample data and control data, and remove missing observations.
+Observations with missing location (coordinates, watershed) data were removed, as well as observations which were missing count values for microplastic concentration.
+UOWN data was cleaned to remove variables with incomplete observations and variables uninformative for analysis.
+Only variables defined in the study hypothesis were included in the cleaned data.
+Census data was cleaned and wrangled to match long form FIPS codes with short form FIPS codes.
+Land cover data was cleaned to extract FIPS codes, to include only data for the state of Georgia, to remove superfluous variables, to format and rename key variables, and to reassign null values.
+Null land cover proportion values were reassigned to the value zero on the basis that absent values indicated a lack of that particular land cover type in the region (for example, perennial ice/snow was absent in the study region of Georgia).
 
-This section will contain the key products of my exploratory analysis (located in exploration.Rmd) when complete.
-Since I am keeping this document in manuscript style, I will keep processing/exploration/analysis code separate.
+## Analysis
+Microplastic concentration, in terms of particles per liter, was assessed as the key outcome in this study and was treated as a continuous variable.
+The univariate relationship between microplastic concentration and each hypothesized predictor was assessed via linear regression.
+Multiple linear regression was performed to determine a multivariate effect on microplastic concentration.
+The final model was selected based on root mean square error (RMSE) and model evaluation was performed via 5-fold cross-validation (5 times repeated).
+All analysis were completed using R (version 4.0.2) and the tidymodels package was used for cross-validation [@R].
+The code and data to reproduce the results are available on [GitHub](https://github.com/ameliafoley/AmeliaFoley-MADA-project).
 
-Table \@ref(tab:summarytable) shows a table summarizing the data.
-```{r summarytable,  echo=FALSE}
+# Results
 
-#summarytable=readRDS("../../results/summarytable.rds")
-#knitr::kable(summarytable, caption = 'Data summary table.')
+## Univariate Analysis: Microplastic Levels
+The average value of microplastic concentration was 104.29 particles/L and the median was 66.67 particles/L.
+The values ranged from a minimum of 16.67 particles/L to a maximum of 1193.33 particles/L.
+The distribution of microplastic concentration is right-skewed, with a few large outliers.
+Figure \@ref(fig:distribution) shows a histogram of microplastic concentration observations.
+There was some seasonal difference in average microplastic concentration at all sites; Figure \@ref(fig:concentrationbydate) shows a boxplot of concentrations by sample date.
+At each individual sample site, there was also seasonal variation in microplastic levels (Supplement Figure 3).
+Sample sites may be grouped by watershed to show that some watersheds exhibited greater variation in microplastic levels than others (Supplement Figure 4).
+A temporal trend in average watershed microplastic level is reflected over the duration of the study; average microplastic concentration in individual watersheds appears to decrease from November 2020 to July 2021.
+Figure \@ref(fig:watershed-overtime) shows a line graph of the mean watershed microplastic concentrations at each seasonal sampling date.
 
-#I eventually want a nice looking summary statistic table. This is a place holder for now  
-```
-
-### Distribution of Microplastic Concentration
-
-Figure \@ref(fig:distribution) shows a histogram of microplastic concentration observations.
-The minimum concentration is 16.67 particles/L and the maximum is 1193.33 particles/L.
-The mean concentration is 104.39 particles/L, and the median is 66.67 particles/L.
-
-```{r distribution,  fig.cap='Distribution of Microplastic Concentration', echo=FALSE}
+```{r distribution,  fig.cap='\\label{fig::distribution}Distribution of Microplastic Concentration', echo=FALSE}
 knitr::include_graphics("../../results/distribution.png")
 ```
-
-Microplastic concentrations remained in similar ranges throughout the study period. Figure \@ref(fig:concentrationbydate) shows a boxplot of concentrations by sample date.
 
@@
 
-There is some seasonal variation in concentration at each individual site. Figure \@ref(fig:seasonalconc-bysite) shows a plot of concentrations at each site.
-
-```{r seasonalconc-bysite,  fig.cap='Seasonal Variation in Particles/L', echo=FALSE}
-knitr::include_graphics("../../results/seasonalconc_bysite.png")
-```
-
-Figure \@ref(fig:watershed-overtime) shows a line graph of the mean watershed microplastic concentrations at each seasonal sampling date.
-
 ```{r watershed-overtime,  fig.cap='Microplastic Concentration Over Time', echo=FALSE}
@@
 
-### Predictors
-Population, land cover/use, and bacteria levels are hypothesized predictors of microplastic concentration.
-Figure \@ref(fig:logparticlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.
+### Mapping
+Using sample site coordinates and corresponding microplastic concentrations for each seasonal sample date, we generated maps of the study region which demonstrate microplastic concentration at each site.
+Figure \@ref(fig:map-grid) demonstrates the geographic distribution of sample sites and corresponding microplastic levels.
 
-```{r logparticlesvpop,  fig.cap="Particles/L vs Population", echo=FALSE}
-knitr::include_graphics("../../results/logparticlesvpop.png")
+```{r map-grid,  fig.cap='Seasonal Maps of Microplastic Concentration', echo=FALSE}
+knitr::include_graphics("../../results/watershed_overtime.png")
 ```
 
-```{r logmpvcfu,  fig.cap='Log particles/L vs CFU', echo=FALSE}
-knitr::include_graphics("../../results/logmpvcfu.png")
+## Bivariate Analysis
+For the bivariate analysis, we aim to explore the relationship between microplastic concentrations and a suite of hypothesized predictors of microplastic pollution.
+Simple linear regression was used to examine the relationships.
+Figure \@ref(fig:bivar-grid) demonstrates the assessment of linear relationship between microplastic particles/L and the following variables: population, E.
+coli CFU, turbidity, and distance from nearest water reclamation facility.
+The correlation matrix in Figure \@ref(fig:landmatrix) assesses the linear relationship between microplastic particles/L and nearby land cover type.
+There is not a significant relationship between any of these variables and microplastic concentration.
+
+```{r bivar-grid,  fig.cap='Linear modeling of bivariate relationships: microplastic concentration in relation to hypothesized predictors', echo=FALSE}
+knitr::include_graphics("../../results/bivar_grid.png")
 ```
 
-### Mapping
-I created some maps for concentrations at each sample sites throughout the seasons (saved as html files "map_season.html") for now, but I am unsure of how best to display these maps in the manuscript word document output.
-```{r}
-#mapview::knit_print.mapview(here("results", "map_spring.html"))
+```{r landmatrix,  fig.cap='Land cover matrix', echo=FALSE}
+knitr::include_graphics("../../results/landmatrix.png")
 ```
 
-## Full analysis
+## Full Analysis
+Beyond the bivariate analysis, a full model was built to incorporate all predictors in modeling microplastic concentration.
+Three modeling techniques were performed in order to perform a multivariate analysis and subset selection: decision tree, LASSO, and random forest.
+A null model is used to compare model performance and therefore interpret variable importance.
+The null model results in an RMSE of 109.4.
+This metric serves as an informative baseline; any additional model should result in a better (lower RMSE) than the null model.
+The process of building a better model via subset selection and variable regularization reveals important variable relationships and determines variable importance.
 
-Preliminary modeling reveals that there is not a strong relationship between microplastic concentration and population level. Figure \@ref(fig:m-logconcvpop) demonstrates a linear model fit.
+Analysis was performed only on complete cases, which includes a total of 92 observations divided into training (n = 63) and testing (n = 29) sets.
+The initial full model with all predictors returns an RMSE of 101.39.
+A table summarizing the linear model fit is found in the Supplement (Table 1).
+Of the three additional models built to predict microplastic concentration, the LASSO model was selected as the best performing (Figure \@ref(fig:model-grid)).
+With the training data, the RMSE for the LASSO model is 100.5, while the RMSE values for the decision tree and random forest models are 102.2 and 108.8, respectively.
 
-```{r m-logconcvpop-no-outlier,  fig.cap='Concentration vs Population Linear Model', echo=FALSE}
-knitr::include_graphics("../../results/m_logconcvpop.png")
-```
+The predicted outcomes are plotted against actual microplastic concentration in Figure \@ref(fig:last-lasso-pred) and variable importance in the model is demonstrated in Supplement Figure 14.
+None of the hypothesized predictors appear as important variables in the model.
+However, the RMSE for the LASSO model on the test data is 196.7; this performance is significantly worse than both the initial full linear model prior to feature selection and regularization and the original null model.
 
-Figure \@ref(fig:m-logconcvcfu) shows a linear model of microplastic concentration vs CFU (both variables log-transformed).
-
-```{r m-logconcvcfu,  fig.cap='Concentration vs CFU Linear Model', echo=FALSE}
-knitr::include_graphics("../../results/m_logconcvcfu.png")
-```
-
-Figure \@ref(fig:concvturbidity) demonstrates a linear model of particles/L vs turbidity.
-
-```{r m_concvturbidity,  fig.cap='Concentration vs Turbidity Linear Model', echo=FALSE}
-knitr::include_graphics("../../results/m_concvturbidity.png")
-```
-
-Table \@ref(tab:lmfittable) shows a table summarizing a linear model fit predicting particles/L with 6 predictors.
-
-```{r lmfittable,  echo=FALSE}
-lm_fit_table=readRDS("../../results/lm_fit_table.rds")
-knitr::kable(lm_fit_table, caption = 'Linear model fit table.')
-```
-
-Beyond the basic linear model, we have applied additional methods to improve model performance, including LASSO regularization and building decision trees and random forests for model comparison.
-The predictions, outcomes, and residuals resulting from each type of plot are demonstrated in Figure \@ref(fig:model-grid)
-
-```{r model-grid, fig.cap='Model Quality', echo=FALSE}
+```{r model-grid, fig.cap='Comparison of Model Performance on Training Data', echo=FALSE}
 knitr::include_graphics("../../results/model_grid.png")
@@
 
-Based on the results of the three different models, the LASSO model is the best option for this dataset, though the minimal difference in RMSE when compared to the null model suggests that even though LASSO is the better model method compared to others, it still does not produce a great model for predicting microplastic concentration.
-
-# Results
+```{r last-lasso-pred,  fig.cap='Microplastic concentration predicted by full LASSO linear regression model', echo=FALSE}
+knitr::include_graphics("../../results/last_lasso_pred.png")
+```
 
@@
 At the start of this project, there was little to no existing research on microplastics in Athens.
-Therefore, the data resulting from this study establish the presence of microplastics in the area [mention support of other studies showing MP in urban vs rural areas].
+Therefore, the data resulting from this study establish the presence of microplastics in the area.
+
+## Summary and Interpretation
+In this analysis, we examined the distribution of freshwater microplastics in Athens, GA, as well as the relationship between the main outcome, microplastic concentration, and a few key hypothesized predictors: population level, microbial water quality, land cover, and proximity to wastewater treatment.
+Microplastic levels were monitored over a period of one year, with water samples being collected and processed quarterly in November 2020, February 2021, April 2021, and July 2021.
+Initial data collection provided sample location data as well as microplastic concentration data.
+Additional data was later imported, including microbial water quality, local population, land cover, and distance from the nearest water treatment facilities.
 
@@
 There is growing evidence that microplastic pollution in permeates ecosystems in unexpected, supposedly pristine regions across the globe; one study found between 22 and 117 particles per liter in the Antarctic, demonstrating that microplastic particles are being found in places you would never expect.
-## Variation of Microplastic Concentration
+
+Exploratory data analysis suggested that there were not strong relationships between microplastic concentration and any of the hypothesized predictors.
+Linear regression analysis confirmed that there were no significant correlations between microplastic concentration and predictors, contrary to our hypothesis based on existing research.
+
+Takeaways include that no anthropogenic indicator can be established as an predictor of microplastic concentration at this time.
+However, the presence of freshwater microplastics in the Upper Oconee Watershed is confirmed via this analysis, and the quantification and distribution of microplastic pollution may act as a baseline for future studies and analyses.
+
+## Strengths and Limitations
 Limitations to the scope of this project include that it was designed as a student-led pilot with minimal funding.
@@
 
+This analysis focused on identifying linear relationships between the outcome and predictors.
+However, no strong linear relationships were identified.
+While there may be no relation between the outcome and the predictors, linear models may be insufficient for capturing these relationships.
+
+A strength of the study design includes that it incorporates citizen science methods of sample collection and analysis and implements cost-effective procedures.
+The design of this exploratory analysis makes the study accessible and feasible to reproduce in the field and laboratory.
+To the knowledge of the authors, this is the first study to examine microplastic pollution in Athens, GA.
+Therefore, this study accomplishes exploratory aims and meets the growing demand for freshwater microplastics research and monitoring.
+## Conclusions
+The present research establishes the existence of microplastics in Athens, GA's freshwater bodies, though there is no clear relationship between microplastic concentration and the hypothesized predictive factors, according to the data recorded and analyzed in this study.
+Continued research is required in order to identify sources of microplastic pollution so that pollution levels may be mitigated in the future.
+Areas of focus for the future include furthering temporal, routine data collection, refining quantification methods, and increasing funding for microplastic research technologies.
+
 This year-long pilot study is followed by a continued partnership with the Upper Oconee Watershed Network, who will continue sampling for microplastics quarterly through citizen-science volunteer efforts.
-Areas of focus for the future include furthering temporal data collection, refining quantification methods, observing the relationship between microplastic levels and other water quality indicators, and exploring other hypothesized water quality predictors.
 While limited in scope, this study provides valuable initial data on the presence of microplastics in the region and establishes a basis for continued and refined research on the growing subject of freshwater microplastics.
 
-## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
-
-## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
-
-## Conclusions
-_What are the main take-home messages?_
-
 # References

### 2021-11-29 products/manuscript/Supplement.Rmd
project pt 5

@@
     fig_caption: yes
-    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
   bookdown::html_document2: 
@@
 
-## Exploratory analysis
+# Documentation for Supplementary Files
+Code and data files required to reproduce this analysis are available on [GitHub](https://github.com/ameliafoley/AmeliaFoley-MADA-project)
 
-This section will contain the key products of my exploratory analysis (located in exploration.Rmd) when complete.
-Since I am keeping this document in manuscript style, I will keep processing/exploration/analysis code separate.
+* "**AmeliaFoley-MADA-project.Rproj**" Establishes relative file paths
+* "**README.md**" Provides brief order of scripts for reproducing; summarizes project parts
+* "**code**" folder contains all code for processing and analysis
+  + "**analysis_code**"
+    + analysis.Rmd performs linear regression on single predictors and full predictor sets
+    + exploration.Rmd contains the exploratory data analysis
+    + land_model.Rmd contains statistical models on land cover data
+    + modeling.Rmd is the main analysis file containing final models (LASSO)
+    + subset_modeling.Rmd is an additional analysis file performing models on subsets of data; it is not used for the manuscript
+  + "**processing_code**"
+    + censusland_code.Rmd combines land cover data with microplastic data
+    + censusprocessing.Rmd imports census tract population data
+    + map_images.Rmd combines individual map images to produce a grid
+    + mpprocessing.R processes the original microplastic data file and saves cleaned data
+    + uownprocessing.Rmd imports additional water quality data from UOWN.org
+    + wrfprocessing.Rmd imports distance for nearest water reclamation facility
+  + "**readme.md**" provides an order and brief description of dependencies/outputs for processing code files
+* "**data**" folder contains all raw and processed data
+* "**products**" folder contains files to produce the Manuscript and Supplement
+  + "**manuscript**" contains files for producing the manuscript
+    + apa.csl optional format for citations
+    + elsevier.csl optional format for citations
+    + library.bib contains references
+    + Manuscript.docx result of knitting the markdown
+    + Manuscript.Rmd creates basic manuscript word document
+    + style_template.docx placeholder to add special formatting to manuscript
+    + Supplement.Rmd creates basic supplemental material word document
+    + Supplement.docx result of knitting markdown file
+* "**results**" folder contains all plots, tables, figures produced in data analysis
 
-Table \@ref(tab:summarytable) shows a table summarizing the data.
-```{r summarytable,  echo=FALSE}
+# Reproducing results
+Reproducing this project requires R, RStudio, and Microsoft Word. Files should be run in the following order.
 
-#summarytable=readRDS("../../results/summarytable.rds")
-#knitr::kable(summarytable, caption = 'Data summary table.')
+In the code > processing_code folder:
 
-#I eventually want a nice looking summary statistic table. This is a place holder for now  
-```
+1. mpprocessing.R
+2. uownprocessing.Rmd
+3. censusprocessing.Rmd
+4. censusland_code.Rmd
+5. wrfprocessing.Rmd
+6. exploration.Rmd
+
+In the code > analysis_code folder:
+
+7. analysis.Rmd
+8. modeling.Rmd
+9. land_model.Rmd
+10. map_images.Rmd
+
+In the products folder
+
+11. Manuscript.Rmd
+12. Supplement.Rmd
+
+# Supplementary results
 
@@
 Population, land cover/use, and bacteria levels are hypothesized predictors of microplastic concentration.
-Figure \@ref(fig:particlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.
+Figure \@ref(fig:particlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/100mL), respectively.
 
@@
 
-Figure \@ref(fig:concvturbidity) demonstrates a linear model of particles/L vs turbidity.
+Figure \@ref(fig:m-concvturbidity) demonstrates a linear model of particles/L vs turbidity.
 
-```{r m_concvturbidity,  fig.cap='Concentration vs Turbidity Linear Model', echo=FALSE}
+```{r m-concvturbidity,  fig.cap='Concentration vs Turbidity Linear Model', echo=FALSE}
 knitr::include_graphics("../../results/m_concvturbidity.png")
@@
 ```
+
+Beyond the basic linear model, we have applied additional methods to improve model performance, including LASSO regularization and building decision trees and random forests for model comparison.
+The predictions, outcomes, and residuals resulting from each type of plot are demonstrated in Figure \@ref(fig:model-grid)
+
+```{r model-grid, fig.cap='Model Quality', echo=FALSE}
+knitr::include_graphics("../../results/model_grid.png")
+```
+
+Based on the results of the three different models, the LASSO model is the best option for this dataset, though the minimal difference in RMSE when compared to the null model suggests that even though LASSO is the better model method compared to others, it still does not produce a great model for predicting microplastic concentration.
+
+Figure \@ref(fig:var-importance) demonstrates variable importance in the final selected LASSO model. None of the hypothesized predictors appear as important variables in this model.
+
+```{r var-importance, fig.cap='Variable importance', echo=FALSE}
+knitr::include_graphics("../../results/var_importance.png")
+```

### 2021-11-29 code/processing_code/map_images.Rmd
improve manuscript map

@@
 ggsave(map_grid, filename = here("results", "map_grid.png"))
+
+map_grid1 <- grid.arrange(rasterGrob(map_fall), rasterGrob(map_winter), ncol = 2)
+ggsave(map_grid1, filename = here("results", "map_grid1.png"))
+map_grid2 <- grid.arrange(rasterGrob(map_spring), rasterGrob(map_summer), ncol = 2)
+ggsave(map_grid2, filename = here("results", "map_grid2.png"))
 ```

### 2021-11-29 products/manuscript/Manuscript.Rmd
improve manuscript map

@@
 
-```{r map-grid,  fig.cap='Seasonal Maps of Microplastic Concentration', echo=FALSE}
-knitr::include_graphics("../../results/watershed_overtime.png")
+```{r map-grid,  fig.cap='Seasonal Maps of Microplastic Concentration', echo=FALSE, fig.height=3}
+knitr::include_graphics("../../results/map_grid.png")
 ```

### 2021-12-04 reviews/Project Review_Alexandra Gil.Rmd
Project Review A. Gil

# Overview

Title of project:  "Plastic in the Urban Freshwater Environment"

Name of project author(s): Amelia M. Foley

Name of project reviewer: Alexandra Gil

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/knits into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments
Solid background, well conceptualization.The summary and background provided a complete context about the current issue with microplastic pollution. Solid background, well conceptualization. The way that you describe analysis is really detailled.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The analysis includes multiple questions, those are clear and the data set is clearly oriented to solve the question. Good Job!

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data is well described. Every data source was explained clearly. A codebook was not included but the information provided for every data set makes it clear.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Exploration.Rmd shows a good exploration of your data. I really like the way that you go analyzed every plot and explain the reasons that you toke in consideration for performed the exploration.
I really like the way that you present the correlation, it was the easiest way to clearly see the correlation between the amount of variables that you have. Other plots were very clear too.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The analysis was very extructured. The model evaluation with turbidity - linear regression shows better performance base on the RMSE

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The analysis presents a good amount of plots and tables for present the outcomes.
The visualization was good, and consistend according with the analysis.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion/conclusion is base on the outcomes, showing that the data set an the way that the analysis was performed answer the questions of the project.

### Summary assessment

* strong, complete and clear discussion

## Further comments

_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Overall the project is well structured, there is not unused files. However, it would be beneficial includes the best way to start running the code.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The documentation around every script was clear and very informative. I like the way you describe the outcomes and the way that you will try to solve issues.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

- Since you included so many processing codes, could be a good idea includes the best way to start the running the code. This could be described in README portion.

- Unfortunately for my, I couldn't reproduced the LASSO modeling. My R aborted multiple time. But this is not anything wrong with your code...After multiple intents even in homeworks I conclude that is something associate with my laptop.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The way that you structure your project was very detailed. For modeling portion the analysis includes LASSO, Decision Tree, and Random Forest. The plot model_grid.png was a amazing idea! Good Job :)

### Summary assessment

* strong level of thorougness

## Further comments

Good Job! in my opinion this is a good topic well analyzed, because so many areas were analyzed.

### 2021-12-09 code/analysis_code/exploration.Rmd
Clean up

@@
 Quantities of main interest:
+
  - particles per liter of microplastics
@@
 
-Nice!
 Going forward, I should think about whether to keep or remove the watersheds that were only sampled in the last (July) sampling.
-I also see that some watershed are missing intermediate data points and are not connected.
-Would we want to connect and skip over a season, or leave them unconnected (looking at Oconee River)
+Note: I also see that some watershed are missing intermediate data points and are not connected.
+Would we want to connect and skip over a season, or leave them unconnected (looking at Oconee River)?

### 2021-12-09 code/analysis_code/subset_modeling.Rmd
Clean up

@@
 However, the implementation of the LASSO method eliminated the need for this by essentially automating subset selection.
-The contents of this document are needed for the final analysis/manuscript product, but are kept in the project repository for understanding of the whole analysis process.
+The contents of this document are *not* needed for the final analysis/manuscript product, but are kept in the project repository for understanding of the whole analysis process.
 
@@
 Include:
+
 1. Model specification
@@
 
-next I want to see what predictors are used in the best lasso model. also need to compare to null model
+Next, I want to see what predictors are used in the best lasso model. also need to compare to null model
 
@@
 The results we get from the initial machine learning modeling and from the modeling of this subset data are not great. Let's think of some ways that we can go about improving our model.
+
 - An option is variable/predictor removal. We tried that with the subset data. We could attempt further subsets going forward. 
 - Variable transformation. Some microplastic studies use log transformation on the concentration values to get a normal distribution. 
-- Alternatives: account for different distribution.
-The distribution of microplastic concentration may be considered more Poisson distributed than normally distributed.
-First, we can test to see whether it falls in either of these distribution categories, and based on the results we can apply a GLM with Poisson distribution.
-(Tranformation could make intrepretability more difficult - may want to avoid)
-- however, regularization may accomplish our goals more efficiently - we are advised that subset selection is not always a great idea and doesn't neccesarily do the best job of addressing overfitting
-- could try ridge regression for alternate form of regularization 
-- could try decision tree (though we learned these usually don't perform as well as LASSO)
+- However, regularization may accomplish our goals more efficiently - we are advised that subset selection is not always a great idea and doesn't neccesarily do the best job of addressing overfitting
+- Could try ridge regression for alternate form of regularization 
+- Could try decision tree (though we learned these usually don't perform as well as LASSO)
 

### 2021-12-09 code/processing_code/ARCHIVE/uownprocessing.Rmd
Clean up

--- 2021-10-07 code/processing_code/uownprocessing.Rmd
@@
 glimpse(uowndata)
+```
+
+# Merging mass UOWN dataset and July data, which was received later (10/27/21)
+```{r}
+july <- read.csv(here("data", "raw_data", "uown_summer2021_qtrly.csv"))
+glimpse(july)
+#The july data also contains 17 columns, just like the mass dataset. This indicates that these datasets should be easy to merge. 
+#adjusting class of some variables so that they are compatible between these two datasets before joining
+july <- july %>% mutate(ID = as.character(ID), 
+               pH = as.numeric(pH)) 
+#adjust class of e.coli.cfu from character to numeric to join datasets (and to work with later in analysis)               
+uowndata <- uowndata %>% mutate(e.coli.cfu = as.numeric(e.coli.cfu))
+
+uowndata <- full_join(uowndata, july, all = TRUE)
 ```

### 2021-12-09 code/processing_code/processing_code.Rmd
Clean up

---
title: "processing"
output: html_document
---
# MICROPLASTICS PROCESSING #

```{r}
###############################
# processing script
#
#this script loads the raw data, processes and cleans it 
#and saves it as Rds file in the processed_data folder

#load  packages
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)

#path to data
data_location1 <- here::here("data","raw_data","mpdata_ALL.xlsx")
data_location2 <- here::here("data","raw_data","uown 8.26.xlsx")

#load microplastics data. 
rawdata1 <- read_excel(data_location1)
rawdata2 <- read_excel(data_location2)

#take a look at the data
glimpse(rawdata1)
glimpse(rawdata2)

#view the dataset on screen
print(rawdata1)
print(rawdata2)

#we have 11 variables in the first dataset, and 18 in the second. These variables
#should correspond so that we can combine the datasets

rawdata <- full_join(rawdata1, rawdata2)

# now, we have a combined dataset to work with

glimpse(rawdata)

# the notes column won't be neccessary for data analysis. Columns 16, 17, and 18 
# are also unnecessary. Let's remove those.

processeddata <- rawdata %>% select(-"notes", -(17:19))

# assessing missing values
is.na(processeddata) %>% summary()

# Since there are just a few observations that are missing location (watershed, lat, and long)
# and/or count data, and there is not a great way to impute them. Since I am working
# with data that I collected myself, I know that those observations simply didn't
# get counted, got lost, or site names were misread/input incorrectly. Let's clean 
# up the data set.

processeddata <-subset(processeddata, site!="MIDO 504" & site!="MIDO 503")

# I've removed the rows by name instead of filtering out rows with missing watershed information.
# This is because I want to retain the "control" rows for now, which originated 
# in the lab and don't have a corresponding sample location

# Another thing I can do is tidy the data by filling in all "id" information (all missing ids for previous
# samplings would be me (a)).

processeddata$id[is.na(processeddata$id)]<-"a"

# Next, let's remove observations where the samples were originally recorded in the
# data sheet, but apparently never got counted during sample processing. Again, 
# we just want to remove sample type observations and not control type observations.

sampledata <- processeddata %>% filter(obs_type == "sample") %>% subset(a_count!="NA" & b_count!="NA")
controldata <- processeddata %>% filter(obs_type == "control")

# There are a lot of missing values for the c_count and d_count. These 
# variables resulted from the fourth quarterly sampling in the dataset, when
# volunteers were enlisted to participate in a microplastics visual identificaiton
# training session, and volunteers then completed counts. IDs (a, b, c, etc) were 
# recorded for volunteers. Each biological samples usually includes two technical 
# replicates - the a_count and b_count. For the volunteer session, each set of 
# replicates was counted twice (once by two different volunteers) in order to 
# assess the efficacy of the training program/reproducability of counts.

# Ideas to address this: 
# 1. Subset the data into one group of all samples with just a_count
# and b_count and another group with counts a, b, c, and d for the July sampling

# volunteer contains original particle_l values
volunteer <- sampledata %>% filter(date == "july_21") #subsetting data with excess counts

# I can look at this subset later to examine differences in counts. For the purposes
# of the overall data set, I will average the four counts for the july samples
# in order to calculate the particles per liter value. Since particles_l is already calculated
# for a_count and b_count, and c_count and d_count, respectively, I'll average that value.

# julydata contains averaged particle_l values to be added to the sampledata dataset
julydata <- sampledata %>% filter(date == "july_21") %>% mutate(
   particles_l = ((particles_l+rep_particles_l)/2), 
   rep_particles_l = NULL)

otherdata <- sampledata %>% filter(!date== "july_21") #to rejoin later

#merging mutated data
cleansampledata <- full_join(otherdata, julydata) %>% select(-"id",
                                                         -"id2",
                                                         -"rep_particles_l",
                                                         -"a_count",
                                                         -"b_count", 
                                                         -"c_count",
                                                         -"d_count",
                                                         -"volume_ml",
                                                         -"obs_type")

# save data as RDS

# location to save file
save_data_location <- here::here("data","processed_data","cleansampledata.rds") 
save_data_location1 <- here::here("data","processed_data","volunteer.rds") #saving july volunteer data to look at later
save_data_location2 <- here::here("data","processed_data","controldata.rds")

saveRDS(cleansampledata, file = save_data_location)
saveRDS(volunteer, file = save_data_location1)
saveRDS(controldata, file = save_data_location2)

# Here, I'll also load some data we'll want to look at later. The csv file being 
# imported contains the coordinates for Athens' three water reclamation facilities (WRF)
# Since WRF efflunet is a hypothesized predictor of microplastic levels, we'll examine
# this later in the exploratory analysis.

# load Water Reclamation Facility location data

#path to data
data_location3 <- here::here("data","raw_data","wrflocation.csv")

#load WRF data
wrflocation <- read.csv(data_location3)

#save data as RDS
save_data_location3 <- here::here("data", "processed_data", "wrflocation.rds")

saveRDS(wrflocation, file = save_data_location3)

```

# UOWN PROCESSING #
This section of code loads the raw UOWN data, processes and cleans it 
and saves it as Rds file in the processed_data folder. While the microplastics data is the focus of my project, I am interested in comapring UOWN's data on bacteria levels with the recorded microplastic levels at samples sites throughout the study timeframe.

# load packages
```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
```

# path to data & load data
```{r}
data_location1 <- here::here("data","raw_data","UOWN_spring2021.csv")

uowndata <- read.csv(data_location1)
```

# Take a look at data
```{r}
glimpse(uowndata)
```

# Merging mass UOWN dataset and July data, which was received later (10/27/21)
```{r}
july <- read.csv(here("data", "raw_data", "uown_summer2021_qtrly.csv"))
glimpse(july)
#The july data also contains 17 columns, just like the mass dataset. This indicates that these datasets should be easy to merge. 
#adjusting class of some variables so that they are compatible between these two datasets before joining
july <- july %>% mutate(ID = as.character(ID), 
               pH = as.numeric(pH)) 
#adjust class of e.coli.cfu from character to numeric to join datasets (and to work with later in analysis)               
uowndata <- uowndata %>% mutate(e.coli.cfu = as.numeric(e.coli.cfu))

uowndata <- full_join(uowndata, july, all = TRUE)
```

# Subsetting UOWN data
```{r}
uowndata <- uowndata %>% filter((year == 2020 & month == 11) | (year == 2021 & quarter == 1) | (year == 2021 & quarter == 2 )| (year == 2021 & quarter == 3))
```

Now we have the UOWN data from just the samples dates that we also have the microplastic data for. Right now, only Nov 2020, Feb 2021, and April 2021 data are available from the UOWN website, but July 2021 data should be available soon and I will add it in whenever it is posted. The code chunk above should already include the code to load the July data when the updated file is available.

# Cleaning up the data
```{r}
cleanuowndata <- uowndata %>% select(-"ecoli.method.known", 
                                     -"biological_score",
                                     -"po4.mgL", 
                                     -"conductivity.uscm",
                                     -"no3.mgL",
                                     -"pH") %>%
  rename(site=WSID)

```

# Save UOWN data alone
```{r}
# save data as RDS

# location to save file
save_data_location <- here::here("data","processed_data","cleanuowndata.rds")

saveRDS(cleanuowndata, file = save_data_location)
```

# Combine UOWN data with microplastic sample data
```{r}
# load MP data
data_location <- here::here("data","processed_data","cleansampledata.rds")
cleansampledata <- readRDS(data_location)

# edit site column to match UOWN data
# remove white space in site name 
cleansampledata <- cleansampledata %>% mutate(site = 
                                     gsub(" ","",cleansampledata$site))

# join data sets by site name (cleansampledata) / WSID (cleanuowndata)
joindata <- left_join(cleansampledata, cleanuowndata, by = "site")

# we joined the data by site, but we also need to join it by date. we'll have to clean up some of the columns in order to do that

# edit date format by creating new columns for month and year
cleansampledata <- cleansampledata %>% mutate(year = ifelse(date == "nov_20", "2020", "2021")) %>% mutate(month = ifelse(date == "nov_20", "11", 
                                                                                                              ifelse(date == "feb_21", "2", 
                                                                                                                     ifelse(date == "apr_21", "4", 
                                                                                                                            ifelse(date == "july_21", "7", "no")))))
cleansampledata$month <- as.numeric(cleansampledata$month)
cleansampledata$year <- as.numeric(cleansampledata$year)
# try joining again
joindata <- left_join(cleansampledata, cleanuowndata, by = c("site", "month", "year"))

# clean up joined dataset
joindata <- joindata %>% select(-"day",
                                -"WS", 
                                -"ID")
# save combined data as RDS
save_data_location <- here::here("data", "processed_data", "combodata.rds")

saveRDS(joindata, file = save_data_location)

```

# CENSUS PROCESSING #

At this point, I have written a few processing files loading the microplastics data, land cover/use data, and UOWN data for this project. I'd also like to look at small area population. I think I can achieve this by looking at population by zip code in Athens Clarke County, which is available from the US Census. My goal is to convert the latitude and longitude coordinates for each sample site into a zip code, and then examine whether microplastic levels vary based on local population size/population density.

The data will come from the American Community Survey 5-year estimate 2019, and I will plan to use API. To convert sample coordinates to comparable zip codes, I can use the `tidygeocoder` package.

Citation for zipcode population data (saving for later): 
United States Census Bureau. B01001 SEX BY AGE, 2019 American Community Survey 5-Year Estimates. U.S. Census Bureau, American Community Survey Office. Web. 10 December 2020. http://www.census.gov/.

Downloaded at: https://www.georgia-demographics.com/zip_codes_by_population

```{r}
#load  packages
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(revgeo) #for reverse geocoding coordinates to zipcode

#load data. 
combodata <- readRDS(here("data", "processed_data", "combodata.rds"))
```

# renaming variable to differentiate use in processing
```{r}
popdata <- combodata
#save data as RDS
saveRDS(popdata, file = here("data", "processed_data", "popdata.rds"))

```

Update 10/25/21

I am realizing that we can get more localized population levels by using census tracts, which are smaller levels of census geography. Let's try to use FIPS codes (which include tract info) to obtain population data that might be more informative. I've removed the previous code that converted coordinates to zip code and got population based on zip code. This code is now located in `ARCHIVE.Rmd`

# Loading data with FIPS variable, converted from coordinates

FIPS codes were obtained through a census API run on the local computer, then saved locally and copied into the project directory under `raw_data` for this analysis. This ensures that the code will be reproducible and consistent even if the API changes in the future. 
```{r}
fips <- readRDS(here("data", "raw_data", "fips_copy.RDS"))
#saveRDS(fips, here("data", "processed_data", "fips.RDS"))
```

Here, after attemping to use `get_decennial()` we receive an error message suggesting to use NHGIS data since the most recent small census geography data is not available through `get_decennial()`.

# Using NHGIS data for more recent block-level population counts
```{r}
#Downloaded data from http://NHGIS.org and placed in "raw_data" for this R project. 
library(ipumsr)

#read in census tract level population data
tract <- read.csv(here("data","raw_data","nhgis0001_csv", "nhgis0001_ds244_20195_tract.csv"))

#clean up dataset to remove unneeded columns/variables
tract <- tract %>% filter(STATE == "Georgia") %>% select(1:3, 6:9, 12, 40, 43:46)

#Now, we have the FIPS codes for each sample site (down to the tract/block level), and we have population counts by FIPS code.

#First, we need to remove the last four digits of the FIPS code in the fips dataset so that it will correspond with the #FIPS code in the tract dataset. We'll also need to remove the G from the beginning of the tract FIPS codes.

#combine individual codes for state, county, tract to get long form FIPS code
a <- tract %>% mutate(FIPS = paste(STATEA, COUNTYA, sep = "")) %>%
  mutate(FIPS = paste(FIPS, TRACTA, sep = "0"))

#keep only the variables we need, FIPS code and population size (ALUBE001)
a <- a %>% select("FIPS", "ALUBE001")
#remove last four digits from fips codes associates with samples, so that they coordinate with population census tract FIPS codes
b <- fips %>% mutate(FIPS = (substr(FIPS,1,nchar(FIPS)-4)))
#combine datasets by FIPS code, so that population size is added to our main dataset based on corresponding FIPS code
blockpop <- left_join(b, a, by = "FIPS")

#We still have some missing values, but when I search for the FIPS code that are missing population levels, I am able to find them in the tract dataset. It looks like the FIPS code/GEOID are sometimes the same but sometimes differ by one digit (usually a zero inserted in the middle of the string somewhere). Let's adjust the GEOID variable to make it comparable to the FIPS codes we are still missing population data for, and then we'll join the datasets again to try to fill in missing values.

#let's try GEOID minus first 7 digits
tract$GEOID <- gsub("^.{0,7}", "", tract$GEOID) 
#selecting only the variables we want to add in our final combination dataset
tractpop <- tract %>% select("GEOID", "ALUBE001")
#joining based on FIPS code to fill in missing population data
c <- left_join(blockpop, tractpop, by = c("FIPS" = "GEOID"))

#This last attempt was successful and we now have a column that gives us the census tract level population count, with no missing values. We no longer need the first column, or the zip code and zip code population data that we looked at initially. Let's clean up this dataset for further exploration

#Remove first ALUBE001 column and zip, population columns
tractpopdata <- c %>% select(-"ALUBE001.x")
tractpopdata <- rename(tractpopdata, "population" = "ALUBE001.y")

#save as RDS
saveRDS(tractpopdata, file = here("data", "processed_data", "tractpopdata.rds"))

```

# LAND PROCESSING #
This portion of code will load the land cover data and associate it with each sample site.

##Load Dependencies
The following code loads the package dependencies for our analysis:
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
#install.packages("tidycensus", repos = "http://cran.us.r-project.org")
#install.packages("censusapi", repos = "http://cran.us.r-project.org")
#install.packages("censusGeography", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/chron_2.3-55.tgz")
#install.packages("rerddap", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/rerddap_0.6.5.tgz")

library(tidyverse)
library(dplyr) #data wrangling
library(tidycensus) #access census data
library(censusapi)
library(rerddap)
library(RCurl)

```

##Load Census API (load only once)
```{r}
#tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE) 
```
^This is not a code chunk you will need in every notebook. As long as "install = TRUE", you will only have to do this once. 
#restart R session after installing the first time

##Obtain FIPS codes for sampling sites
```{r}
#create function
latlong2fips <- function(latitude, longitude) {
  url <- "https://geo.fcc.gov/api/census/block/find?format=json&latitude=%f&longitude=%f"
  url <- sprintf(url, latitude, longitude)
  json <- RCurl::getURL(url)
  json <- RJSONIO::fromJSON(json)
  as.character(json$County['FIPS'])
}
#vectorize so function will take more than one argument
latlong2fips_vec <- Vectorize(latlong2fips, vectorize.args = "latitude", "longitude")
```

#accessing NHGIS land use/land cover data from downloaded file
source:https://www.nhgis.org/user-resources/environmental-summaries#download
```{r}
#path to data
data_location1 <- here::here("data","raw_data","landcover2015.csv")

#load data. 
landcover <- read.csv(data_location1)

class(landcover$GISJOIN)
#class = character, so I can use strsplit(). Data source uses extra letters and zeroes in their GIS code. I need to extract the original FIPS code to match my data

#extract state fip from GISJOIN
landcover = landcover %>% mutate(state_fip = substring(landcover$GISJOIN, 2,3)) %>% relocate(state_fip)

#extract county fip from GISJOIN
landcover = landcover %>% mutate(county_fip = substring(landcover$GISJOIN, 5, 7)) %>% relocate(county_fip)

#combine to state, county fip for whole fips codes
landcover$fips = paste0(landcover$state_fip, landcover$county_fip)

landcover <- landcover %>%
  select(fips, everything()) # moving fips column to far left of data set

#clean dataset by selecting only data for the state of Georgia
landcover <- landcover %>%
  filter(state_fip == "13")

#get rid of the 4th and the 8th zero in GISJOIN to make compatible with FIPS codes for samples + population count
landcover = landcover %>% mutate(FIPS = paste(state_fip, county_fip, substring(landcover$GISJOIN, 9, 14), sep = "")) %>% relocate(FIPS)

#filter to clean data set
landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID, fips)
#remove empty variables and rename variables according to key described in readme file for raw NHGIS landcover data
landcoverclean <- landcoverclean %>% select(-(4:67)) #remove variable columns from years 2001 and 2006, keeping only 2011 data 
landcoverclean <- landcoverclean %>% rename(open_water = AREA11_2011,
                                            perennial_ice = AREA12_2011, 
                                            developed_open = AREA21_2011, 
                                            developed_low = AREA22_2011, 
                                            developed_medium = AREA23_2011, 
                                            developed_high = AREA24_2011, 
                                            barren_land = AREA31_2011, 
                                            deciduous_forest = AREA41_2011, 
                                            evergreen_forest = AREA42_2011, 
                                            mixed_forest = AREA43_2011,
                                            shrub = AREA52_2011, 
                                            grassland = AREA71_2011, 
                                            pasture = AREA81_2011, 
                                            crops = AREA82_2011, 
                                            woody_wetlands = AREA90_2011, 
                                            herb_wetlands = AREA95_2011) 
landcoverclean <- landcoverclean %>% select(-(20:35), 
                                            -"perennial_ice") #removing empty perennial ice column, removing proporation variables (keeping area values rather than proportions. we have the total area and can recalculate proportion later if needed)

#Based on the NAs in the perennial_ice columns and knowing the context/region of the data, I am thinking that the other NA values for certain land cover types simply indicate that there is no land cover of that type in the area - therefore, these values are essentially zero. Let's replace NA values with zeroes so that we are working with complete (not missing) data.

landcoverclean[is.na(landcoverclean)] <- 0 #replace NA with 0

#Now we have landcover data in a form that is compatible with our sample site data (using small geography census info). Let's join datasets to create a dataset with sample site info and land cover info
tractpopdata <- readRDS(here("data", "processed_data", "tractpopdata.rds"))
athensland <- left_join(tractpopdata, landcoverclean, by = "FIPS")

#cleanup resultant dataset
athensland <- athensland %>% select(-"fips", 
                              -"FIPS", 
                              -"AREA", 
                              -"quarter", 
                              -"month",
                              -"year")

#save as RDS
saveRDS(athensland, file = here("data", "processed_data", "athensland.rds"))
```

# WRF PROCESSING #
This portion of code will use water reclamation facility location data to calculate the distance between each sample site and the nearest wastewater treatment plant.

I also wanted to look at proximity of sample sites to water reclamation facilities (WRFs). Let's try that. 
# load WRF location
```{r}
library(here)
library(geodist) #to calculate distance from sample site to each water reclamation facility
library(dplyr) #for data processing
library(tidyverse)

#load MP data
mpdata <- readRDS(here("data", "processed_data", "popdata.rds"))
#path to data
wrfdata_location <- here::here("data","processed_data","wrflocation.rds")

#load data
wrflocation <- readRDS(wrfdata_location)
```

Eventually, I would like to be able to view this information on a map - where WRFs are labelled, and samples sites are color-coded by density of microplastics found. For now, let's see if we can calculate distance otherwise.

```{r}
#calculate distance between sample coordinates and WRF location
wrfdistance <- geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat)
```

I think the product we have here is a matrix of the distance of each sample site from each of the three WRFs, in the order that they appear in the WRF data (Cedar Creek, North Oconee, Middle Oconee). But, the labelling of the data is not super clear, so I need to verify this to be sure.

```{r}
test <- mpdata %>% mutate(wrfdist = geodist_vec(mpdata$long, mpdata$lat, wrflocation$long, wrflocation$lat))
```

This produced another column like I wanted, but I don't know which WRF the distance was calculated for. Maybe I need to save WRF locations as individual lat and long values, and produce three new columns that way? However, I am not sure what the final product would look like if we do that, because a sample site could be closely associated with one WRF but much further from another one. Maybe I need to narrow this down to proximity to the nearest WRF, so that each sample site is only associated with the distance to ONE WRF, not three.

#Working on gathering WRF distance
```{r}
#set individual WRF locations and save as values
cedarcreeklat = 33.87351 
cedarcreeklong = -83.33639
northoconeelat = 33.93402
northoconeelong = -83.36018
middleoconeelat = 33.90862
middleoconeelong = -83.39214

mpdata <- mpdata %>% mutate(ccwrf = geodist_vec(mpdata$long, mpdata$lat, cedarcreeklong, cedarcreeklat))
mpdata <- mpdata %>% mutate(nowrf = geodist_vec(mpdata$long, mpdata$lat, northoconeelong, northoconeelat))
mpdata <- mpdata %>% mutate(miwrf = geodist_vec(mpdata$long, mpdata$lat, middleoconeelong, middleoconeelat))

#Since I only care about the proximity to the closest WRF, my next task is to write code that creates a new variable returning only the lowest value out of ccwrf, nowrf, and miwrf for each sample site.

#subset data into just distance from each WRF
wrf <- mpdata %>% select("ccwrf", "nowrf", "miwrf", "site", "date") 
#find minimum distance to any WRF out of all three
val <- do.call(pmin, c(wrf, na.rm = TRUE)) 
minwrf <- transform(wrf, min=val, minCol = names(wrf)[max.col(wrf == val, 'first')])

#minwrf contains the variable min, which tells us the distance from the sample site to the nearest WRF. This is all of the information that we need, so now we will re-join to the original DF
mpwrfdata <- left_join(mpdata, minwrf)

#remove variables we won't need anymore
mpwrfdata <- mpwrfdata %>% select(-"ccwrf", 
                                  -"nowrf",
                                  -"miwrf")
#rename columns
mpwrfdata <- mpwrfdata %>% rename("dist" = "min", 
                     "wrf" = "minCol")

#save as RDS
saveRDS(mpwrfdata, file = here("data", "processed_data", "mpwrfdata.rds"))
```

Update 10/25/21
I've updated the population data by accessing census tract level population rather than zip code level population. This should give us more localized population levels and might paint a better picture of the relationship between MP particles per liter and local population level.

#Adding WRF data to updated population data
```{r}
#load data
tractpopdata <- readRDS(here("data", "processed_data", "tractpopdata.rds"))

#add wrf data to tractpopdata
wrfjoin <- mpwrfdata %>% select("site", "date", "wrf", "dist") #selecting just the variables we want to add
mpwrfcombo <- left_join(tractpopdata, wrfjoin, by = c("site", "date")) #joining the two datasets

glimpse(mpwrfcombo)
#dist variable is saved as character when it needs to be numeric
#convert dist from character to numeric
mpwrfcombo <- mpwrfcombo %>% mutate(dist = as.numeric(dist))

#ungroup df
mpwrfcombo <- ungroup(mpwrfcombo)

#save as RDS
saveRDS(mpwrfcombo, file = here("data", "processed_data", "mpwrfcombo.rds"))
```

Now, `mpwrfcombo` contains the sample info and the new WRF distance info. However, something to work on going forward is that a sample site may be near a WRf but not downstream of it. It might be more important to know the distance to the nearest upstream WRF. I'll have to go back and consider this as I continue my exploration/analysis.

Other than not including landcover data, `mpwrfcombo` is the most comprehensive dataset, including almost all of the data that I have compiled so far.

### 2021-12-09 products/manuscript/Manuscript.Rmd
Clean up

@@
 bibliography: library.bib 
-csl: apa.csl
+csl: elsevier.csl
 ---
@@
 Through community-based sample collection and analysis, volunteers quantified microplastics through membrane filtration and visual identification.
-Preliminary data consistently show that local freshwater microplastic levels are greater than the U.S.
-average in the Upper Oconee Watershed, a portion of the Oconee River basin with both a high population size and high population density relative to other regions of Georgia.
-Average and median concentrations of 95 particles/L and 46 particles/L, respectively, were recorded throughout one year of seasonal sampling.
-Further visual and statistical analysis will confirm the project’s results, and discussion of these results in context will reveal the efficacy of this community-based approach to the quantification of microplastics.
+Preliminary data consistently show that freshwater microplastics are present in the local Upper Oconee Watershed, a portion of the Oconee River basin with both a high population size and high population density relative to other regions of Georgia.
+Average and median concentrations of 104 particles/L and 66 particles/L, respectively, were recorded throughout one year of seasonal sampling.
+Additionally, this project assesses the relationship between microplastic levels and hypothesized predictors of pollution, finding no present correlations between microplastic levels and population, land cover, wastewater treatment, or general water quality.
+Primarily, this project establishes the presence of microplastics in the region and demonstrates the use of citizen science, community based efforts for advances in science and monitoring, while further studies may continue to investigate predictive factors.
 
@@
 This preliminary study sought to identify the presence of microplastics in the Upper Oconee Watershed, estimate the levels of microplastic pollution present in the freshwater systems of Athens, GA, observe temporal changes and trends, and explore the relationship between microplastic pollution and hypothesized predictors including population level, land use, bacteria level, and wastewater treatment plant proximity (@Murphy2016).
-Thus, the key questions that we ask in this study are: what are microplastic levels like in and aroudn Athens, GA?
-Do microplastic levels vary by season?
-Does local population level, local wastewater treatment, land cover, or general water quality influence microplastic levels?
+Thus, the key questions that we ask in this study are:
+
+ - What are microplastic levels like in and around Athens, GA? 
+ - Do microplastic levels vary by season? 
+ - Does local population level, local wastewater treatment, land cover, or general water quality influence microplastic levels?
+
 We hypothesized that microplastics would be present in the Upper Oconee Watershed, that levels would be greater in more highly-trafficked areas with greater anthropogenic impact, and that microplastic pollution would correlate to other indicators of water quality [@McCormick2014a].
@@
 The Upper Oconee Watershed is located in Northeast Georgia, spanning from Gainesville on the northern end to Lake Sinclair in Milledgeville, Georgia, on the southern end.
-The Upper Oconee Watershed encompasses 7590 km2, and includes more than 4000 km of streams [@Fisher].
+The Upper Oconee Watershed encompasses 7590 km^2, and includes more than 4000 km of streams [@Fisher].
 Land within the Upper Oconee Watershed is historically agricultural, but has been subject to increasing urbanization in the past two decades; the study area contains developed and densely populated urban residential areas [@Cho; @Fisher].
-The Oconee Watershed is one of 14 river basins in the state of Georgia; the Oconee River feeds into the Altamaha River, which eventually deposits into the Atlantic Ocean off the Georgia coast (source: UOWN website).
+The Oconee Watershed is one of 14 river basins in the state of Georgia; the Oconee River feeds into the Altamaha River, which eventually deposits into the Atlantic Ocean off the Georgia coast.
 The Oconee River begins where the North and Middle Oconee Rivers meet, south of Athens, Georgia.
@@
 Quarterly sampling events occur throughout the year, and the present research project was made possible via collaboration and eventual partnership with UOWN and the quarterly sampling events.
-The UOWN study area consists of 600 km2.
+The UOWN study area consists of 600 km^2.
 
@@
 ### UOWN Quarterly Data
-This study relies on bublic data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org.
+This study relies on public data from Upper Oconee Watershed Network, Athens, Georgia; retrieved from www.uown.org.
 UOWN supplied information on samples site names, coordinates, and corresponding watersheds during primary data collection.
@@
 The data was cleaned to adjust variable format, calculate microplastic concentration averages, subset sample data and control data, and remove missing observations.
-Observations with missing location (coordinates, watershed) data were removed, as well as observations which were missing count values for microplastic concentration.
+Observations with missing location data (coordinates, watershed) were removed, as well as observations which were missing count values for microplastic concentration.
 UOWN data was cleaned to remove variables with incomplete observations and variables uninformative for analysis.
@@
 Beyond the bivariate analysis, a full model was built to incorporate all predictors in modeling microplastic concentration.
-Three modeling techniques were performed in order to perform a multivariate analysis and subset selection: decision tree, LASSO, and random forest.
+Three modeling techniques were implemented in order to perform a multivariate analysis and subset selection: decision tree, LASSO, and random forest.
 A null model is used to compare model performance and therefore interpret variable importance.
@@
 ## Summary and Interpretation
-In this analysis, we examined the distribution of freshwater microplastics in Athens, GA, as well as the relationship between the main outcome, microplastic concentration, and a few key hypothesized predictors: population level, microbial water quality, land cover, and proximity to wastewater treatment.
+In this analysis, we examined the distribution of freshwater microplastics in Athens, GA, as well as the relationship between the main outcome, microplastic concentration, and a few key hypothesized predictors: population level, microbial water quality, land cover, and proximity to water reclamation facilities.
 Microplastic levels were monitored over a period of one year, with water samples being collected and processed quarterly in November 2020, February 2021, April 2021, and July 2021.
@@
 However, the body of research on microplastics as a whole reflects mixed evidence of concentrations.
-Recorded levels in the literature range from 1 or fewer particles per liter at some sites to greater than 1000 particles per liter at others.
-There is growing evidence that microplastic pollution in permeates ecosystems in unexpected, supposedly pristine regions across the globe; one study found between 22 and 117 particles per liter in the Antarctic, demonstrating that microplastic particles are being found in places you would never expect.
+Recorded levels in the literature range from 1 or fewer particles per liter at some sites to greater than 1000 particles per liter at others [@Li2018;@Tibbetts2018].
+There is growing evidence that microplastic pollution permeates ecosystems in unexpected, supposedly pristine regions across the globe; one study found between 22 and 117 particles per liter in the Antarctic, demonstrating that microplastic particles are being found in places you would never expect [@Waller2017].
 
@@
 Limitations to the scope of this project include that it was designed as a student-led pilot with minimal funding.
-Visual identification via dissecting microscope was the method used due to its simplicity and accessibility; visual identification is used in most baseline microplastic studies, but can be up to 70% inaccurate and therefore is usually accompanied by spectroscopy technology which allows chemical identification of suspected microplastic particles and increasing the accuracy of quantification results.
+Visual identification via dissecting microscope was the method used due to its simplicity and accessibility; visual identification is used in most baseline microplastic studies, but can be up to 70% inaccurate and therefore is usually accompanied by spectroscopy technology which allows chemical identification of suspected microplastic particles and increasing the accuracy of quantification results [@Primpke2018].
 
@@
 Therefore, this study accomplishes exploratory aims and meets the growing demand for freshwater microplastics research and monitoring.
+
 ## Conclusions
@@
 Continued research is required in order to identify sources of microplastic pollution so that pollution levels may be mitigated in the future.
-Areas of focus for the future include furthering temporal, routine data collection, refining quantification methods, and increasing funding for microplastic research technologies.
+Areas of focus for the future include furthering  routine and temporal data collection, refining quantification methods, and increasing funding for microplastic research technologies.
 

### 2021-12-09 products/manuscript/Supplement.Rmd
Clean up

@@
   + "**processing_code**"
+    + processing_code.Rmd is the master script for all processing. It combines all of the scripts from below, which have been moved to the `ARCHIVE` folder. 
       + censusland_code.Rmd combines land cover data with microplastic data
       + censusprocessing.Rmd imports census tract population data
-    + map_images.Rmd combines individual map images to produce a grid
       + mpprocessing.R processes the original microplastic data file and saves cleaned data
@@
       + wrfprocessing.Rmd imports distance for nearest water reclamation facility
+    + map_images.Rmd combines individual map images to produce a grid
   + "**readme.md**" provides an order and brief description of dependencies/outputs for processing code files
@@
 
-1. mpprocessing.R
-2. uownprocessing.Rmd
-3. censusprocessing.Rmd
-4. censusland_code.Rmd
-5. wrfprocessing.Rmd
-6. exploration.Rmd
+1. processing_code.Rmd
 
@@
 
-7. analysis.Rmd
-8. modeling.Rmd
-9. land_model.Rmd
-10. map_images.Rmd
+2. analysis.Rmd
+3. modeling.Rmd
+4. land_model.Rmd
+
+In the code > processing_code folder:
+
+5. map_images.Rmd
 
@@
 
-11. Manuscript.Rmd
-12. Supplement.Rmd
+6. Manuscript.Rmd
+7. Supplement.Rmd
 

### 2021-12-09 code/processing_code/ARCHIVE/mpprocessing.R

@@ -1,0 +1,136 @@
+###############################
+# processing script
+#
+#this script loads the raw data, processes and cleans it 
+#and saves it as Rds file in the processed_data folder
+
+#load  packages
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing
+library(here) #to set paths
+library(tidyverse)
+
+#path to data
+data_location1 <- here::here("data","raw_data","mpdata_ALL.xlsx")
+data_location2 <- here::here("data","raw_data","uown 8.26.xlsx")
+
+#load microplastics data. 
+rawdata1 <- read_excel(data_location1)
+rawdata2 <- read_excel(data_location2)
+
+#take a look at the data
+glimpse(rawdata1)
+glimpse(rawdata2)
+
+#view the dataset on screen
+print(rawdata1)
+print(rawdata2)
+
+#we have 11 variables in the first dataset, and 18 in the second. These variables
+#should correspond so that we can combine the datasets
+
+rawdata <- full_join(rawdata1, rawdata2)
+
+# now, we have a combined dataset to work with
+
+glimpse(rawdata)
+
+# the notes column won't be neccessary for data analysis. Columns 16, 17, and 18 
+# are also unnecessary. Let's remove those. 
+
+processeddata <- rawdata %>% select(-"notes", -(17:19))
+
+# assessing missing values
+is.na(processeddata) %>% summary()
+
+# Since there are just a few observations that are missing location (watershed, lat, and long)
+# and/or count data, and there is not a great way to impute them. Since I am working
+# with data that I collected myself, I know that those observations simply didn't
+# get counted, got lost, or site names were misread/input incorrectly. Let's clean 
+# up the data set. 
+
+processeddata <-subset(processeddata, site!="MIDO 504" & site!="MIDO 503")
+
+# I've removed the rows by name instead of filtering out rows with missing watershed information.
+# This is because I want to retain the "control" rows for now, which originated 
+# in the lab and don't have a corresponding sample location
+
+# Another thing I can do is tidy the data by filling in all "id" information (all missing ids for previous
+# samplings would be me (a)). 
+
+processeddata$id[is.na(processeddata$id)]<-"a"
+
+# Next, let's remove observations where the samples were originally recorded in the
+# data sheet, but apparently never got counted during sample processing. Again, 
+# we just want to remove sample type observations and not control type observations. 
+
[…]

### 2021-12-09 code/processing_code/mpprocessing.R

@@ -1,136 +1,0 @@
-###############################
-# processing script
-#
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed_data folder
-
-#load  packages
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing
-library(here) #to set paths
-library(tidyverse)
-
-#path to data
-data_location1 <- here::here("data","raw_data","mpdata_ALL.xlsx")
-data_location2 <- here::here("data","raw_data","uown 8.26.xlsx")
-
-#load microplastics data. 
-rawdata1 <- read_excel(data_location1)
-rawdata2 <- read_excel(data_location2)
-
-#take a look at the data
-glimpse(rawdata1)
-glimpse(rawdata2)
-
-#view the dataset on screen
-print(rawdata1)
-print(rawdata2)
-
-#we have 11 variables in the first dataset, and 18 in the second. These variables
-#should correspond so that we can combine the datasets
-
-rawdata <- full_join(rawdata1, rawdata2)
-
-# now, we have a combined dataset to work with
-
-glimpse(rawdata)
-
-# the notes column won't be neccessary for data analysis. Columns 16, 17, and 18 
-# are also unnecessary. Let's remove those. 
-
-processeddata <- rawdata %>% select(-"notes", -(17:19))
-
-# assessing missing values
-is.na(processeddata) %>% summary()
-
-# Since there are just a few observations that are missing location (watershed, lat, and long)
-# and/or count data, and there is not a great way to impute them. Since I am working
-# with data that I collected myself, I know that those observations simply didn't
-# get counted, got lost, or site names were misread/input incorrectly. Let's clean 
-# up the data set. 
-
-processeddata <-subset(processeddata, site!="MIDO 504" & site!="MIDO 503")
-
-# I've removed the rows by name instead of filtering out rows with missing watershed information.
-# This is because I want to retain the "control" rows for now, which originated 
-# in the lab and don't have a corresponding sample location
-
-# Another thing I can do is tidy the data by filling in all "id" information (all missing ids for previous
-# samplings would be me (a)). 
-
-processeddata$id[is.na(processeddata$id)]<-"a"
-
-# Next, let's remove observations where the samples were originally recorded in the
-# data sheet, but apparently never got counted during sample processing. Again, 
-# we just want to remove sample type observations and not control type observations. 
-
[…]

### 2021-12-13 reviews/Project_Review_Brent Cameron.Rmd

---
title: Project Review Template 
date: "`r file.mtime(knitr::current_input())`"
#bibliography: ../media/references.bib
output: 
  html_document:
    toc_depth: 3
    number_sections: true
---

# Overview

Title of project: Plastic in the Urban Freshwater Environment

Name of project author(s): Amelia Foley

Name of project reviewer: Brent Cameron

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments
Great job on your explanation of the background relating to micro-plastics, I 
enjoyed reading about your description of the history of micro-plastics and 
their sources, as well as how primary and secondary micro-plastics relate 
to one another. The description of waste water treatment plants as a major source
of secondary micro-plastic pollution was well done.

I was impressed with how many sources you cited, definitely loos like you've 
done your research.

I think it would be beneficial to talk more in depth about motivation in the 
project, you mention why micro-plastics are dangerous and bad for the environment,
but I think adding a little about the potential benefits of reducing 
micro-plastics to go along with why they are dangerous would make the project 
overall feel more well-rounded.

Also, it could be beneficial to mention one or two of your sources directly when
discussing research to add credibility without needing to access citations directly (in abstract for instance).

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
I believe you did a good job in explaining what you project involves, especially
in regards to what the project involves (collect relevant data on micro-plastics)
and how this data was collected (community-based sample collection...). However,
I believe that you could go a little more in depth (in abstract), about what 
questions in particular you are trying to answer, as well as how those questions
relate to the data.

You state what the project is and what it does (established presence of micro-plastics), but if you also added some information on your main hypothesis
being asked (in the abstract), I think that could help in immediate understanding when reading your project. I know you go more in depth on 2.1.4 Present Research,
but even adding just your main question to the abstract could be good for 
understanding.

### Summary assessment
* question/hypotheses somewhat clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
You do a good job describing the data you're using as well as where it comes from
and the methods they are using to create the data (vacuum filtration, using Whatman, etc.). I also like how instead of having a general data section, you
include all supplemental data in its' own section for ease of access and understanding along with citations to the respective organizations.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
The data cleaning that you did was well explained, and every additional
step you did to create more pertinent data had sufficient explanation as to why
you felt that it was necessary or needed.

For your exploratory analysis you did a good job of explaining what your main
outcome was as well as your methods you used to determine statistical 
significance, the only thing I would suggest is to include all variables used 
in your analysis (or maybe even data cleaning) section to allow for further 
understanding when reading your manuscript.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The analysis methods used seem to be appropriate for your data and the analysis
seems to have been done properly. I think it would be good to include more 
information in your analysis about why you chose the variables you did, as 
well as why you decided to choose RMSE for your performance measure (I know you talked about it in 4.3 Full Analysis, but I think a bit of info would be helful here as well). Doesn't have to be a ton of information but just a few lines could add a lot here.
(analysis section of manuscript)

I know you include what the variables are under 4.2 Bivariate Analysis, but 
I think a few lines identifying what they are before hand would be good.

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
The results were very well presented, the tables and figures were easy to read and
understand, in particular I really found the seasonal maps interesting (although a little difficult to read) and was really impressed with them. In addiition,
the Land cover matrix was an interesting chart and was more exciting to look at 
then traditional linear regression charts (looks good too).

I also like how you make sure to specify why you chose the LASSO model and mentioned the RMSE of each model as comparisons.
### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
I thought it was cool that this research establishes a baseline of the presence of micro-plastics in the Athens, GA area for the future. I also think that you did 
a good job describing conclusion (nothing can be confirmed via the analysis
but stating how micro-plastics do contaminate even in remote places). I also
think you did well going into more detail about strength and limitations, including the fact that the research is a student-led pilot and the accuracy of the methods used (as well as traditional methods accompanying research).

### Summary assessment
* discussion is complete and well done

## Further comments
Overall I am really impressed with your project, everything looks really professional and there are a large number of cited sources. Tables and charts are (for the most part- maps are a little hard to read), easily read and understood
and are beneficial to understanding research made.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
The structure of the project is well done and replicating is easy with all steps
adequately explained in read.me on github, as well as in the actual paper itself

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments
The project is documented well, there is ample explanation for why each step of the analysis was done the way it was, and the files do a suitable job at explaining the process and reasons behind it.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments
Project is completely reproducible without any manual intervention

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments
Great job in explaining the process of the research and what your goal was,
you could possibly elaborate more in your conclusion section (or even a bit before), about possible alternative methods that could be used or measures of association besides RMSE that may be suitable.

### Summary assessment
* decent level of thoroughness

## Further comments
I don't have any further comments than what I have already written, great project,
really impressed with the whole thing.

### 2022-03-30 code/analysis_code/results.Rmd
adding results
volunteer data and CV

---
title: "results"
output: html_document
---

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)

mp <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
sampledata <- readRDS(here("data", "processed_data", "fullsampledata.rds"))
```

# find CV for first three sample dates (within subject A -> B)
```{r}
sample_three <- sampledata %>% filter(!date == "july_21") # remove July samples (counted by volunteers, not by self)

sample <- sample_three %>% select(a_count, b_count)
matrix_sample <- data.matrix(sample)

sample <- sample %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))

mean(sample$cv)
median(sample$cv)
```

Woah - CVs for samples counted by me personally are averaging 20 - 26 for CV%. This is quite different from the volunteer counts. What does this say about the precision of volunteer counted-data?

# CV vs MP concentration
```{r}
# this was performed in the Barrows paper - looks at how variation correlates with concentration - previous findings showed that high-concentration sites had less variability among replicates

# create dataframe for all sites with CV values and concentration values
# average A and C and B and D counts so that they can be compared to the other sample dates for variation
avg_cd <- sampledata %>% filter(date == "july_21") %>% mutate(a_count = ((a_count+c_count)/2)) %>% mutate(b_count = ((b_count+d_count/2)))

# combine averaged july data with other three seasons of data
all_counts_cv <- full_join(avg_cd, sample_three)

# keep only useful variables
all_counts_part <- all_counts_cv %>% select(a_count, b_count, particles_l, date)
all_counts_cv <- all_counts_cv %>% select(a_count, b_count) # don't include particles/L so it won't be calculated with CV

# create matrix
all_counts_matrix <- data.matrix(all_counts_cv)
# mutate new variable to add CV
all_counts_cv <- all_counts_cv %>% mutate(cv = (Rfast::rowcvs(all_counts_matrix, ln = FALSE)*100))
# add MP concentration back in using join
all_counts_cv <- left_join(all_counts_cv, all_counts_part)
```
# plot MP concentration vs CV
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

Wonder if I can make this graph and have point colors correlate to sample dates. That would be interesting
# Adding color to graph
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv, color = date)) + geom_point() + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

There appears to be a negative correlation between CV and microplastic concentration, which is consistent with what's reported in the Barrows paper

## Next, let's do a Kruskal Wallis test (basically a non-parametic one-way ANOVA, that considers more than two groups)
```{r}
kruskal.test(particles_l ~ date, data = all_counts_cv)
```

P=value less than 0.05 - we conclude that there IS a significant difference between groups (though we don't yet know between what groups). From here, we may use the pairwise.wilcox.test() to calculate pairwise comparison
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(all_counts_cv$particles_l, all_counts_cv$date,
                 p.adjust.method = "BH")
```

The only significant difference is between April and February sample dates. (in Barrows, March had the highest concentration)

Could perform regression analysis to compare original samples and corresponding duplicates over 1 year study period (using this language from the Barrows paper)
(or does this mean between A and B duplicates?) Let's try
# linear regression for duplicates over study period
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit_dup <- lm_mod %>% fit(a_count ~ b_count, data = all_counts_cv)
#save table
lm_fit_dup_table <- tidy(lm_fit_dup)
#view summary of fit
(lm_fit_dup)
summary(lm_fit_dup)
#save file
saveRDS(lm_fit_dup_table, file = here("results", "lm_fit_dup_table.rds"))

model = lm(a_count ~ b_count, data = all_counts_cv)
summary(model)
```

p-value is <0.05 and multiple r-squared value is 0.7824
p-value indicates that the model reflects a signficant relationship
r-squared is "degree to which the data is explained by the model" --> 78% of variability is explained by the model

```{r}
#visualize linear model
lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count")
lm_dup
```

It appears that the duplicate counts are pretty well correlated.

# want to test for significant difference within a group to see if sites are signficantly different/if the distribution of MP concentrations is signficant
```{r}
kruskal.test(particles_l ~ site, data = sampledata)
```

Seen in Barrows paper - other researchers found concentrations at 70 - 100 particles/L

Try to make plot to compare sites over seasons
```{r}
sampledata %>% filter(date == "nov_20") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

```{r}
# feb 2021
sampledata %>% filter(date == "feb_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# april 2021
sampledata %>% filter(date == "apr_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# july 2021
sampledata %>% filter(date == "july_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

It could be interesting to have a table listing all of the site, and the concentration for each sample date in columns to the side. this might be easier to interpret than a graph for this type of data. 
```{r}
table <- sampledata %>% select(particles_l, site, date)

table %>% rowwise()
reshape(sampledata, idvar = "site", timevar = "date", direction = "wide")
```
```{r}
table$date <- factor(table$date , levels=c("nov_20", "feb_21", "apr_21", "july_21")) #reorder dates to appear chronologically
# transform to wide data
data_wide <- spread(table, date, particles_l)
data_wide #display new data
# display decimal places upto 1
data_wide$nov_20<-format(round(data_wide$nov_20,1),nsmall=1)
data_wide$feb_21<-format(round(data_wide$feb_21,1),nsmall=1)
data_wide$apr_21<-format(round(data_wide$apr_21,1),nsmall=1)
data_wide$july_21<-format(round(data_wide$july_21,1),nsmall=1)
```

goals: 
- look at differences in average watershed concentration over time

- test for difference between watershed groups
```{r}
#difference between watershed groups
kruskal.test(particles_l ~ watershed, data = sampledata)
```

P-value is greater than 0.05, there is NOT a signficant difference between watershed concentrations

#avg watershed concentration over time
```{r}
watershed_time <- group_by(sampledata, watershed, date) %>%
             summarise(particles_l = mean(particles_l, na.rm = TRUE))
kruskal.test(particles_l ~ date, data = watershed_time)

barber_creek <- watershed_time %>% filter(watershed == "Barber Creek")
kruskal.test(particles_l ~ date, barber_creek)

hunnicutt_creek <- watershed_time %>% filter(watershed == "Hunnicutt Creek")
kruskal.test(particles_l ~ date, hunnicutt_creek)

north_oconee_river <- watershed_time %>% filter(watershed == "North Oconee River")
kruskal.test(particles_l ~ date, north_oconee_river)

```

### 2022-03-30 code/analysis_code/volunteer.Rmd
adding results
volunteer data and CV

---
title: "volunteer_analysis"
output: html_document
---

# VOLUNTEER ANALYSIS#
Here, I will compare the counts performed by two separate volunteers on the same set of samples, so evaluate the use of citizen science volunteers to precisely count MPs (precision > accuracy)

# load packages and data
```{r}
library(tidyverse)
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(MASS) #for Wilcoxon signed rank test
library(Rfast) # for column-wise coefficient of variation

data <- readRDS(here("data", "processed_data", "volunteer.rds"))
```

I want to compare a_count and c_count, and b_count and d_count. I could do this using a paired t-test for matched pairs. Paired t-test can only be used for normal distribution. This can be checked with Shapiro-Wilk test. 
# Shapiro-wilk test for normality
```{r}
shapiro.test(data$a_count)
shapiro.test(data$b_count)
shapiro.test(data$c_count)
shapiro.test(data$d_count)
```

P-value < 0.05, meaning that each variable IS significantly different from a normal distribution.

Instead, we can use the non-parametric Wilcoxon signed rank test

# Wilcoxon signed rank
```{r}
# comparing the A filter by a_count and c_count
wilcox.test(data$a_count, data$c_count, paired=TRUE) 
# comparing the B filter by b_count and d_count
wilcox.test(data$b_count, data$d_count, paired=TRUE) 
```

## A filter
P-value > 0.05, we can conclude that a_count and c_count are NOT significantly different; we FAIL to reject the null hypothesis

## B filter
P-value > 0.05, we can conclude that b_count and d_count are NOT significantly different; we FAIL to reject the null hypothesis

# Is the total particles/L count impacted by small differences in individual counts? 
```{r}
# Let's test particles_l and rep_particles_l to be sure
wilcox.test(data$particles_l, data$rep_particles_l, paired=TRUE) 
```

Here, we see that the p-value still exceeds 0.05, indicating that there is not a signficant difference between the replicates.

Let's also use the coefficient of variation to evaluate the difference in volunteer counts of the same filters.
# Coefficient of variation
```{r}
# let's make a subset of data, so that we can do some row-wise calculations\
subset_a <- data %>% dplyr::select(a_count, c_count)
subset_b <- data %>% dplyr::select(b_count, d_count)

matrix_a <- data.matrix(subset_a)
matrix_b <- data.matrix(subset_b)

# calculate cv row-wise
rowcvs(matrix_a, ln = FALSE)
rowcvs(matrix_b, ln = FALSE)
```

I think we need to multiply these values by 100 to achieve CV %
Ideally, we'd have these as another column in our dataset for easy reading. Let's see if we can use mutate to do this. 
```{r}
subset_a <- subset_a %>% mutate(cv = (rowcvs(matrix_a, ln = FALSE)*100))
subset_b <- subset_b %>% mutate(cv = (rowcvs(matrix_b, ln = FALSE)*100))

#average CV for A filters
mean(subset_a$cv)
#average CV for B filters
mean(subset_b$cv)

#median CV for A filters
median(subset_a$cv)
#median CV for B filters
median(subset_b$cv)

```

Here, we see that the average CV lies between approximately 50-60%. This is a little high. We also want to look at the median to account for possible outliers (a few rogue volunteers)

We see that the medians fall between 43 and 60. This is not too different from our means. This may indicate substantial variation in results between replicate counts.

## we can also calculate CV based on replicate concentration (rather than indvidual filters)
```{r}
conc_rep <- data %>% dplyr::select(particles_l, 
                            rep_particles_l)
matrix_c <- data.matrix(conc_rep)
conc_rep <- conc_rep %>% mutate(cv = (rowcvs(matrix_c, ln = FALSE)*100))

mean(conc_rep$cv)
median(conc_rep$cv)
```

Here, we see similar CV values at 56 (mean) and 60 (median) - this indicates that the comparison of calculated concentrations may be representative of the comparison between individual filter set replicates.

Could also be helpful to calculate CV of variation between all counts for July volunteer data. We could then comapre this CV to the CV for my own counts from the three prior sample dates. 
# Cumulative CV
```{r}
cumul_cv <- data %>% dplyr::select(a_count, b_count, c_count, d_count)

matrix_cumul <- data.matrix(cumul_cv)

cumul_cv <- cumul_cv %>% mutate(cv = (rowcvs(matrix_cumul, ln = FALSE)*100))

mean(cumul_cv$cv)
```

The average overall coefficient of variation is 58 for volunteer-counted results.

### 2022-03-30 code/processing_code/processing_code.Rmd
adding results
volunteer data and CV

@@
 #merging mutated data
-cleansampledata <- full_join(otherdata, julydata) %>% select(-"id",
+fullsampledata <- full_join(otherdata, julydata) # to preserve a and b counts
+cleansampledata <- fullsampledata %>% select(-"id",
                                                          -"id2",
@@
 save_data_location2 <- here::here("data","processed_data","controldata.rds")
+save_data_location3 <- here::here("data", "processed_data", "fullsampledata.rds")
 
@@
 saveRDS(controldata, file = save_data_location2)
+saveRDS(fullsampledata, file = save_data_location3)
 

### 2022-04-08 code/analysis_code/particle_type.Rmd
new figures

---
title: "particle_type"
output: html_document
---

```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(readxl) #for loading Excel files

type <- read_excel(here("data", "raw_data", "copy_part.xlsx"))
```

# cleaning data
```{r}
# this data set is a litte larger than we need. let's remove some unneccesary variables
a_and_b <- type %>% select("fragment", 
                           "fiber", 
                           "bead", 
                           "foam", 
                           "pellet",
                           "film",
                           "fragment_b", 
                           "fiber_b", 
                           "bead_b", 
                           "foam_b", 
                           "pellet_b", 
                           "film_b")

c_and_d <- type %>% select("fragment_c", 
                           "fiber_c", 
                           "bead_c", 
                           "foam_c", 
                           "pellet_c",
                           "film_c",
                           "fragment_d", 
                           "fiber_d", 
                           "bead_d", 
                           "foam_d", 
                           "pellet_d", 
                           "film_d")
summary(a_and_b)
```

```{r}
#summary for a and b counts
colSums(a_and_b, na.rm = TRUE)

#summary for c and d counts
colSums(c_and_d, na.rm = TRUE)
```

```{r}
part <- type %>% select("fragment", 
                           "fiber", 
                           "bead", 
                           "foam", 
                           "pellet",
                           "film",
                           "fragment_b", 
                           "fiber_b", 
                           "bead_b", 
                           "foam_b", 
                           "pellet_b", 
                           "film_b", 
                        "fragment_c", 
                           "fiber_c", 
                           "bead_c", 
                           "foam_c", 
                           "pellet_c",
                           "film_c",
                           "fragment_d", 
                           "fiber_d", 
                           "bead_d", 
                           "foam_d", 
                           "pellet_d", 
                           "film_d")
part[is.na(part)] <- 0 # replace NA with 0

part <- part %>% mutate(fragment_avg_a = ((fragment + fragment_c)/2)) 
part <- part %>% mutate(fragment_avg_b = ((fragment_b + fragment_d)/2)) 
part <- part %>% mutate(fiber_avg_a = ((fiber + fiber_c)/2)) 
part <- part %>% mutate(fiber_avg_b = ((fiber_b + fiber_d)/2))
part <- part %>% mutate(bead_avg_a = ((bead + bead_c)/2))  
part <- part %>% mutate(bead_avg_b = ((bead_b + bead_d)/2))     
part <- part %>% mutate(foam_avg_a = ((foam + foam_c)/2)) 
part <- part %>% mutate(foam_avg_b = ((foam_b + foam_d)/2)) 
part <- part %>% mutate(pellet_avg_a = ((pellet + pellet_c)/2))
part <- part %>% mutate(pellet_avg_b = ((pellet_b + pellet_d)/2)) 
part <- part %>% mutate(film_avg_a = ((film + film_c)/2))
part <- part %>% mutate(film_avg_b = ((film_b + film_d)/2))

sum <- part %>% select(25:36)
colSums(sum)

sum <- sum %>% mutate(fragment = fragment_avg_a + fragment_avg_b, 
                      fiber = fiber_avg_a + fiber_avg_b, 
                      bead = bead_avg_a + bead_avg_b, 
                      foam = foam_avg_a + foam_avg_b, 
                      pellet = pellet_avg_a + pellet_avg_b, 
                      film = film_avg_a + film_avg_b) %>% select(13:18)

colSums(sum)
# this gives us a small dataframe with just the total for each particle type
part_sum <- enframe(colSums(sum))
```

Now, we have a breakdown of total particle type (added between A and B filters). A and B counts were determined by averaging the duplicate volunteer counts performed on singular filters.

Let's see how we can visualize this breakdown. Remember that this is just for the July sample date, not for all 4 sample dates (we are looking at a subset).

```{r}
part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
```

Perhaps we could add error bars, significant difference indications in this graph.

### 2022-04-08 particle_type.Rmd
new figures

--- 2022-04-08 code/analysis_code/particle_type.Rmd
@@
 ```{r}
+
 part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
-  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
+  geom_col() + geom_text(aes(label = percent((value/559)), #total particles is 559
+                             vjust = -0.5, #above bar
+                             position = position_dodge(width=0.9)) + ylim(0, 350) #makes 60% visible
 ```

### 2022-04-10 code/analysis_code/particle_type.Rmd
color updates, editing figures

@@
 part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
-  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
+  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5)) + ylim(0,350) + guides(fill=guide_legend("Particle Type")) + labs(y = "Microplastic Composition")
+#position = position_dodge(width=0.9))) + ylim(0, 350) 
 ```
@@
 Perhaps we could add error bars, significant difference indications in this graph.
+
+#particle color breakdown (aggregate, for July only)
+```{r}
+# load particle color data
+color <- read_excel(here("data", "raw_data", "mp_color.xlsx"))
+
+# plot color data
+color %>% ggplot(aes(x = color, y = number, fill = color)) + 
+  geom_col() + geom_text(aes(label = percent(number/968), vjust = -0.5)) + ylim(0,420) + scale_fill_manual(values = c("black" = "gray0", "blue" = "royalblue2", "green" = "seagreen2", "other" = "yellow1", "red" = "red1", "transparent/white"= "white")) + labs(y = "Occurence") + guides(fill=guide_legend("Particle Color"))
+
+#stacked
+
+#reorder
+color$color <- reorder(color$color, color$number) #largest to smallest
+# plot reordered stacked plot
+ggplot(data = color, aes(x = obs, y = number, fill = color)) + geom_bar(stat = "identity", position = "fill") + scale_fill_manual(values = c("black" = "gray0", "blue" = "royalblue2", "green" = "seagreen2", "other" = "yellow1", "red" = "red1", "transparent/white"= "white")) + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+ # geom_text(aes(label = paste0(percent(number/968))), position=position_stack(vjust=0.5), colour="white", size = 2)
+
+ # here is some code for adding % labels, but I can't get it to work without messing up the graph                                                       #geom_text(aes(label=paste0(sprintf("%1.1f", percent*100),"%")),
+                   #  position=position_stack(vjust=0.5), colour="white", size = 2)  
+```

### 2022-04-10 code/analysis_code/results.Rmd
color updates, editing figures

@@
 library(dplyr)
+library(gridExtra) #for savings images in grids
+library(gridExtra) #for savings images in grids
 
@@
 #visualize linear model
-lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count")
+lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count") + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + scale_x_log10() + scale_y_log10()
 lm_dup
@@
 It appears that the duplicate counts are pretty well correlated.
+
+We could use this same type of graph to display the comparative CV for my counts vs volunteer counts by doing AvsB for sample dates 1-3 and then AvsB for sample date 4 or AvsC and BvsD
+
+#visualizing CV
+```{r}
+#visualize linear model
+#subset data
+all_counts_me <- all_counts_cv %>% filter(!date == "july_21") 
+#plot
+lm_me <- ggplot(all_counts_me, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Duplicate Counts Nov, Apr, July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
+lm_me #show plot
+
+#subset data
+all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
+#plot
+lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Volunteer Counts July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
+lm_vol #show plot
+```
+
+# CV for each sample date
+```{r}
+#nov
+nov <- all_counts_cv %>% filter(date == "nov_20")
+#plot
+lm_nov <- ggplot(nov, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Nov", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + expand_limits(x = 0, y = 0)
+lm_nov #show plot
+
+#feb
+feb <- all_counts_cv %>% filter(date == "feb_21")
+#plot
+lm_feb <- ggplot(feb, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Feb", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
+lm_feb #show plot
+
+#april
+april <- all_counts_cv %>% filter(date == "apr_21")
+#plot
+lm_apr <- ggplot(april, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "April", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
+lm_apr #show plot
+
+#july
+all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
+#plot
+lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "July", x = "A Count", y = "B Count") + ylim(0,25) + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + xlim(3,25) + ylim(3,30)
+lm_vol #show plot
+
+bivar_grid <- grid.arrange(lm_nov, lm_feb, lm_apr, lm_vol, ncol = 2)
+bivar_grid #show grid
+```
 
@@
 ```
+# main rivers vs tributaries
+```{r}
+river <- sampledata %>% filter(watershed == "North Oconee River"|watershed == "Middle Oconee River"|watershed == "Oconee River")
+
+trib <- sampledata %>% filter(!watershed == "North Oconee River", !watershed == "Middle Oconee River", !watershed == "Oconee River")
+
+river <- river %>% mutate(body = "river")
+trib <- trib %>% mutate(body = "trib")
+
+#rejoin
+body <- full_join(river, trib)
+
+#plot
+body %>% ggplot(aes(x = body, y = particles_l, color = )) + 
+  geom_boxplot()  +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10()
+
+#kruskal wallis/mann whitney test - nonparametric test
+kruskal.test(particles_l ~ body, data = body)
+```
+
+P=value less than 0.05 - we conclude that there IS a significant difference between groups. Let's do a pairwise.wilcox.test() to double check. 
+# pairwise wilcox test to see differences between groups
+```{r}
+pairwise.wilcox.test(body$particles_l, body$body,
+                 p.adjust.method = "BH")
+```
+
+This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps.

### 2022-04-10 code/analysis_code/density.Rmd
figures + calculations

---
title: "density"
output: html_document
---

```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(tigris)
library(sp)

options(tigris_year = 2019)
```

```{r}
df <- tracts(state = 'GA', county = c('Clarke'), )
plot(df)
```
The GEOID in the df dataframe matches our FIPS code for the population/sample data. Let's add the land area variable to the tractpopdata dataframe and then mutate a new variable to calculate population density. Then, we can compare population density to MP concentration.

```{r}
pop <- readRDS(here("data", "processed_data", "tractpopdata.rds"))

area <- df %>% select(GEOID, ALAND)

combine <- left_join(pop, area, by = c("FIPS" = "GEOID"))

#calculate pop density
combine <- combine %>% mutate(density = population/ALAND)

# plot MP concentration vs pop density

combine %>% ggplot(aes(x = density, y = particles_l)) + geom_point() +
  labs(title = "Microplastic Concentration vs Population Density", x = "Population Density", y = "Particles/Liter") + scale_y_log10() + scale_x_log10()

```

# use linear regression to confirm relationship
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit <- lm_mod %>% fit(particles_l ~ density, data = combine)
#save table
lm_fittable <- tidy(lm_fit)
#view summary of fit
(lm_fit)
summary(lm_fit)
#save file
#saveRDS(lm_fit_dup_table, file = here("results", "lm_fit_dup_table.rds"))

model = lm(particles_l ~ density, data = combine)
summary(model)
```

#calculate 60th percentile
```{r}
quantile(sampledata$particles_l, .60)
```

### 2022-04-10 code/analysis_code/results.Rmd
figures + calculations

@@
 This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps. 
+# more figures for main rivers vs tribs
+```{r}
+body %>% ggplot(aes(x = date, y = particles_l, color = body, group = body)) + 
+  geom_point() + geom_path() + scale_color_discrete(name = "Water Body")+ 
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(title = "Water Body Microplastic Concentration Over Time", x = "Water Body", y = "Particles/Liter") + scale_x_discrete(labels=c("Fall 2020",
+                            "Winter 2021",
+                            "Spring 2021",
+                            "Summer 2021"))+ 
+  ylim(0,300) #removing outliers to better visualize majority of concentrations
+```

### 2022-04-10 code/analysis_code/volunteer.Rmd
figures + calculations

@@
 The average overall coefficient of variation is 58 for volunteer-counted results. 
+# CV bw A and B for July
+```{r}
+sample_july <- sampledata %>% filter(date == "july_21") # remove July samples (counted by volunteers, not by self)
+
+july <- sample_july %>% select(a_count, b_count)
+matrix_sample <- data.matrix(july)
+
+july <- july %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))
+
+mean(july$cv)
+median(july$cv)
+```
+
+#CV for each sample date
+```{r}
+#november
+sample_nov <- sampledata %>% filter(date == "nov_20")
+
+nov <- sample_nov %>% select(a_count, b_count)
+matrix_sample <- data.matrix(nov)
+
+nov <- nov %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))
+
+mean(nov$cv)
+median(nov$cv)
+
+#february
+sample_feb <- sampledata %>% filter(date == "feb_21")
+
+feb <- sample_feb %>% select(a_count, b_count)
+matrix_sample <- data.matrix(feb)
+
+feb <- feb %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))
+
+mean(feb$cv)
+median(feb$cv)
+
+#april
+sample_april <- sampledata %>% filter(date == "apr_21") # remove July samples (counted by volunteers, not by self)
+
+april <- sample_april %>% select(a_count, b_count)
+matrix_sample <- data.matrix(april)
+
+april <- april %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))
+
+mean(april$cv)
+median(april$cv)
+```

### 2022-04-11 code/analysis_code/density.Rmd
sunday updates

@@
 ```{r}
-df <- tracts(state = 'GA', county = c('Clarke'), )
-plot(df)
+df_clarke <- tracts(state = 'GA', county = c('Clarke'))
+plot(df_clarke)
+
+df <- tracts(state = 'GA', county = c('Clarke', 'Oconee'))
 ```
@@
 ```
+
+#land cover distribution
+```{r}
+land_long <- athensland %>% pivot_longer(12:26)
+
+land_long %>% ggplot(aes(x = name, y = value)) + geom_col() + 
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```

### 2022-04-11 code/analysis_code/particle_type.Rmd
sunday updates

@@
 ```
+# particle type by water body
+```{r}
+ggplot(data = part_sum, aes(x = body, y = value, fill = name)) + geom_bar(stat = "identity", position = "fill") + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+```

### 2022-04-11 code/analysis_code/results.Rmd
sunday updates

@@
 library(gridExtra) #for savings images in grids
+library(ggpubr) #adding stats to figures
 
@@
 Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
+```
+
+#adding stats to graph
+```{r}
+ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() +
+  stat_cor(method = "spearman")
+
+ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(title = "Coefficient of Variance vs.
+Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") +
+  stat_cor(method = "spearman") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
+
+df %>% ggplot(aes(x = x , y = y)) +
+  geom_point() +
+  stat_cor(method = "spearman")
+
+  geom_smooth(method = lm) + labs(title = "Coefficient of Variance vs.
+Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
+
+figure_s1_data %>% ggplot(aes(x = cases.reported.7dma/1.2, y = value)) +
+  geom_point() +
+  facet_grid(variable~target, scales = "free",  labeller = label_wrap_gen()) +
+  stat_cor(method = "spearman") +
+  ylab("Value") +
+  xlab("Reported Cases per 100K (7-Day Moving Average)")
+
 ```
@@
   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10()
+  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10() +
+  stat_cor(method = "spearman")
 
@@
 kruskal.test(particles_l ~ body, data = body)
+wilcox.test(particles_l ~ body, data = body)
+
+summary(river)
+summary(trib)
 ```

### 2022-04-12 code/analysis_code/exploration.Rmd
sunday updates
quad figure 1

@@
 
+# stats for land use, p-value for land use
+```{r}
+ggplot(data = land_long, aes(x = value, y = particles_l)) + geom_point(size = .5) + facet_wrap(~name) + stat_cor(method = "spearman") + labs(x = "Total Area (sq.
+meters)", y = "Particles per Liter") + theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
 In this plot, we are mainly looking at the row containing particles/L (the very bottom row). We do not see any strong correlations between microplastic level and local land use.

### 2022-04-12 code/analysis_code/particle_type.Rmd
quad figure 1

--- 2022-04-08 code/analysis_code/particle_type.Rmd
@@
 library(readxl) #for loading Excel files
+library(ggpubr)
 
@@
 
-sum <- sum %>% mutate(fragment = fragment_avg_a + fragment_avg_b, 
-                      fiber = fiber_avg_a + fiber_avg_b, 
-                      bead = bead_avg_a + bead_avg_b, 
-                      foam = foam_avg_a + foam_avg_b, 
-                      pellet = pellet_avg_a + pellet_avg_b, 
-                      film = film_avg_a + film_avg_b) %>% select(13:18)
+sum <- sum %>% mutate(Fragment = fragment_avg_a + fragment_avg_b, 
+                      Fiber = fiber_avg_a + fiber_avg_b, 
+                      Bead = bead_avg_a + bead_avg_b, 
+                      Foam = foam_avg_a + foam_avg_b, 
+                      Pellet = pellet_avg_a + pellet_avg_b, 
+                      Film = film_avg_a + film_avg_b) %>% select(13:18)
 
@@
 ```{r}
-part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
-  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
+part_sum %>% ggplot(aes(x = name, y = value)) + 
+  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5)) + ylim(0,350) + guides(fill=guide_legend("Particle Type")) + labs(y = "Microplastic Composition")
+#position = position_dodge(width=0.9))) + ylim(0, 350) 
 ```
@@
 Perhaps we could add error bars, significant difference indications in this graph.
+
+#PARTICLE TYPE FIGURE
+```{r}
+fig_part_type <- part_sum %>% ggplot(aes(x = name, y = value)) + 
+  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5), size = 3) + ylim(0,380) + guides(fill=guide_legend("Particle Type")) + labs(y = "No.
+of Particles", x = "") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.text.y=element_text(size=12))
+plot(fig_part_type)
+
+```
+
+#PARTICLE COLOR FIGURE
+```{r}
+library(ggrepel)
+# load particle color data
+color <- read_excel(here("data", "raw_data", "mp_color.xlsx"))
+
+# plot reordered stacked plot
+color <- color %>%  mutate(color= str_to_title(color)) 
+color <- color %>% rename(Color = color)
+color$Color <- reorder(color$Color, color$number) #largest to smallest
+
+# plot color data
+fig_part_color <- color %>% ggplot(aes(x = Color, y = number, fill = Color)) + 
+  geom_col() + geom_text(aes(label = percent(number/968), vjust = -0.5), size = 3) + ylim(0,460) + scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "TRN/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") + labs(y = "No.
+of Particles")  + theme_bw() + theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
+plot(fig_part_color)
+#stacked
+#reorder
+
+ggplot(data = color, aes(x = obs, y = number, fill = Color)) + geom_bar(stat = "identity", position = "fill") + scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "Transparent/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
+   geom_text(aes(label = percent(number/968)), size = 3, color = "white", position = position_fill(vjust = 0.5)) + theme_bw()+ theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+plot(fig_part_color)
+
+```
+# particle type by water body
+```{r}
+ggplot(data = part_sum, aes(x = body, y = value, fill = name)) + geom_bar(stat = "identity", position = "fill") + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+```

### 2022-04-12 code/analysis_code/results.Rmd
quad figure 1

--- 2022-04-10 code/analysis_code/results.Rmd
@@
 library(here) #for data loading/saving
+library(tidyverse)
 library(tidymodels) #for modeling
@@
 
+#adding stats to graph
+```{r}
+# with line of best fit
+ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(x = "Particles per Liter", y = "Coefficient of Variance %") +
+  stat_cor(method = "spearman") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points
+
+#no line of best fit
+ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + labs(x = "Particles per Liter", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points
+```
+
 Wonder if I can make this graph and have point colors correlate to sample dates. That would be interesting
@@
                  p.adjust.method = "BH")
+
+compare_means(particles_l ~ date, data = all_counts_cv, p.adjust.method = "BH")
 ```
@@
 
+#rename
+body$body <- fct_recode(body$body,
+                                 "Main Rivers" = "river",
+                                 "Tributaries" = "trib")
+pv <- body %>% summarize(p.value = round(kruskal.test(particles_l ~ body)$p.value, digits = 3))
+
 #plot
@@
   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10()
+  labs(x = "Water Body", y = "Particles per Liter") + ylim(0,500) + scale_y_log10() +
+    geom_text(data=pv, aes(x=1.5, y=900, label=paste0("Kruskal-Wallis\n p=",p.value))) + theme_bw()
 
@@
 This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps. 
+# figure with stats for seasons
+```{r}
+compare_means(particles_l ~ date,  data = body)
+my_comparisons <- list(c("nov_20", "feb_21"), c("nov_20", "apr_21"), c("nov_20", "july_21"), c("feb_21", "apr_21"), c("feb_21", "july_21"), c("apr_21", "july_21"))
+
+fig_season <- body %>% ggplot(aes(x = date, y = particles_l, group = date)) + 
+  geom_boxplot() + scale_color_discrete(name = "Water Body")+ 
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  labs(x = "Sample Date", y = "Particles per Liter") + scale_x_discrete(labels=c("Fall 2020",
+                            "Winter 2021",
+                            "Spring 2021",
+                            "Summer 2021"))+ theme_bw() + stat_compare_means(comparison=my_comparisons, p.adjust.method = "BH") + stat_compare_means(label.y = 5) + scale_y_log10()
+plot(fig_season)
+```
+#arrange figures
+```{r}
+#map
+#Load Libraries 
+library(tidyverse)
+library(viridis)
+library(ggpubr)
+#Load data
+sites = readRDS(here::here("data","processed_data","mpwrfcombo.rds"))
+sites$date = as.factor(sites$date)
+sites$date = recode(sites$date, nov_20 = "Nov 2020", feb_21 = "Feb 2021", apr_21 = "April 2021", july_21 = "July 2021")
+sites$date = ordered(sites$date, levels = c("Nov 2020", "Feb 2021", "April 2021", "July 2021"))
+sites %>% ggplot(aes(x = date, y = site, fill= particles_l)) + 
+  geom_tile() + 
+  scale_fill_viridis() + 
+  theme_bw() +
+  ylab("Site") + 
+  xlab("Collection Period") + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_text(aes(x = date, y = site, label = round(particles_l))) +
+  labs(fill = "Particles per Liter")  
+  #ggtitle("Seasonality of Microplastics Concentration")
+sites %>% ggplot(aes(x = date, y = site, fill= log10(particles_l))) + 
+  geom_tile() + 
+  scale_fill_viridis() + 
+  theme_bw() +
+  ylab("Site") + 
+  xlab("") + 
+  scale_x_discrete(position = "top") +
+  #theme(axis.text.x = element_text(angle = 45)) +
+  geom_text(aes(x = date, y = site, label = round(log10(particles_l), 2))) +
+  labs(fill = "Log10(Particles/Liter)") + 
+  theme(legend.position = "bottom")
+  #ggtitle("Seasonality of Microplastics Concentration")
+hist_all = sites %>% ggplot(aes(x = particles_l)) + 
+  geom_histogram() + 
+  theme_bw() + 
+  #facet_wrap(~date) +
+  xlab("Particles per Liter") + 
+  ylab("No. of Observations") +
+  theme(axis.text.y=element_text(size=12))
+hist_season = sites %>% ggplot(aes(x = particles_l)) + 
+  geom_histogram() + 
+  theme_bw() + 
+  facet_wrap(~date) +
+  xlab("Particles per Liter") + 
+  ylab("No. of Observations") +
+  theme(axis.text.y=element_text(size=12))
+fig2 = ggarrange(hist_all, hist_season, ncol = 2)
+plot(fig2)
+sites %>% group_by(date) %>% tally()
+
+#particle type = fig_part_type
+#particle color = fig_part_color
+quad1 <- ggarrange(hist_all, hist_season, fig_part_type, fig_part_color, nrow = 2, ncol = 2, labels = "AUTO", vjust = 1) 
+#%>% ggexport(filename = here("results","manuscript_figures", "quad_1.png"))
+plot(quad_1)
+
+ggsave(filename = here("results","manuscript_figures", "quad_1.png")) #export resolution than ggexport
+```

### 2022-04-13 code/analysis_code/particle_type.Rmd
fig updates

@@
 fig_part_type <- part_sum %>% ggplot(aes(x = name, y = value)) + 
-  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5), size = 3) + ylim(0,380) + guides(fill=guide_legend("Particle Type")) + labs(y = "No.
-of Particles", x = "") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.text.y=element_text(size=12))
+  geom_col() + 
+  geom_text(aes(label = percent(value/559), vjust = -0.5), size = 3) + 
+  ylim(0,380) + 
+  guides(fill=guide_legend("Particle Type")) + 
+  labs(y = "No. of Particles", x = "") + 
+  theme_bw() + 
+  theme(axis.text.x = element_text(angle = 45, hjust=1), 
+        axis.text.y=element_text(size=12)) + 
+  annotate("text", x = 5, y = 300, label = "n = 559")
 plot(fig_part_type)
@@
 fig_part_color <- color %>% ggplot(aes(x = Color, y = number, fill = Color)) +
-  geom_col() + geom_text(aes(label = percent(number/968), vjust = -0.5), size = 3) + ylim(0,460) + scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "TRN/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") + labs(y = "No.
-of Particles")  + theme_bw() + theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust=1))
+  geom_col() + 
+  geom_text(aes(label = percent(number/968), vjust = -0.5), size = 3) + 
+  ylim(0,300) + 
+  scale_fill_manual(values = c("Black" = "gray0", "Blue" = "royalblue2", "Green" = "seagreen2", "Other" = "orange1", "Red" = "red1", "TRN/White"= "gray61")) + labs(x = "", y = "Microplastic Composition") +
+  labs(y = "No. of Particles")  + 
+  theme_bw() + 
+  theme(legend.position = "none")+ 
+  theme(axis.text.x = element_text(angle = 45, hjust=1), 
+        axis.text.y = element_text(size = 12)) + 
+  annotate("text", x = 2, y = 200, label = "n = 484")
 plot(fig_part_color)
@@
         axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
+        axis.ticks.x=element_blank(), 
+        axis.text.y = element_text(size = 3)) 
 plot(fig_part_color)

### 2022-04-13 code/analysis_code/results.Rmd
fig updates

@@
 #%>% ggexport(filename = here("results","manuscript_figures", "quad_1.png"))
-plot(quad_1)
+plot(quad1)
 
